---
title: "Processing Recipes"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The most recent version of this file and data can be found at
```
git clone https://github.com/PeterClaussenSDSU/JoyOfCooking2018.git
```

# List Recipes

The first step to create a recipe database is to process individual recipe files. The files structure was chosed to allow a group of individuals to contribute, and allow individual recipes to be added to the whole, independent of any coordinating files or tables. These files will be stored in:
```{r}
path.to.recipes <- "~/github/JoyOfCooking2018/RecipeTables"
```


We'll make no assumptions about file types. The expected suffix is `.tab`, but we won't assume that all files conform. If we can guarantee all files conform to a specific format, we might use
```{r,eval=FALSE}
recipe.files <- list.files(path = path.to.recipes, pattern = '*.tab')
```

Instead, we'll attempt to read all files in the table directory.

```{r}
recipe.files <- list.files(path = path.to.recipes)
length(recipe.files)
```

The original specification was for three data columns, and two were to be added by processing file names. A fourth data column was later added, but we won't assume all files conform.

```{r}
expected.columns <- c("Amount","Measure","Ingredient")
added.columns <- c("NDB_No","Year","Recipe")
```

For processing, we'll have have a single organizing structures - a list of tables that have been read from file. We'll use file name for indexes.

```{r}
recipe.list <- vector("list", length(recipe.files))
names(recipe.list) <- recipe.files
```

## Utility functions

### Minimum headers

Does a list of names include the minimum required header names?

```{r}
# returns true if the list of names has the minimum required headers
minimum.headers <- function(names) {
  for(name in expected.columns) {
    if(!(name %in% names)) {
      #found a missing identifier
      return(FALSE)
    }
  }
  #all were found
  return(TRUE)
}
```

**To do** A complementary function may be implemented to use partial matching and return the list of names, in the order they match the input names. This returned list can be used to map columns in data to the expected column order. This function may return an empty list if matches are not made.

### Parse File Name

Later, we'll need parse the file name into a recipe name and a date. This is a placeholder for that function.

```{r}
#place holder for file name processing function
parse.filename <- function(name) {
  return(list(Year=0,Recipe=name))
}
```

# Read Files

We'll try to read each file by first guessing the delimiter. For the time being, I'm assuming either tab-delimited or space-delimited text. 

We'll assume the first row is a header row, and that it contains only single words, so we can use `strsplit` to token the row into names. If that doesn't work, we'll try with spaces. 

For the files that don't have conforming first lines, we print the file name and header line for further inspection.

```{r}
conforming.files <- 0
nonconforming.files <- c()

for(recipe in recipe.files) {
  file.name <- paste(path.to.recipes,'/',recipe,sep='')
  current <- NULL
  #read the first line
  connection <- file(file.name)
  first.line <- readLines(connection,n=1)
  close(connection) #R won't be happy if we don't close the file
  
  #if the first line can be split by tabs, then use tabs as delimiter
  first.tokens <- strsplit(first.line,split='\t')[[1]]
  #some files will be written with quoted text. 
  #The various `read` functions will strip the quotes, 
  #but here we have to do this manually.
  first.tokens <-gsub('\"','', first.tokens)
  if(minimum.headers(first.tokens)) {
    #read as tab-delimited 
    recipe.list[[recipe]] <- read.table(file.name,
                                        sep='\t',
                                        stringsAsFactors = FALSE,
                                        header=TRUE)
    conforming.files <- conforming.files+1
  } else {
    #second guess is spaces.
    first.tokens <- strsplit(first.line,split=' ')[[1]]
    first.tokens <-gsub('\"','', first.tokens)
    if(minimum.headers(first.tokens)) {
      recipe.list[[recipe]] <- read.table(file.name,
                                          header=TRUE,
                                          stringsAsFactors = FALSE)
      conforming.files <- conforming.files+1
    } else {
      #save a list for later
      nonconforming.files <- c(nonconforming.files,recipe)
      print(recipe)
      print(first.line)
    }
  }
}
conforming.files
```

**To do** Determine how to partially match data columns to read non-conforming files.

# Combining Recipes

Our final table will have 6 columns. We'll create a enpty data frame to build from, and keep an empty copy to test the join for each individual recipe table.

```{r}
join.names <- c(expected.columns, added.columns)
join.frame <-data.frame(matrix(vector(), nrow=0, ncol=length(join.names),
                dimnames=list(c(), join.names)),
                stringsAsFactors=F)
test.frame <- join.frame
```


## Fix column names.

There are some simple inconsistencies in the column names that prevent merging. One simple one to correct is NBD vs NDB.

```{r}
for(idx in 1:length(recipe.files)) {
  file.name <- recipe.files[idx]
  tbl <- recipe.list[[file.name]]
  if('NBD_No' %in% names(tbl)) {
    nbd.idx <- which(names(tbl)=='NBD_No')
    names(tbl)[nbd.idx] <- added.columns[1]
    recipe.list[[file.name]] <- tbl
  }
}
```

**To do** Partial matches for other mispellings.


Some recipes already have `Year` and `Recipe` columns. If we assume these have correct information, we can join directly. If not, we parse the file name and add `Year` and `Recipe` columns, using the previously defined utility function.


Some may not have `NDB_No`. We can add a dummy variable and attempt to match to Ingredients later.


```{r}
conforming.tables <- 0
nonconforming.tables <- c()

for(idx in 1:length(recipe.files)) {
  file.name <- recipe.files[idx]
  if(!(file.name %in% nonconforming.files)) {
    tbl <- recipe.list[[file.name]]
    if(length(names(tbl)) == 3) {
      #We tested against Ingredient,Amount and Measure, so append
      #an empty NBD_No. We will attempt to look up thos later.
      #if we sort the data by ingredient, and multiple recipes have the 
      #same incredients, we may be able to save some searches.
      tbl$NDB_No <- NA
    }
    if(length(names(tbl)) == 4) {
      # we assume, for now, that the table is missing Year and Recipe,
      # so parse the file name for this information.
      ids <- parse.filename(file.name)
      tbl$Year <- ids$Year
      tbl$Recipe <- ids$Recipe
    }
    test.tbl <- NULL
    #put the join in a try block. This will allow us to
    #process all files without having to stop and fix errors.
    tryCatch(test.tbl <- rbind(test.frame, tbl[,join.names]),
             error=function(e) {
               print(file.name)
               print(head(tbl))
           })
    #if we can bind to the empty table, we can merge this table with
    #the rest
    if(!is.null(test.tbl)) {
      join.frame <- rbind(join.frame,tbl[,join.names])
      conforming.tables <- conforming.tables + 1
    } else {
      nonconforming.tables <- c(nonconforming.tables,file.name)
    }
  }
}
conforming.tables
```

```{r}
summary(join.frame)
```

**To do** 
Iterate over non-conforming tables, determine which can be fixed programmatically.

