---
title: "Midterm Project"
author: "Set 14"
date: "July 21, 2018"
output: 
   pdf_document: default
---

I have gone through the Lond Description in the FOOD_DES,I try to find the best description that matches the description of each ingredient of the recipes in my dataset.Based on the matched description, I have added the column NDB_No to each recipe table.


To generate ChocolateCharlotte.2006 recipe data table:

```{r}
ChocolateCharlotte.1936 <- data.frame(
  Amount=c(1.00,0.25,0.75,1.00,2.00,1.00,0.88,1.00,18.00),
  Measure= c("envelope (1 tbsp)", "cup","cup","oz","cup, whipped","large","cup","tsp","ladyfinger"),
  Ingredient = c("Gelatins, dry powder, unsweetened", "Water, bottled, generic", "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D", "Chocolate, dark, 45- 59% cacao solids", "Cream, fluid, heavy whipping", "Egg, whole, raw, fresh", "Sugars, granulated","Vanilla extract","Cookies, ladyfingers, without lemon juice and rind"),
  NDB_No=c(19177,14555,1211,19902,1053,1123,19335,02050,18423)
  )

ChocolateCharlotte.1936
```


To generate ChocolateCharlotte.2006 recipe data table: In this data, I have combined the "sugars,granulated" three vaules using the following conversions: 1 cup = 16 table spoon, and 1 table spoon = 3 tea spoon.

```{r}
ChocolateCharlotte.2006<- data.frame(
  Amount=c(18.00,0.56,0.44,0.44,0.50,6.00,3.00,1.00,0.50),
  Measure= c("ladyfinger", "cup","cup","cup","envelope (1 tbsp)","oz","tbsp","large","cup, whipped"),
   Ingredient = c("Cookies, ladyfingers, without lemon juice and rind","Sugars, granulated","Water, bottled, generic","Beverages, coffee, brewed, prepared with tap water","Gelatins, dry powder, unsweetened","Chocolate, dark, 45- 59% cacao solids","Butter, salted","Egg, whole, raw, fresh","Cream, fluid, heavy whipping"),
  NDB_No=c(18423,19335,14555,14209,19177,19902,1001,1123,1053)
)

ChocolateCharlotte.2006
```

To generate MoldedCustard.1936 recipe data table:

```{r}
MoldedCustard.1936<- data.frame(
  Amount=c(1.00,1.75,2.00,0.33,1.00),
  Measure= c("envelope (1 tbsp)","cup","large","cup","tsp"),
  Ingredient = c("Gelatins, dry powder, unsweetened","Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D", "Egg, whole, raw, fresh", "Sugars, granulated","Vanilla extract"),
  NDB_No=c(19177,1077,1123,19335,02050)
)

MoldedCustard.1936
```
To generate MoldedCustard.2006 recipe data table:

```{r}
MoldedCustard.2006<- data.frame(
  Amount=c(3.00,0.75,0.50,3.00,1.00),
  Measure= c("cup","envelope (1 tbsp)","cup","large","tsp"),
  Ingredient = c("Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D","Gelatins, dry powder, unsweetened", "Sugars, granulated", "Egg, whole, raw, fresh","Vanilla extract"),
  NDB_No=c(1077,19177,19335,1123,02050)
)

MoldedCustard.2006
```


Writing the ChocolateCharlotte 1936 recipe:
```{r}
write.table(ChocolateCharlotte.1936,"ChocolateCharlotte.1936", sep="\t")
```


Writing the ChocolateCharlotte 1936 recipe:
```{r}
write.table(ChocolateCharlotte.1936,"ChocolateCharlotte.2006", sep="\t")
```


Writing the MoldedCustard 1936 recipe:

```{r}
write.table(ChocolateCharlotte.1936,"MoldedCustard.1936", sep="\t")
```


Writing the MoldedCustard 2006 recipe:

```{r}
write.table(ChocolateCharlotte.1936,"MoldedCustard.2006", sep="\t")
```

Reading the ChocolateCharlotte 1936:
```{r}
#getwd()
read.table("ChocolateCharlotte.1936")

```

Reading the ChocolateCharlotte 2006:

```{r}
read.table("ChocolateCharlotte.2006")
```


Reading the MoldedCustard 1936 recipe:

```{r}
read.table("MoldedCustard.1936")
```


Reading the MoldedCustard 2006 recipe:

```{r}
read.table("MoldedCustard.2006")
```


Reading the BeefStroganoff.1997 recipe:


```{r}
BeefStroganoff.1936<-read.table("BeefStroganoff.1997.tab" ,header=TRUE)
BeefStroganoff.1936

```


Reading the BeefStroganoff.2006 recipe:

```{r}
BeefStroganoff.2006<-read.table("BeefStroganoff.2006.tab",header=TRUE)
BeefStroganoff.2006
```


since the given files, BeefStroganoff.1997.tab and BeefStroganoff.2006.tab, do not have NDB_No. It would be diffcult to add combine them with the MoldedCustard 1936/2006 recipe and ChocolateCharlotte 1936/2006 recipe. Thus, let us add NDB_No to BeefStroganoff.1997 and 2006 recipes. Similar to what I have done above, I will use the best description matching approach to assign NDB_No to the ingredients.


Adding the column NDB_No to BeefStroganoff.1936:

```{r}

BeefStroganoff.1936$NDB_No=c(23336,1001,20082,06170,01074,02046)
BeefStroganoff.1936
```


Adding the column NDB_No to BeefStroganoff.2006:

```{r}
BeefStroganoff.2006$NDB_No=c(23336,04582,1001,11282,11238,06170,01074,02046)       
BeefStroganoff.2006
```

Combining all three recipes from above (for both 1936 and 2006), we get:

```{r}
combined_recipes<-rbind(ChocolateCharlotte.1936,ChocolateCharlotte.2006,MoldedCustard.1936,MoldedCustard.2006,BeefStroganoff.1936,BeefStroganoff.2006)

combined_recipes

```

I have proved the accuracy of my data and checked for errors using tables, tests and plots:

```{r}
datacheck1<-merge.data.frame(ChocolateCharlotte.1936,ChocolateCharlotte.2006,by = "NDB_No")
datacheck1
all(ChocolateCharlotte.1936$Measure.x== ChocolateCharlotte.2006$Measure.y)
```


From the above table and the test for equality, we can see that the measurement units for each NDB_No in both 1936 and 2006 data are matched.


Similarly, MoldedCustared recipes:

```{r}
datacheck2<-merge.data.frame(MoldedCustard.1936,MoldedCustard.2006,by = "NDB_No")
datacheck2
all(MoldedCustard.1936$Measure.x== MoldedCustard.2006$Measure.y)
```
From the above table and the test for equality, we can see that the measurement units for each NDB_No in both 1936 and 2006 data are matched.



To check data, let us plot the amount.1936 and amount.2006 for ChocolateCharlotte recipes:


```{r}
datacheck1<-merge.data.frame(ChocolateCharlotte.1936,ChocolateCharlotte.2006,by = "NDB_No")
#datacheck
plot(datacheck1$Amount.x,datacheck1$Amount.y)
```



To check data, let us plot the amount.1936 and amount.2006 for MoldedCustared recipes:

```{r}
datacheck2<-merge.data.frame(MoldedCustard.1936,MoldedCustard.2006,by = "NDB_No")
#datacheck
plot(datacheck2$Amount.x,datacheck2$Amount.y)
```

For both recipes, the plots of the amount look reasonable since they are close to each other.


Reading the Recipes.csv:


```{r}
Recipe_Table<-read.table("Recipes.csv" ,header=TRUE,sep = ",")

Recipe_Table
```


Inputting the Maximum and Minimum Servings 

```{r}
Add_data<- data.frame(
  Recipe=c("ChocolateCharlotte","ChocolateCharlotte","MoldedCustard","MoldedCustard"),
  Year= c(1997,2006,1997,2006),
  MinServings = c(6,8,6,8),
  MaxServings=c(6,8,6,8)
)
Add_data
```

Combining rows from Recipes.csv containing BeefStroganoff, ChocolateCharlotte, and MoldedCustard:


```{r}
Recipe_comb<-rbind(Recipe_Table, Add_data)
Recipe_comb
```



Writing the modified Recipes.csv out:


```{r}
write.table(Recipe_comb,"Recipes.csv", sep=",")
```

Reading the Recipes.csv in:

```{r}
Recipes.csv<-read.table("Recipes.csv" ,header=TRUE,sep = ",")
Recipes.csv
```











