---
title: "Joy Of Cooking Midterm Project Prep"
author: "Set 28"
date: ''
output:
  pdf_document: default
  html_document: default
---

# Part 1 - Assign Recipes

First, generate a list of all files in the Recipes folder.

### This list has all the names of the recipes provided.

```{r,eval=FALSE}
recipes <- list.files(path = "./", pattern = '*.pdf')
head(recipes)
length(recipes)
```

Make this reproducible.
```{r,eval=FALSE}
set.seed(10010)
```

Create two randomly sampled list of recipe files and cycle this list through 80 rows

### The data length for this part is project is 107, not 80. The nrow in code had to be changed to reflect the correct number.

```{r,eval=FALSE}
recipe.pairs <- matrix(c(sample(recipes),sample(recipes)),ncol=2,nrow=107)
head(recipe.pairs)
```

Are there any pairs across rows?

```{r,eval=FALSE}
sum(recipe.pairs[,1]==recipe.pairs[,2])
```

Write to a table

```{r,eval=FALSE}
write.csv(recipe.pairs, file="Recipe.Table.csv",row.names = FALSE)

```

Read the table in and print:

### Created a Recipe.Table as instructed. And then imported its content to this Markdown.

```{r,eval=FALSE}
Recipe.Table <- read.csv("Recipe.Table.csv",header=FALSE,skip=1)
head(Recipe.Table, 5)
```

# Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.

### The code below was able to read the data witout any errors.
```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
#View(weight.dat)
summary(weight.dat)

```

```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
#write.table(food_des.dat,"food_dest.csv", sep = ",")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")

```
### I created an object to test the grep function below. The function can read through the vector and find the variable being queried. 

```{r}
Ingredient=c("Lobster meat, can or cooked","oil, olive","vinegar", "celery, fresth,diced", "Mayonnaise")
if(length(i <- grep("coo", Ingredient)))
  cat("'coo' appears at least once in\n\t", Ingredient, "\n")
i 
Ingredient[i]
```

### The above function did not work very well to identify the recipe's name. The end goal was to find a recipe name and find the `NDB_No`.

```{r}

Ingredient=food_des.dat$Long_Desc
if(length(i <- grep("Mayonnaise", Ingredient)))
  cat("Mayonnaise' appears at least once in\n\t", Ingredient, "\n")
i # 2 and 4
Ingredient[807]



#summary(food_des.dat)
APPNAPPTART <- grepl("^celery, fresth,diced", food_des.dat$Long_Desc)

#PMC too many row to print (~8000)
#food_des.dat[1:13]

```
### I tried using the function above to get a recipe name connected to the `NDB_No`, instead of doing it manually. Failed.
```{R}
Ingredient=food_des.dat$Long_Desc
if(length(i <- grep("lobster", Ingredient)))
  cat("Ingredient' appears at least once in\n\t", Ingredient, "\n")
i 
Ingredient[i]
Ingredient[4232 ]
Ingredient[4311 ]


#summary(food_des.dat)
APPNAPPTART <- grep("^celery, fresth,diced", food_des.dat$Long_Desc, value = T)
APPNAPPTART
#PMC - too many rows to print
#food_des.dat[1:13]
```

```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
#summary(nut_data.dat)
```
## 1. Create data tables for your assigned recipes, with columns Amount, Unit and Ingredient.

### Lobster Salad or Shrimp 1936

```{r}
lobsterSaladorShrimp.1936<-data.frame(Recipe=c("Lobster Salad or Shrimp","Lobster Salad or Shrimp", "Lobster Salad or Shrimp","Lobster Salad or Shrimp","Lobster Salad or Shrimp"),Year=c(1936, 1936, 1936, 1936,1936),
Amount=c(4,1,1,1,0.5),
Measure=c("oz", "tbs", "tbs", "cup", "cup"),
Ingredient=c("Lobster meat, can or cooked","oil, olive","vinegar", "celery, fresth,diced", "Mayonnaise"))#, NDB_No=c("NA","NA", "NA","NA","NA"))
lobsterSaladorShrimp.1936$NDB_No<-NA

write.table(lobsterSaladorShrimp.1936, "lobsterSaladorShrimp.1936.tab", sep = "\t", row.names = FALSE)
lobsterSaladorShrimp.1936<- as.data.frame(read.table("lobsterSaladorShrimp.1936.tab", header = T, sep = "", fill = TRUE))
lobsterSaladorShrimp.1936
```

### Created a data frame for Lobster Salad or Shrimp 1936

```{r}
lobsterSaladorShrimp.2006<-data.frame(Recipe=c("Lobster Salad or Shrimp","Lobster Salad or Shrimp","Lobster Salad or Shrimp"),Year=c(2006, 2006, 2006),
Amount=c(8, .25, 1),
Measure=c("oz", "cup", "tsp"),
Ingredient=c("Lobster, cooked","sour cream, cultured", "juice, lemeon, fresh"))
lobsterSaladorShrimp.2006$NDB_No<-NA
lobsterSaladorShrimp.2006
write.table(lobsterSaladorShrimp.2006, "lobsterSaladorShrimp.2006.tab", sep = "\t")
lobsterSaladorShrimp.2006
lobsterSaladorShrimp.2006 <- as.data.frame(read.table("lobsterSaladorShrimp.2006.tab", header = T, sep = "", fill = TRUE))
lobsterSaladorShrimp.2006
```

### Created a data frame for LemonSouffle 1936
```{r}
lemonSouffle.1936<-data.frame(Recipe=c("Lemon Souffle", "Lemon Souffle","Lemon Souffle"),Year=c(1936,1936,1936),Amount=c(.75, 5, 3), Measure=c("cup", "lrg", "tbs"),Ingredients=c("sugar, white, granulated,granulated", "egg, raw", "juice,lemon,fresh") )

lemonSouffle.1936$NDB_No<-NA
write.table(lemonSouffle.1936, "lemonSouffle.1936.tab", sep = "\t")

lemonSouffle.1936
```
###  Write each individual recipe to a tab delimited file. The name of the tab file should be comprised of the recipe file name and the year of the recipe (1936 or 2006, in most cases).

Read in each tab file, and merge the tables with the example Beef Stroganoff tables. 
```{r}
lemonSouffle.1936 <- as.data.frame(read.table("lemonSouffle.1936.tab", header = T, sep = "", fill = TRUE))
lemonSouffle.1936
```

###LemonSouffle 2006
```{r}
lemonSouffle.2006<-data.frame(Recipe=c("Lemon Souffle", "Lemon Souffle","Lemon Souffle","Lemon Souffle", "Lemon Souffle","Lemon Souffle","Lemon Souffle", "Lemon Souffle"),Year=c(2006, 2006,2006,2006,2006,2006,2006,2006),
Amount=c(3,.25,5,1,.5,.5,.5,.33), 
Measure=c("tbs", "cup", "lrg","ea","cup","cup","cup","cup"), Ingredient=c("butter,salted", "flour, all purpose,self rising","egg, raw", "egg white, raw, large","milk, whole, 3.25%", "cream, whipping, heavy", "sugar, white, granulated, granulated", "juice, lemon, fresh"))
lemonSouffle.2006$NDB_No<-NA
```
###  Write each individual recipe to a tab delimited file. The name of the tab file should be comprised of the recipe file name and the year of the recipe (1936 or 2006, in most cases).

 Read in each tab file, and merge the tables with the example Beef Stroganoff tables. 
```{r}
write.table(lemonSouffle.2006, "lemonSouffle2006.tab", sep = "\t")
lemonSouffle.2006
lemonSouffle.2006 <- as.data.frame(read.table("lemonSouffle2006.tab", header = T, sep = "", fill = TRUE))
lemonSouffle.2006
```
### Recipe table for Min and Max Servings
```{r}
recipe<-data.frame(Recipe=c("Lobster Salad or Shrimp","Lobster Salad or Shrimp", "Lemon Souffle", "Lemon Souffle"), Year=c(1936,2006,1936,2006), MinServings=c(4,2,4,8), MaxServings =c(4,2,4,8))
recipe
```


### Merging all recipes. The idea is to create a large data set that could be split into subsection and using to mine the assignment objectives. 
- merge.recipe5 has data frames:
BeefStroganoff.1997
BeefStroganoff.2006
lobsterSaladorShrimp.1936
lobsterSaladorShrimp.2006
lemonSouffle.1936
lemonSouffle.2006


```{r}
BeefStroganoff.1997 <- as.data.frame(read.table("BeefStroganoff.1997.tab", header = T, sep = "", fill = TRUE))
BeefStroganoff.1997$YEAR<-c(1997, 1997, 1997,1997, 1997, 1997)
#PMC View shouldn't be used in typeset code
#View(BeefStroganoff.1997)
BeefStroganoff.2006 <- as.data.frame(read.table("BeefStroganoff.2006.tab", header = T, sep = "", fill = TRUE))
BeefStroganoff.2006$YEAR<-c(2006, 2006,2006, 2006,2006, 2006,2006, 2006)
#View(BeefStroganoff.2006)
merge.recipe<-merge(BeefStroganoff.1997, BeefStroganoff.2006, all = T)
#View(merge.recipe)
merge.recipe2<-merge(merge.recipe, lobsterSaladorShrimp.1936, all = T)
#View(merge.recipe2)
merge.recipe3<-merge(merge.recipe2, lemonSouffle.1936, all = T)
merge.recipe4<-merge(merge.recipe3  , lobsterSaladorShrimp.2006, all = T)
merge.recipe5<-merge(merge.recipe4, lemonSouffle.2006, all= T)
#View(merge.recipe5)


```

### Graph and Contingecy table for Beefstroganoff
```{r}
Beef.Merge<-merge(BeefStroganoff.1997, BeefStroganoff.2006, by="Ingredient", all.x  = T, all.y = T)
#View(Beef.Merge)
plot(Beef.Merge$Amount.x, Beef.Merge$Amount.y, type = "l", col="red")
table(Beef.Merge$Measure.x, Beef.Merge$Measure.y)


```
### This is function that list 4 data frames into one, then it merges the tables provided for this exercise.

- nut_data.dat

- food_des.dat

- weight.dat

# Again, the idea is to combined the NDB_No to the ingredients. In the end, it managed to merge all the data frames but I do not know with NDB_no's are assigned to the correct ingredient.
```{r}
Combine.Recipes<- function(x, y){
  df <- merge(x, y, by= "Ingredient", all.x=T, all.y=T )
  return(df)
}
new.df <-Reduce(Combine.Recipes, list(BeefStroganoff.1997,lobsterSaladorShrimp.1936, BeefStroganoff.2006, lemonSouffle.2006))



table(new.df$Amount.x, new.df$Amount.x)
table.1<-merge(new.df, weight.dat, all = T)

#this line crashes my system; it's far too large a file for all.x=T
#table.2<-merge(table.1, nut_data.dat, by="NDB_No", all.x=T )

#head(table.2)
#dim(table.2)







```

