---
title: "Stat 700 MIDterm"
author: "Set 29"
date: "July 17, 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---


# Part 1 - lets Assign Recipes


```{reval=FALSE}
getwd()
recipes = list.files(path = "./", pattern = '*.pdf')
length(recipes)
head(recipes)

```

Make this reproducible. 
```{r}
set.seed(10010) #To make the reproducable data
```


Create two randomly sampled list of recipe files and cycle this list through 109 rows.
```{r,eval=FALSE}
recipe.pairs <- matrix(c(sample(recipes),sample(recipes)),ncol=2,nrow=109)
length(recipe.pairs)
head(recipe.pairs)
```

Are there any pairs across rows?
2 pairs exist across the row. 
```{r,eval=FALSE}
sum(recipe.pairs[,1]==recipe.pairs[,2])
```

Write to a table. 
```{r,eval=FALSE}
write.table(recipe.pairs, file="Recipe.Table.tab")
```

Reading the table and printing: 

```{r,eval=FALSE}

Recipe.Table1 = read.table("Recipe.Table.tab",header = TRUE,stringsAsFactors = FALSE) #Read the .tab file
head(Recipe.Table1)
```

# lets do Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.
```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)
```

```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```

```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)
```


What am I doing?

My assigned recipes were Pineapple Snow and Coffee Bavarian Cream. I followed the same format of pdf for naming those recipes using the year number. I created CSV file using the Amount, Measure and Ingredient,name of recipes and year.
 

```{r}
# Now working on my Recipe
#Data frame for recipe1 == Pineapple Snow for years 1936 & 2006

### Pineapple Snow for year 1936 ################

PineappleSnow36 <- data.frame(
  Recipes = c("Pineapple Snow "),
  Year=c(1936),
  Amount = c(1.00, 0.25, 2.00, 1.00, 2.00),
  Measure = c("Tbs", "cup", "cup", "cup", "cup"),
  Ingredient = c("gelatin,dry,unsweetened", "water,generic", "pineapple,canned,crushed,in juice,drained", "sugar,white,granulated,granulated", "cream,whipping,heavy"))
PineappleSnow36[,-c(1,2)] #Taking the unwanted column out

### Pineapple Snow for year 2006 ################

PineappleSnow06 <- data.frame(
  Recipes = c("Pineapple Snow"),
  Year=c(2006),
  Amount = c(2.25, 0.25, 20.00, 1.00, 1.00, 0.50),
  Measure = c("tsp", "cup", "oz", "cup", "cup", "tsp"),
  Ingredient = c("gelatin,dry,unsweetened","water,generic","pineapple,canned,crushed,in juice", "sugar,white,granulated,granulated", "cream,whipping,heavy", "extract,vanilla"))
PineappleSnow06[,-c(1,2)] #Taking the unwanted column out

#Data table for recipie1 for year 1936 and 2006
PineappleSnow36
PineappleSnow06

#Writing in tab file
write.table(PineappleSnow36, file = "PineappleSnow36.tab") 
write.table(PineappleSnow06, file = "PineappleSnow06.tab")


#Merging table for recipie1 for year 1936 and 2006 by ingredent after taking the Name and Year column out
PineappleSnow36_06 = merge(PineappleSnow36[,-c(1,2)],PineappleSnow06[,-c(1,2)], by = "Ingredient" )
PineappleSnow36_06 

#Writing in tab file
write.table(PineappleSnow36_06, file = "PineappleSnow36_06.tab") 

#Checking the number of row of the recipe1 for year 1936 and 2006 indivisually
nrow(PineappleSnow36)
nrow(PineappleSnow06)

#Checking the number of row of the recipe1 of both year 1936 & 2006 after merging by ingrident 
nrow(PineappleSnow36_06)



#Finding the mismatching name if any
setdiff(PineappleSnow36$Ingredient, PineappleSnow06$Ingredient)


par(mfrow=c(1,2))
plot(as.factor(PineappleSnow36$Measure) ,PineappleSnow36$Amount,las = 3,xlab = "Measure(Unit)", ylab = "Amount",main = "1936")
plot(as.factor(PineappleSnow06$Measure) ,PineappleSnow06$Amount,las = 3,xlab = "Measure(Unit)", ylab = "Amount",main = "2006")

par(mfrow=c(1,2))
plot(as.factor(PineappleSnow36$Ingredient) ,PineappleSnow36$Amount,las = 3,xlab = "Ingredient", ylab = "Amount",main = "1936")
plot(as.factor(PineappleSnow06$Ingredient) ,PineappleSnow06$Amount,las = 3,xlab = "Ingredient", ylab = "Amount",main = "2006")

# The above plot shows that there are two units mismatch from year 1936 to 2006 and there is similarity of only "water,generic" Ingredient.

```




```{r}
#Data frame for recipe2 == Coffee Bavarian Cream  for years 1936 & 2006

### Coffee Bavarian Cream for year 1936 ################


CoffeeBavarianCream36 <- data.frame(
  Recipes = c("Coffee Bavarian Cream"),
  Year=c(1936),
  Amount = c(2.00, 1.00, 1.00, 1.00, 2.00, 2.00),
  Measure = c("Tbs", "cup", "cup", "cup", "ea", "cup"),
  Ingredient = c("gelatin,dry,unsweetened","milk,whole,3.25%","coffee,brewed,with tap water","sugar,white,granulated,granulated","egg white,raw,large", "cream,whipping,heavy"))
CoffeeBavarianCream36[,-c(1,2)]

### Coffee Bavarian Cream for year 2006 ################

CoffeeBavarianCream06 <- data.frame(
  Recipes = c("Coffee Bavarian Cream"),
  Year=c(2006),
  Amount = c(3.00, 1.50, 2.25, 1.00, 0.25, 1.00, 1.00),
  Measure = c("Tbs","cup", "tsp", "Tbs", "cup", "tsp", "cup"),
  Ingredient = c("milk,whole,3.25%","milk,whole,3.25%","gelatin,dry,unsweetened","instant coffee,regular","sugar,white,granulated,granulated","extract,vanilla","cream,whipping,heavy"))
CoffeeBavarianCream06[,-c(1,2)]
#Data table for recipie2 for year 1936 and 2006
CoffeeBavarianCream36
CoffeeBavarianCream06

#Writing in tab file
write.table(CoffeeBavarianCream36, file = "CoffeeBavarianCream36.tab") 
write.table(CoffeeBavarianCream06, file = "CoffeeBavarianCream06.tab") 

#Merge table for recipie2 for year 1936 and 2006 by ingredent 
CoffeeBavarianCream36_06 = merge(CoffeeBavarianCream36[,-c(1,2)],CoffeeBavarianCream06[,-c(1,2)], by = "Ingredient" )
CoffeeBavarianCream36_06 

#Writing in tab file
write.table(CoffeeBavarianCream36_06, file = "CoffeeBavarianCream36_06.tab") 

#Checking the number of row of the recipe2 for year 1936 and 2006 indivisually
nrow(CoffeeBavarianCream36)
nrow(CoffeeBavarianCream06)

#Checking the number of row of the recipe2 of both year 1936 & 2006 after merging by ingrident 
nrow(CoffeeBavarianCream36_06)



#Find the mismatching name
setdiff(CoffeeBavarianCream36$Ingredient, CoffeeBavarianCream06$Ingredient)



par(mfrow=c(1,2))
plot(as.factor(CoffeeBavarianCream36$Measure) ,CoffeeBavarianCream36$Amount,las = 3,xlab = "Measure(Unit)", ylab = "Amount",main = "1936")
plot(as.factor(CoffeeBavarianCream06$Measure) ,CoffeeBavarianCream06$Amount,las = 3,xlab = "Measure(Unit)", ylab = "Amount",main = "2006")

par(mfrow=c(1,2))
plot(as.factor(CoffeeBavarianCream36$Ingredient) ,CoffeeBavarianCream36$Amount,las = 3,xlab = "Ingredient", ylab = "Amount",main = "1936")
plot(as.factor(CoffeeBavarianCream06$Ingredient) ,CoffeeBavarianCream06$Amount,las = 3,xlab = "Ingredient", ylab = "Amount",main = "2006")

# The above plot shows that there is one unit mismatch from year 1936 to 2006 and the Similarity in the amount of "sugar,white,granulated,granulated" Ingredient.



```




```{r}
#lets now read all the.tab files


PineappleSnow36 = read.table("PineappleSnow36.tab",header = TRUE,stringsAsFactors = FALSE) #Read the above saved PineappleSnow36 .tab file
PineappleSnow06 = read.table("PineappleSnow06.tab",header = TRUE,stringsAsFactors = FALSE) #Read the above saved PineappleSnow06 .tab file
PineappleSnow36_06 = read.table("PineappleSnow36_06.tab",header = TRUE,stringsAsFactors = FALSE) #Read the above saved PineappleSnow36_06 .tab file
CoffeeBavarianCream36 = read.table("CoffeeBavarianCream36.tab",header = TRUE,stringsAsFactors = FALSE) #Read the above saved CoffeeBavarianCream36 .tab file
CoffeeBavarianCream06 = read.table("CoffeeBavarianCream06.tab",header = TRUE,stringsAsFactors = FALSE) #Read the above saved CoffeeBavarianCream06 .tab file
CoffeeBavarianCream36_06 = read.table("CoffeeBavarianCream36_06.tab",header = TRUE,stringsAsFactors = FALSE) #Read the above saved CoffeeBavarianCream36_06 .tab file



##Now lets read the given .tab file of BeefStroganoff data provided by instructor.
BeefStroganoff.1997 = read.table("BeefStroganoff.1997.tab",header = TRUE,stringsAsFactors = FALSE) #Read the provided BeefStroganoff.1997 .tab file
BeefStroganoff.1997

BeefStroganoff.2006 = read.table("BeefStroganoff.2006.tab",header = TRUE,stringsAsFactors = FALSE) #Read the provided BeefStroganoff.2006 .tab file
BeefStroganoff.2006


 ####### Now Merge data using merge() #############
  

BeefStroganoff97_06 = merge(BeefStroganoff.1997,BeefStroganoff.2006,by = "Ingredient") #Merge both BeefStroganoff recipe for both year 1997 and 2006
BeefStroganoff97_06

#Checking the number of row of the BeefStroganoff for year 1997 and 2006 indivisually
nrow(BeefStroganoff.1997)
nrow(BeefStroganoff.2006)

#Checking the number of row of the BeefStroganoff of both year 1997 & 2006 after merging by ingrident 
nrow(BeefStroganoff97_06)



```



```{r}
#Now lets Bind all the 1936 data together
WholeData.1936 = rbind(PineappleSnow36[,-c(1,2)],CoffeeBavarianCream36[,-c(1,2)],BeefStroganoff.1997) 
WholeData.1936

#Lets Bind all the 2006 data together
WholeData.2006 = rbind(PineappleSnow06[,-c(1,2)],CoffeeBavarianCream06[,-c(1,2)],BeefStroganoff.2006) 
WholeData.2006

#Merge all the data  together

WholeData.1936_2006 = merge(WholeData.1936,WholeData.2006,by = "Ingredient") 
WholeData.1936_2006


WholeData.1936_20061 = merge(WholeData.1936,WholeData.2006,by = "Ingredient",all = TRUE)
WholeData.1936_20061

#Ploting  the data
par(mfrow=c(1,2))
plot(as.factor(WholeData.1936$Measure),WholeData.1936$Amount,las = 3,xlab = "Measure.1936(Unit)", ylab = "Amount",main = "1936")
plot(as.factor(WholeData.2006$Measure),WholeData.2006$Amount,las = 3,xlab = "Measure.2006(Unit)", ylab = "Amount",main = "2006")
```

Plotting the WholeData seems that there is one unit mismatch from year 1936 to 2006. 

The following code is to read the Recipes.csv file and append rows for my recipes, then write the table back to Recipes.csv
```{r}
#Read the Recipes.csv file
Given_Recipe <- read.csv("Recipes.csv",header=TRUE,skip=0)
print(Given_Recipe) #Print the Recipes.csv file

#Creating the my Recipes data table file
My_Recipe <- data.frame(
  Recipe = c("Pineapple Snow","Pineapple Snow","Coffee Bavarian Cream","Coffee Bavarian Cream"),
  Year=c(1936, 2006,1936, 2006),
  MinServings = c(8.0,8.0, 8.0, 8.0),
  MaxServings = c(8.0,8.0, 8.0, 8.0))
print(My_Recipe)

#Append the Recipes file with my Recipies file
Recipes = rbind(Given_Recipe,My_Recipe)
print(Recipes) #Print the recipes file
write.table(Recipes, file = "Recipes1.tab") #Write in .tab file

```

The following code is to check my recipes ingredient with the standard USDA ingredient list, and see if it matches or not.

```{r}

#Reading the given .txt file of FOOD_DES
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")

#Now lets Check for the Ingredent match in FOOD_DES$Long_Decs file
grep('(G|g)elatin|(D|d)ry|(U|u)nsweetened',food_des.dat$Long_Desc)
grep('(W|w)ater|(G|g)eneric',food_des.dat$Long_Desc)
grep('(P|p)ineapple|(C|c)anned|(C|c)rushed|(I|i)n juice|(D|d)rained',food_des.dat$Long_Desc)
grep('(S|s)ugar|(W|w)hite|(G|g)ranulated|(G|g)ranulated',food_des.dat$Long_Desc)
grep('(C|c)ream|(W|w)hipping|(H|h)eavy',food_des.dat$Long_Desc)
grep('(E|e)xtract|(V|v)anilla',food_des.dat$Long_Desc)
grep('(M|m)ilk|(W|w)hole,3.25%',food_des.dat$Long_Desc)
grep('(C|c)offee|(B|b)rewed|(W|w)ith tap water',food_des.dat$Long_Desc)
grep('(E|e)gg white|(R|r)aw|(L|l)arge',food_des.dat$Long_Desc)
grep('(I|i)nstant coffee|(R|r)egular',food_des.dat$Long_Desc)

```


Thus it seems that my individual ingredients are listed in the USDA data which is validated by searching with grep function. But it seems like the USDA recipes list are different than ours,,i mean the way the ingredients are listed. 




