---
title: "Mid-term Project"
author: "Set 2"
date: "July 18, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prep

## Renaming Ingredients

Many of the ingredients don't have names that exist in `FOOD_DES.txt`, so I'm defining a function here that will convert any of those ingredient names into the proper one.

```{r}

# replaceIngredient - takes as input a dataset from Joy of Cooking,
# a name of one of the Joy of Cooking ingredients,
# and the name of the corresponding FOOD_DES ingredient.
# Replaces the Joy of Cooking ingredient with the FOOD_DES ingredient,
# and returns the result.
replaceIngredient <- function(joyOfCookingData, joyOfCookingIngredient, foodDesIngredient) {
  joyOfCookingData$Ingredient <- ifelse(
    tolower(joyOfCookingData$Ingredient) == tolower(joyOfCookingIngredient),
    foodDesIngredient,
    joyOfCookingData$Ingredient)
  return(joyOfCookingData)
}

# replaceIngredients - takes as input a dataset from Joy of Cooking
# and outputs the same dataset, but with
# ingredient names changed to correspond to what is in FOOD_DES.txt
replaceIngredients <- function(joyOfCookingData) {
  # Fix the names of ingredients so that we can cross-reference them.
  
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "celery, fresh, diced",
                                        "celery, raw")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "flour, all purpose, self rising, enriched",
                                        "wheat flour, white, all-purpose, self-rising, enriched")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "cream, whipping, heavy",
                                        "cream, fluid, heavy whipping")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "bread crumbs, plain, dry, grated",
                                        "bread, white, commercially prepared (includes soft bread crumbs)")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "beef flank steak, raw, choice, 0\" trim, urmis 2399",
                                        "beef, flank, steak, separable lean and fat, trimmed to 0\" fat, choice, raw")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "egg, raw",
                                        "egg, whole, raw, fresh")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "juice, tomato, canned, unsalted",
                                        "tomato juice, canned, without salt added")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "onion, yellow, fresh, chopped",
                                        "onions, sweet, raw")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "milk, whole, 3.25%",
                                        "milk, whole, 3.25% milkfat, without added vitamin a and vitamin d")
  joyOfCookingData <- replaceIngredient(joyOfCookingData,
                                        "potato, baked, large, 3\" to 4 1/4\"",
                                        "potatoes, white, flesh and skin, baked")
  
  return(joyOfCookingData)
}

```

## Merging in NDB_No

We'll need to add an `NDB_No` column to each recipe data frame. Here's the code to do it:

```{r}

# Read the FOOD_DES.txt table and set column names
food_des <- read.table("FOOD_DES.txt", sep="^", quote="~")
colnames(food_des) <- c("NDB_No", "FdGrp_Cd", "Long_Desc", "Shrt_Desc", "ComName",
                        "ManufacName", "Survey", "Ref_desc", "Refuse", "SciName",
                        "N_Factor", "Pro_Factor", "Fat_Factor", "CHO_Factor")
write.csv(food_des, file = "FOOD_DES.csv")
head(food_des)

# addNDB_No - a function that, given recipe data, will add an NDB_No column to it.
addNDB_No <- function(recipeData) {
# Make the ingredient names be all lower case, as I don't want to reject a match
  # just because of case
  recipeData$Ingredient <- tolower(recipeData$Ingredient)
  food_des$Long_Desc <- tolower(food_des$Long_Desc)
  
  recipeData <- merge(recipeData, food_des, by.x="Ingredient", by.y="Long_Desc",
                               all.x = TRUE, all.y = FALSE)

  # Discard all columns except the ones we're interested in
  recipeData <- recipeData[,c("Amount", "Unit", "Ingredient", "NDB_No")]
  
  return(recipeData)
}

```

# Create Recipe Tables

This part is where I create recipes as data frames in the required format and export these to a file.

## Flank Steak with Dressing - 1936

Here, we'll need to create the data frame and `.tab` file for the Flank Steak with Dressing recipe from 1936.

```{r}

# Create the data frame
Amount <- c(1, .25, 0.06161167, 1, 3, 1, 3, 0.12322334, 1)
Unit <- c("steak", "cup", "onion", "cup, crumbs", "Tbsp", "large", "Tbsp","cup", "cup")
Ingredient <- c("beef flank steak, raw, choice, 0\" trim, URMIS 2399",
                "butter, salted",
                "onion, yellow, fresh, chopped",
                "bread crumbs, plain, dry, grated",
                "celery, fresh, diced",
                "egg, raw",
                "oil, canola",
                "flour, all purpose, self rising, enriched",
                "juice, tomato, canned, unsalted")
flankSteakWithDressing1936 <- data.frame(Amount=Amount,
                                         Unit=Unit,
                                         Ingredient=Ingredient,
                                         stringsAsFactors = FALSE)

# Rename columns so that we can merge in the NDB_No column
flankSteakWithDressing1936 <- replaceIngredients(flankSteakWithDressing1936)
# Merge in the NDB_No column
flankSteakWithDressing1936 <- addNDB_No(flankSteakWithDressing1936)

# Print the result, to document our code
print(flankSteakWithDressing1936)

# Write the data to a file.
write.table(flankSteakWithDressing1936, file = "FlankSteakWithDressing.1936.tab", sep = " ")

```

Notes:

1. I converted `Tbs` to `Tbsp`
2. I changed `Tbsp` of onions to actual, based on this link: https://www.reference.com/food/much-minced-onion-equals-one-onion-9008e05cc4dd86cd. I assumed each onion produces 2 cups of minced onions.
3. I changed `lrg` to `large`
4. I changed `cup` to `cup, crumbs`
5. I changed `ea` to `steak`

## Flank Steak with Dressing - 2006

Here, we'll need to create the data frame and `.tab` file for the Flank Steak with Dressing recipe from 2006.

```{r}

# Create the data frame
Amount <- c(1, .25, 0.06161167, 1, 3, 1, 3, 0.12322334, 1)
Unit <- c("steak", "cup", "onion", "cup, crumbs", "Tbsp", "large", "Tbsp","cup", "cup")
Ingredient <- c("beef flank steak, raw, choice, 0\" trim, URMIS 2399",
                "butter, salted",
                "onion, yellow, fresh, chopped",
                "bread crumbs, plain, dry, grated",
                "celery, fresh, diced",
                "egg, raw",
                "oil, canola",
                "flour, all purpose, self rising, enriched",
                "juice, tomato, canned, unsalted")
flankSteakWithDressing2006 <- data.frame(Amount=Amount,
                                         Unit=Unit,
                                         Ingredient=Ingredient,
                                         stringsAsFactors = FALSE)

# Rename columns so that we can merge in the NDB_No column
flankSteakWithDressing2006 <- replaceIngredients(flankSteakWithDressing2006)
# Merge in the NDB_No column
flankSteakWithDressing2006 <- addNDB_No(flankSteakWithDressing2006)


# Print the result, to document our code
print(flankSteakWithDressing2006)

# Write the data to a file.
write.table(flankSteakWithDressing2006, file = "FlankSteakWithDressing.2006.tab", sep = " ")

```

Notes:

1. I converted `Tbs` to `Tbsp`
2. I changed `Tbsp` of onions to actual, based on this link: https://www.reference.com/food/much-minced-onion-equals-one-onion-9008e05cc4dd86cd. I assumed each onion produces 2 cups of minced onions.
3. I changed `lrg` to `large`
4. I changed `cup` to `cup, crumbs`
5. I changed `ea` to `steak`

## Check Flank Steak with Dressing for Errors

There are a couple checks we can do to make sure the Flank Steak with Dressing recipe doesn't have errors:

```{r}

# Make sure there aren't any NAs in the data frames.

if (any(is.na(flankSteakWithDressing2006))) {
  print("Warning: there are NAs in the 2006 data frame.")
}
if (any(is.na(flankSteakWithDressing1936))) {
  print("Warning: there are NAs in the 1936 data frame.")
}

# Merge the two data frames to get some extra checks
merged <- merge(flankSteakWithDressing1936, flankSteakWithDressing2006, by="Ingredient", suffixes=c("1936", "2006"))
print(merged)

# Plot the amounts, to see if they are linear
plot(merged$Amount1936, merged$Amount2006)

```

The scales look the same, so that looks good.

## Twice-Baked Potatoes - 1936

Here, we'll need to create the data frame and `.tab` file for the Twice-Baked Potatoes recipe from 1936.

```{r}

# Create the data frame
Amount = c(6,6,3,3)
Unit <- c("potato large (3\" to 4-1/4\" dia)", "Tbsp", "oz", "Tbsp")
Ingredient <- c("potato, baked, large, 3\" to 4 1/4\"",
                "butter, salted", "cheese, blue", "cream, whipping, heavy")
twiceBakedPotatoes1936 <- data.frame(Amount=Amount,
                                     Unit=Unit,
                                     Ingredient=Ingredient,
                                     stringsAsFactors = FALSE)

# Rename columns so that we can merge in the NDB_No column
twiceBakedPotatoes1936 <- replaceIngredients(twiceBakedPotatoes1936)
# Merge in the NDB_No column
twiceBakedPotatoes1936 <- addNDB_No(twiceBakedPotatoes1936)


# Print out the result, to document our code
print(twiceBakedPotatoes1936)

# Write the data to a file.
write.table(twiceBakedPotatoes1936, file = "TwiceBakedPotatoes.1936.tab", sep = " ")

```

Notes:

1. I converted `Tbs` to `Tbsp`
2. I converted `ea` to `potato large (3" to 4-1/4" dia)`

## Twice-Baked Potatoes - 2006

Here, we'll need to create the data frame and `.tab` file for the Twice-Baked Potatoes recipe from 2006.

```{r}

# Create the data frame
Amount = c(6,3,3,0.5)
Unit <- c("potato large (3\" to 4-1/4\" dia)", "Tbsp", "Tbsp", "cup")
Ingredient <- c("potato, baked, large, 3\" to 4 1/4\"", "butter, salted",
                "milk, whole, 3.25%", "cheese, parmesan, grated")
twiceBakedPotatoes2006 <- data.frame(Amount=Amount,
                                     Unit=Unit,
                                     Ingredient=Ingredient,
                                     stringsAsFactors = FALSE)

# Rename columns so that we can merge in the NDB_No column
twiceBakedPotatoes2006 <- replaceIngredients(twiceBakedPotatoes2006)
# Merge in the NDB_No column
twiceBakedPotatoes2006 <- addNDB_No(twiceBakedPotatoes2006)


# Print out the result, to document our code
print(twiceBakedPotatoes2006)

# Write the data to a file.
write.table(twiceBakedPotatoes2006, file = "TwiceBakedPotatoes.2006.tab", sep = " ")

```

Notes:

1. I converted `Tbs` to `Tbsp`
2. I converted `ea` to `potato large (3" to 4-1/4" dia)`

## Check Twice-Baked Potatoes for Errors

There are a couple checks we can do to make sure the Twice-Baked Potatoes recipe doesn't have errors:

```{r}

# Make sure there aren't any NAs in the data frames.

if (any(is.na(twiceBakedPotatoes2006))) {
  print("Warning: there are NAs in the 2006 data frame.")
}
if (any(is.na(twiceBakedPotatoes1936))) {
  print("Warning: there are NAs in the 1936 data frame.")
}

# Merge the two data frames to get some extra checks
merged <- merge(twiceBakedPotatoes1936, twiceBakedPotatoes2006, by="Ingredient", suffixes=c("1936", "2006"))
print(merged)

# Plot the amounts, to see if they are linear
plot(merged$Amount1936, merged$Amount2006)

```

This plot looks weird but this is actually correct:

- The ingredients changed (the type of cheese changed, and cream was replaced with milk). This means that of the four ingredients, only two end up in the merged data, which means only two points are plotted.
- Of the ingredients that didn't change in type of ingredient, only butter changed in amount (6 Tblsp to 3 Tblsp).

# rbind Recipes

Here's where I append the recipe data frames together to produce a single big data frame of all the recipes. Note that to do this properly, I will have to add columns for recipe and year.

```{r}

# loadRecipeFile - takes a file name as input and returns the recipe data frame
loadRecipeFile <- function(file) {
  # Split by the period character to extract recipe name and year  
  fileNameParts <- strsplit(file, ".", fixed = TRUE)[[1]]
  recipe <- fileNameParts[1]
  year <- as.numeric(fileNameParts[2])

  # Read the tab file
  result <- read.table(file, header = TRUE)
  
  # The beef stroganoff recipe doesn't include an NDB_No column,
  # so we'll hack one in with NAs. That way, rbind won't complain.
  if (is.null(result$NDB_No)) {
    result$NDB_No <- NA
  }
  
  # Change the Measure column to Unit column.
  # Workaround for the format in beef stroganoff.
  if (!is.null(result$Measure)) {
    result$Unit <- result$Measure
    result$Measure <- NULL
  }
  
  # Set the columns on the resulting data frame
  result$Recipe <- recipe
  result$Year <- year
  
  return(result)
}

# loadRecipeFiles - takes a character vector of file names and returns a
# single data frame of all the files' contents
loadRecipeFiles <- function(files) {
  # Use lapply to convert the list of files into a list of data frames
  # loaded from the files
  listOfRecipeDataFrames <- lapply(files, loadRecipeFile)
  
  # Merge the list of data frames
  mergedData <- do.call(rbind, listOfRecipeDataFrames)
  
  return(mergedData)
}

# Get a list of .tab files
files <- list.files(pattern=".+\\.tab$")
print(files)

# Load the files
mergedData <- loadRecipeFiles(files)
print(mergedData)

```

# Cross-Reference Ingredients and Units

We'll need to merge (a.k.a. join) the recipe data set with the tables in `FOOD_DES.txt` and `WEIGHT.txt`. First, let's merge with `FOOD_DES.txt`:

```{r}

# Read the FOOD_DES.txt table and set column names
food_des <- read.table("FOOD_DES.txt", sep="^", quote="~")
colnames(food_des) <- c("NDB_No", "FdGrp_Cd", "Long_Desc", "Shrt_Desc", "ComName",
                        "ManufacName", "Survey", "Ref_desc", "Refuse", "SciName",
                        "N_Factor", "Pro_Factor", "Fat_Factor", "CHO_Factor")
write.csv(food_des, file = "FOOD_DES.csv")
head(food_des)

# Load just the recipes we're generating, because we don't want to load the example recipe,
# because its ingredient names have not been changed to match FOOD_DES.txt.

mergedData <- loadRecipeFiles(c(
  "FlankSteakWithDressing.1936.tab", "FlankSteakWithDressing.2006.tab",
  "TwiceBakedPotatoes.1936.tab", "TwiceBakedPotatoes.2006.tab"))

# Make the ingredient names be all lower case, as I don't want to reject a match
# just because of case
mergedData$Ingredient <- tolower(mergedData$Ingredient)
food_des$Long_Desc <- tolower(food_des$Long_Desc)

crossReferencedData <- merge(mergedData, food_des, by="NDB_No",
                             all.x = TRUE, all.y = FALSE)
print(crossReferencedData)

```

Time to check the merge. If there are any `NAs` in the columns that were merged in from food_des, then that means that not all the ingredients had names that matched `FOOD_DES.txt`.

```{r}

if (any(is.na(crossReferencedData$Shrt_Desc))) {
  print("Bad news: there are ingredients in the recipe that are missing in FOOD_DES.txt")
} else {
  print("Good news: the ingredients in the recipe are all found in FOOD_DES.txt")
}

```

Time to merge in `WEIGHT.txt`:

```{r}

# Read the WEIGHT.txt table and set columns
weight <- read.table("WEIGHT.txt", sep="^", quote="~")
colnames(weight) <- c("NDB_No", "Seq", "Amount", "Unit", "Gm_Wgt", "Num_Data_Pts", "Std_Dev")
head(weight)

weight$Unit <- tolower(weight$Unit)
crossReferencedData$Unit <- tolower(crossReferencedData$Unit)

crossReferencedData <- merge(crossReferencedData, weight, by=c("NDB_No", "Unit"),
                             all.x = TRUE, all.y = FALSE)
print(crossReferencedData)

```

Let's check the `WEIGHT.txt` merge too. If there are any `NAs` in `Gm_Wgt`, then we know that the merge wasn't wholly successful:

```{r}

if (any(is.na(crossReferencedData$Gm_Wgt))) {
  print("Bad news: there are units in the recipe that are missing in WEIGHT.txt")
} else {
  print("Good news: the units in the recipe are all found in WEIGHT.txt")
}

```

# Add Recipes to Recipes.csv

Here, I'll add my recipes to `Recipes.csv`:

```{r}

# Load the recipes
recipes.csv <- read.csv("Recipes.csv")
# Create a data frame of my recipes
myRecipes <- data.frame(Recipe=c("Twice-Baked Potatoes", "Twice-Baked Potatoes",
                                 "Flank Steak with Dressing", "Flank Steak with Dressing"),
                        Year=c(1936, 2006, 1936, 2006),
                        MinServings=c(6, 6, 4, 4),
                        MaxServings=c(6, 6, 6, 6))
print(myRecipes)

# Append my recipes onto the end of recipes.csv
recipes.csv <- rbind(recipes.csv, myRecipes)

# Write out the file. Note row.names = FALSE. This is so that we don't get
# an extra column of numbers that indicate the row numbers. We don't need that
# since it wasn't in the original Recipes.csv file.
write.csv(recipes.csv, file = "Recipes.csv", row.names = FALSE)

```

