---
title: "Joy Of Cooking Midterm Project Prep"
author: "Set 38"
date: "7/21/2018"
output:
  html_document: default
  pdf_document: default
---

# Part 1 - Assign Recipes

First, generate a list of all files in the Recipes folder.
```{r}


######################################################
############# Fruit and Berry Cobbler ################
#############     1936      ##########################
######################################################

# A data frame is assigned to FruitandBerryCobbler_1936
FruitandBerryCobbler_1936 <- data.frame(
# Amount, Measure and Ingredients are defined here as per my Recipe in the PDF
  Amount = c(3.00,0.67,1.00,1.00,0.50,2.00,6.00,4.00), 
  Measure = c("cup","cup", "Tbs", "cup", "Tbs", "Tbs","Tbs","Tbs"),
  Ingredient = c("peaches,fresh,sliced", "sugar,white,granulated,granulated","flour,all purpose,self rising,enriched", "flour,cake,white,unsifted,enriched", "sugar,white,granulated,granulated", "butter,salted", "cream, whipping, heavy", "butter, salted"))
```
###Comment: 
As per the requirement I have assigned a data frame to my First Recipe of Year 1936 FruitandBerryCobbler_1936 and then I have declared the Amount, Measure, Ingredient here. It creates table for my recipe, with columns amount, Unit and Ingredient.  

```{r}
######################################################
############# Fruit and Berry Cobbler ################
#############     2006      ##########################
######################################################

# A data frame is assigned to FruitandBerryCobbler_2006
FruitandBerryCobbler_2006 <- data.frame(
# Amount, Measure and Ingredients are defined as per my recipe in the PDF  
  Amount = c(1.75, 0.50, 2.00, 1.75, 4.00, 0.75, 1.00, 2.00),
  Measure = c("lb","cup","Tbs","cup", "Tbs", "cup", "Tbs", "tsp"),
  Ingredient = c("peaches,fresh,sliced", "sugar,white,granulated","starch,corn","flour, allpurpose, self rising, enriched", "butter,salted", "milk,whole,3.25%", "butter, salted", "sugar, white, granulated, granulated"))
```
###Comment: 
As per the requirement I have assigned a data frame to my First Recipe of Year 2006 FruitandBerryCobbler_2006 and then I have declared the Amount, Measure, Ingredient here. It creates table for my recipe, with columns amount, Unit and Ingredient. 

```{r}
######################################################
############# Tapioca Custard ########################
#############     1936      ##########################
######################################################

#  A data frame is assigned to Tapioca Custard_2006
TapiocaCustard_1936 <- data.frame(
# Amount, Measure and Ingredients are defined as per my recipe in the PDF 
  Amount = c(0.33,0.50,2.00,4.00,1.00),
  Measure = c("cup","cup","Irg","cup","tsp"),
  Ingredient = c("tapioca,quick cooking,dry", "sugar,white,granulated,granulated","egg,raw","milk,whole,3.25%","extract,vanilla"))
```
###Comment: 
As per the requirement I have assigned a data frame to my First Recipe of Year 1936 FruitandBerryCobbler_1936 and then I have declared the Amount, Measure, Ingredient here. It creates table for my recipe, with columns amount, Unit and Ingredient. 

```{r}
######################################################
############# Tapioca Custard ########################
#############     2006      ##########################
######################################################
#  A data frame is assigned to Tapioca Custard_2006
TapiocaCustard_2006 <- data.frame(
  # Amount, Measure and Ingredients are defined as per my recipe in the PDF 
  Amount = c(2.50,0.33,3.00,1.00),
  Measure = c("cup","cup","Tbs","tsp"),
  Ingredient = c("milk,whole,3.25%", "sugar,white,granulated,granulated", "tapioca, quickcooking, dry", "extract, vanilla"))
```
###Comment: 
As per the requirement I have assigned a data frame to my First Recipe of Year 1936 FruitandBerryCobbler_1936 and then I have declared the Amount, Measure, Ingredient here. It creates table for my recipe, with columns amount, Unit and Ingredient. 

```{r}
#############################################################################
####### 2. Writing each individual recipe to a tab delimited file ###########
#############################################################################

# write.table using .tab for recipe FruitandBerryCobbler_1936
write.table(FruitandBerryCobbler_1936,"FruitandBerryCobbler_1936.tab",row.names = FALSE)
# write.table using .tab for recipe FruitandBerryCobbler_2006
write.table(FruitandBerryCobbler_2006,"FruitandBerryCobbler_2006.tab",row.names = FALSE)
# write.table using .tab for recipe TapiocaCustard_1936
write.table(TapiocaCustard_1936,"TapiocaCustard_1936.tab",row.names = FALSE)
# write.table using .tab for recipe TapiocaCustard_2006
write.table(TapiocaCustard_2006, "TapiocaCustard_2006.tab",row.names = FALSE)
```
###Comment: 
According to the requirement I have written write.table for FruitandBerryCobbler_1936, FruitandBerryCobbler_2006, TapiocaCustard_1936 and TapiocaCustard_2006. Finally Made sure that the name of the tab file is comprimised of the recipe file name and the year of the recipe. 

```{r}
#############################################################################
####### 3. Read  and merge the tables  ######################################
#############################################################################

# read.table for the year 1997 of recipe BeefStroganoff
BeefStroganoff.1997 = read.table("BeefStroganoff.1997.tab",header = TRUE,stringsAsFactors = FALSE) 
# read.table for the year 2006 of recipe BeefStroganoff
BeefStroganoff.2006 = read.table("BeefStroganoff.2006.tab",header = TRUE,stringsAsFactors = FALSE)
# combining both the recipes FruitandBerryCobbler and TapiocaCustard of Year 1936) 
Rec1936 = rbind(FruitandBerryCobbler_1936,TapiocaCustard_1936)
Rec2006 = rbind(TapiocaCustard_2006,TapiocaCustard_2006)
# Merging all four of them (1936 and 1997) to BeefStroganoff_1936_1997 and (2006) to BeefStroganoff_2006
MergwithBSO_1936_1997 = rbind(Rec1936,BeefStroganoff.1997)
MergwithBSO_2006 = rbind(Rec2006,BeefStroganoff.2006) 

```
###Comment:
According to the requirement I have read in each tab file and merged the tables with the example BeefStroganoff tables. Here in the above code read.table is used for BeefStroganoff.1997, BeefStroganoff.2006 and rbind has been done for both the recipes of years 1936 and 2006 and assigned to Rec1936 and Rec2006 respectively. 
Merge is used for all four BeefStrongoff.1997, BeefStrongoff.2006, Rec1936, Rec2006.

```{r}
#############################################################################
####### 4. data to check for errors  ########################################
#############################################################################

# Fruit and Berry Cobbler and TapiocaCustard merged by ingredient
FruitandBerryCobblerIng <- merge(FruitandBerryCobbler_1936,FruitandBerryCobbler_2006, by = "Ingredient")
TapiocaCustardIng <- merge(TapiocaCustard_1936,TapiocaCustard_2006, by = "Ingredient")
#Measure.x and Measure.y is plotted and checked where the measures are not plotted
plot(FruitandBerryCobblerIng$Measure.x,FruitandBerryCobblerIng$Measure.y, col = "blue")
plot(TapiocaCustardIng$Measure.x,TapiocaCustardIng$Measure.y, col = "orange")
#A contingency table comparing Measure.x and Measure.y to look for mismatches in measures between is created.
table(FruitandBerryCobblerIng$Measure.x,FruitandBerryCobblerIng$Measure.y)
table(TapiocaCustardIng$Measure.x,TapiocaCustardIng$Measure.y)
# Amount.x and Amount.y to look for unexpected changes in values
plot(FruitandBerryCobblerIng$Amount.x,FruitandBerryCobblerIng$Amount.y,col = "blue")
plot(TapiocaCustardIng$Amount.x,TapiocaCustardIng$Amount.y,col = "orange")

```
###Comment: 
Here in the above code Fruit and Berry Cobbler and Tapioca Custard are merged by ingredient and Measure.x and Measure.y is plotted and checked where the measures are not plotted.Followed by a contingency table comparing Measure.x and Measure.y to look for mismatches in measure between is created. Finally Amount.x and Amount.y to look for unxpected changes in values. 

You can see the plots above according to the requirement. 
```{r}
######################################################################################

food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")


######################################################################################

###Fruit and Berry Cobbler_1936 
# Original Ingredients peaches,fresh,sliced
# Matched "Peaches, frozen, sliced, sweetened "
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_1936_1 <- food_des.dat[grep('^peaches[[:punct:]].frozen[[:punct:]].sliced',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "sugar,white,granulated,granulated"
# Matched  "Sugars, granulated "
# It is Partially Matched.
FBC_1936_2 <- food_des.dat[grep('^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "flour,all purpose,self rising,enriched"
# Matched "Wheat flour, white, all-purpose, self-rising, enriched"
# Matched Appropriate or most suitable in the FOOD_DES
# It is Partially Matched.
FBC_1936_3 <- food_des.dat[grep('^wheat.flour[[:punct:]].white[[:punct:]].all-purpose[[:punct:]].self-rising',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "flour,cake,white,unsifted,enriched"
# Matched "Wheat flour, white, cake, enriched"
# Partially Matched.

FBC_1936_4 <- food_des.dat[grep('^wheat.flour[[:punct:]].white[[:punct:]].cake', food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "sugar,white,granulated,granulated"
# Matched "sugar,white,granulated,granulated"
# Partially Matched.

FBC_1936_5 <- food_des.dat[grep('^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "butter,salted"
# Matched ingredient "Butter, salted"
# Matched Exactly in the FOOD_DES
FBC_1936_6 <- food_des.dat[grep('^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "cream, whipping, heavy"
# Matched ingredient are "Cream, fluid, heavy whipping"
# Matched Exactly in the FOOD_DES
FBC_1936_7 <- food_des.dat[grep('^cream[[:punct:]].fluid[[:punct:]].heavy.whipping',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original ingredients are "butter,salted"
# Matched ingredient "Butter, salted"
# Matched Exactly in the FOOD_DES

FBC_1936_8 <- food_des.dat[grep('^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Combining all of them with rbind
FABC_1936 <- rbind(FBC_1936_1,FBC_1936_2,FBC_1936_3,FBC_1936_4,FBC_1936_5,FBC_1936_6,FBC_1936_7,FBC_1936_8)

###Fruit and Berry Cobbler_2006

# Original Ingredients "peaches,fresh,sliced"
# Matched "Peaches, frozen, sliced, sweetened"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_2006_1 <- food_des.dat[grep('^peaches[[:punct:]].frozen[[:punct:]].sliced',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "sugar,white,granulated"
# Matched "Sugars, granulated"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_2006_2 <- food_des.dat[grep('^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "starch,corn"
# Matched "Cornstarch"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_2006_3 <- food_des.dat[grep('^Cornstarch',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "flour, allpurpose, self rising, enriched"
# Matched "Wheat flour, white, all-purpose, self-rising, enriched"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_2006_4 <- food_des.dat[grep('^wheat.flour[[:punct:]].white[[:punct:]].all-purpose[[:punct:]].self-rising',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "butter,salted"
# Matched "Butter, salted"
# Matched Exactly in the FOOD_DES
FBC_2006_5 <- food_des.dat[grep('^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients ""milk,whole,3.25%
# Matched "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_2006_6 <- food_des.dat[grep('^milk[[:punct:]].whole[[:punct:]].3.25% milkfat[[:punct:]].without added vitamin A and vitamin D',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "butter,salted"
# Matched "Butter, salted"
# Matched Exactly in the FOOD_DES
FBC_2006_7 <- food_des.dat[grep('^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "sugar,white,granulated,granulated"
# Matched "Sugars, granulated"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
FBC_2006_8 <- food_des.dat[grep('^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Combining all of them with rbind
FABC_2006 <- rbind(FBC_2006_1,FBC_2006_2,FBC_2006_3,FBC_2006_4,FBC_2006_5,FBC_1936_6,FBC_1936_7,FBC_1936_8)

###Tapioca Custard_1936

# Original Ingredients "tapioca,quick cooking,dry"
# Matched "Puddings, tapioca, dry mix"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_1936_1 <- food_des.dat[grep('^Puddings[[:punct:]].tapioca[[:punct:]].dry mix$',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "sugar,white,granulated,granulated"
# Matched "Sugars, granulated"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_1936_2 <- food_des.dat[grep('^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "egg,raw"
# Matched "Egg, whole, raw, fresh"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_1936_3 <- food_des.dat[grep('^egg[[:punct:]].whole[[:punct:]].raw[[:punct:]].fresh',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "milk,whole,3.25%"
# Matched "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_1936_4 <- food_des.dat[grep('^milk[[:punct:]].whole[[:punct:]].3.25% milkfat[[:punct:]].without added vitamin A and vitamin D',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "extract,vanilla"
# Matched "Vanilla extract"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_1936_5 <- food_des.dat[grep('Vanilla extract$',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Combining all of them with rbind
TPC_1936 <- rbind(TC_1936_1,TC_1936_2,TC_1936_3,TC_1936_4,TC_1936_5)

###Tapioca Custard_2006

# Original Ingredients "milk,whole,3.25%"
# Matched "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_2006_1 <- food_des.dat[grep('^milk[[:punct:]].whole[[:punct:]].3.25% milkfat[[:punct:]].without added vitamin A and vitamin D',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "sugar,white,granulated,granulated"
# Matched "Sugars, granulated"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_2006_2 <-food_des.dat[grep('^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "tapioca, quickcooking, dry"
# Matched "Puddings, tapioca, dry mix"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_2006_3 <-food_des.dat[grep('^Puddings[[:punct:]].tapioca[[:punct:]].dry mix$',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "extract, vanilla"
# Matched "Vanilla extract"
# Matched Appropriate or most suitable in the FOOD_DES
# Partially Matched.
TC_2006_4 <-food_des.dat[grep('Vanilla extract$',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Combining all of them with rbind
TPC_2006 <- rbind(TC_2006_1,TC_2006_2,TC_2006_3,TC_2006_4)

# BeefStroganoff year 1997

# Original Ingredients "beef roast, tenderloin, raw, 0" trim" 
# Matched "Beef, loin, tenderloin roast, boneless, separable lean only, trimmed to 0" fat, select, raw"
# It is Partially Matched 
BS_97_1 <- food_des.dat[grep('^beef[[:punct:]].loin[[:punct:]].tenderloin.roast[[:punct:]].boneless[[:punct:]].separable.lean.only[[:punct:]].trimmed.to.0".fat[[:punct:]].select',food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]

# Original Ingredients "butter,salad" 
# Matched "Butter, salted"
# It is Matched Exactly in the FOOD_DES 
BS_97_2 <- food_des.dat[grep('^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "flour,all purpose,self rising,enriched" 
# Matched "Wheat flour, white, all-purpose, self-rising, enriched"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_97_3 <- food_des.dat[grep('^wheat.flour[[:punct:]].white[[:punct:]].all-purpose[[:punct:]].self-rising',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "stock, beef, prepared from recipe" 
# Matched "Soup, stock, beef, home-prepared"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_97_4 <- food_des.dat[grep("^soup[[:punct:]].stock[[:punct:]].beef", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]


# Original Ingredients "sour cream, cultured" 
# Matched "Cream, sour, cultured"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_97_5 <- food_des.dat[grep("^cream[[:punct:]].sour[[:punct:]].cultured",food_des.dat$Long_Desc, ignore.case = TRUE,  fixed = FALSE),c(1,3)]

# Original Ingredients "mustard, dijon" 
# Matched "Mustard, prepared, yellow"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_97_6 <- food_des.dat[grep("^mustard[[:punct:]]|.(D|d)ijon",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]

BSO_1997 <- rbind(BS_97_1,BS_97_2,BS_97_3,BS_97_4,BS_97_5,BS_97_6)


# BeefStroganoff year 2006

# Original Ingredients "beef roast, tenderloin, raw, 0" trim" 
# Matched "Beef, loin, tenderloin roast, boneless, separable lean only, trimmed to 0" fat, select, raw"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_06_1 <- food_des.dat[grep('^beef[[:punct:]].loin[[:punct:]].tenderloin.roast[[:punct:]].boneless[[:punct:]].separable.lean.only[[:punct:]].trimmed.to.0".fat[[:punct:]].select',food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]

# Original Ingredients "oil, canola" 
# Matched "Oil, canola"
# It is Matched Exactly in the FOOD_DES  
BS_06_2 <-food_des.dat[grep("^oil[[:punct:]].canola", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]


# Original Ingredients "butter,salad" 
# Matched "Butter, salted"
# It is Matched Exactly in the FOOD_DES  
BS_06_3 <-food_des.dat[grep('^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]

# Original Ingredients "onion, yellow, fresh, medium, 2 1/2\" 
# Matched "Onions, yellow, sauteed"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_06_4 <-food_des.dat[grep("^onions[[:punct:]].yellow", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]


# Original Ingredients "mushrooms, fresh" 
# Matched "Mushrooms, white, raw"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_06_5 <-food_des.dat[grep("^mushrooms[[:punct:]].white[[:punct:]].raw", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]

# Original Ingredients "stock, beef, prepared from recipe" 
# Matched "Soup, stock, beef, home-prepared"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_06_6 <-food_des.dat[grep("^soup[[:punct:]].stock[[:punct:]].beef", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]

# Original Ingredients "sour cream, cultured" 
# Matched "Cream, sour, cultured"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_06_7 <-food_des.dat[grep("^cream[[:punct:]].sour[[:punct:]].cultured",food_des.dat$Long_Desc, ignore.case = TRUE,  fixed = FALSE),c(1,3)]

# Original Ingredients "mustard, dijon" 
# Matched "Mustard, prepared, yellow"
# It is Matched Appropriate or most suitable in the FOOD_DES  
# It is Partially Matched 
BS_06_8 <-food_des.dat[grep("^mustard[[:punct:]]|.(D|d)ijon",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]

BSO_2006 <- rbind(BS_06_1,BS_06_2,BS_06_3,BS_06_4,BS_06_5,BS_06_6,BS_06_7,BS_06_8)



# An empty column c() is assigned to FruitandBerryCobbler_1936
FruitandBerryCobbler_1936$Ingredient <- c()
# FABC_1936 which is rbinded with all of FruitandBerryCobbler of Year 1936 is assigned
FruitandBerryCobbler_1936$Ingredient <- FABC_1936$Long_Desc
# A new variable NDB_NO is assigned here
FruitandBerryCobbler_1936$NDB_NO <- FABC_1936$NDB_NO

# An empty column c() is assigned to FruitandBerryCobbler_1936
FruitandBerryCobbler_2006$Ingredient <- c()
# FABC_2006 which is rbinded with all of FruitandBerryCobbler of Year 2006 is assigned
FruitandBerryCobbler_2006$Ingredient <- FABC_2006$Long_Desc
# A new variable NDB_NO is assigned here
FruitandBerryCobbler_2006$NDB_NO <- FABC_2006$NDB_NO

#Print FruitandBerryCobbler_1936
FruitandBerryCobbler_1936
#Print FruitandBerryCobbler_2006
FruitandBerryCobbler_2006

# An empty column c() is assigned to TapiocaCustard_1936
TapiocaCustard_1936$Ingredient <- c()
# TPC_1936 which is rbinded with all of FruitandBerryCobbler of Year 1936 is assigned
TapiocaCustard_1936$Ingredient <- TPC_1936$Long_Desc
# A new variable NDB_NO is assigned here
TapiocaCustard_1936$NDB_NO <- TPC_1936$NDB_NO

# An empty column c() is assigned to TapiocaCustard_2006
TapiocaCustard_2006$Ingredient <- c()
# TPC_2006 which is rbinded with all of FruitandBerryCobbler of Year 2006 is assigned
TapiocaCustard_2006$Ingredient <- TPC_2006$Long_Desc
# A new variable NDB_NO is assigned here
TapiocaCustard_2006$NDB_NO <- TPC_2006$NDB_NO

# Print TapiocaCustard_1936
TapiocaCustard_1936
# Print TapiocaCustard_2006
TapiocaCustard_2006

# An empty column c() is assigned to BeefStroganoff.1997
BeefStroganoff.1997$Ingredient <- c()
# BSO_1997 which is rbinded with all of FruitandBerryCobbler of Year 1997 is assigned
BeefStroganoff.1997$Ingredient <- BSO_1997$Long_Desc
# A new variable NDB_NO is assigned here
BeefStroganoff.1997$NDB_NO <- BSO_1997$NDB_NO

# An empty column c() is assigned to BeefStroganoff.2006
BeefStroganoff.2006$Ingredient <- c()
# BSO_1997 which is rbinded with all of FruitandBerryCobbler of Year 1997 is assigned
BeefStroganoff.2006$Ingredient <- BSO_2006$Long_Desc
# A new variable NDB_NO is assigned here
BeefStroganoff.2006$NDB_NO <- BSO_2006$NDB_NO

# Print BeefStroganoff.1997
BeefStroganoff.1997
# Print BeefStroganoff.2006
BeefStroganoff.2006



All <- rbind(FruitandBerryCobbler_1936,FruitandBerryCobbler_2006,TapiocaCustard_1936,TapiocaCustard_2006,BeefStroganoff.1997,BeefStroganoff.2006)
All

# creates tab file
write.table(FruitandBerryCobbler_1936,"FruitandBerryCobbler_1936.tab", sep="\t", row.names = FALSE)
# creates tab file
write.table(FruitandBerryCobbler_2006,"FruitandBerryCobbler_2006.tab", sep="\t", row.names = FALSE)
# creates tab file
write.table(TapiocaCustard_1936,"TapiocaCustard_1936.tab", sep="\t", row.names = FALSE)
# creates tab file
write.table(TapiocaCustard_2006,"TapiocaCustard_2006.tab", sep="\t", row.names = FALSE)

```



```{r}
########################################################################################################
####### 5.Read the Recipes.csv, append rows for your recipes and write the table back to Recipes.csv.###   ########################################################################################################
########################################################################################################


Recipe <- data.frame(
  Recipe=c("Fruit and Berry Cobbler","Fruit and Berry Cobbler","Tapioca Custard","Tapioca Custard"),
  # Years are written here and Minimum and Maximum Servings are matched from the PDF
  Year=c(1936,2006,1936,2006),
  MinServings=c(6,8,8,6),
  MaxServings=c(6,8,8,6))
# Recipe is printed here
print(Recipe)

# Recipes.csv file is read here 
BeefStroganoffRecipe <- read.csv("Recipes.csv",header=TRUE)
# BeefStroganoff_Recipe is printed here
print(BeefStroganoffRecipe)
# Combining BeefStroganoff_Recipe and Recipe
Combination <- rbind(BeefStroganoffRecipe,Recipe)
# creates csv file
write.csv(Combination, file="Recipes.csv",row.names = FALSE)
# reads csv file
Recipe.Table <- read.csv("Recipes.csv",header=TRUE)
# Recipe.Table is printed here
print(Recipe.Table)


```
###Comment: 
Here we have created a data frame and did the Minimum and Maximum Servings of Fruit and Berry Cobbler and Tapioca Custard. Recipe is printed. Recipes.csv file is read and assigned to BeefStronganoff_Recipe and printed. Combine Recipe and BeefStroganoff_Recipe using rbind and then create csv file, read it and print the Recipe.Table. 

# Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.
```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)
```

```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```

```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)
```
