---
title: "Midterm Report"
author: "Set 33"
date: "July 20, 2018"
output:
  word_document: default
  html_document: default
---
# Introduction

We have been asked to work on two recipes that was randomly given to us from USDA food data base.  The food descriptions, nutrition and weight form USDA database was used for this objectives. I was assigned to work on German potato salad and Irish stew

# Objectives

###To collect information about the recipes, ingredients and amount and measure of servings for the 1936 and 2006.
###To learn how to merge or match the tables in our recipes to other persons or to recipes that was given as an example
###To match the ingredients in the data base and check if any measurement or ingredient inconsistences was obsrved. 





# PART 1. Preparing tables for the selcted recipes

Gathering infromation about the given recipes and table preparation

## Irish Stew 1936
```{r}
Irish.Stew.1936=data.frame(Recipe=c(rep("Irish Stew", 10)),
                           NDB_No=c(rep(NA, 10)),
                           Year=c("1936", "1936", "1936", "1936", "1936", "1936", "1936", "1936", "1936","1936"),
                           Amount=c(1.50, 3.00, 0.25, 3.00, 0.25, 1.50, 0.25, 0.33, 2.00, 0.50),
                           Measure=c("lb", "Tbs", "cup","cup","cup","cup","cup","cup","Tbs","cup"),
                           Ingredient=c('beef, chuck, stew meat, raw, choice,URMIS 1100',
                                        'oil, canola',
                                        'flour, all purpose, self rising, enrcihed',
                                        'water, generic',
                                        'onion, yellow, fresh, sliced',
                                        'potato, fresh, with skin, diced',
                                        'carrotes, fresh, sliced',
                                        'turnip, fresh, cubes',
                                        'flour, all purpose, self rising, enrcihed',
                                        'stock, beef, prepared from recipe'))

  

```


## Irish Stew 2006
```{r}
Irish.Stew.2006=data.frame(Recipe=c(rep("Irish Stew", 6)),
                           NDB_No=c(rep(NA, 6)),
                           Year=c("2006","2006","2006","2006","2006","2006"), 
                           Amount=c(2.00,2.00,3.00,6.00,3.00,4.00),
                           Measure=c("Tbs", "ea","lb","ea","cup","ea"),
                           Ingredient=c('Oil, canola', 
                                        'onion, yelow, fresh, medium, 2 1/2', 
                                        'lamb stew meat,raw lean 1/4"trim', 
                                        'potato, red, fresh, with skin, medium, 2 1/4" to 3 1/4"',
                                        'stock, chiken, prepared from recipe',
                                        'carrots, fresh, medium'))

```

## German Potato Salad 1936
```{r}

German.Potato.Salad.1936=data.frame(Recipe=c(rep("German Potato Salad", 6)),
                                    NDB_No=c(rep(NA, 6)),
                                    Year=c("1936","1936","1936","1936","1936","1936"), 
                                    Amount=c(6.00,4.00,0.25,5.00,5.00,0.25),
                                    Measure=c("Tbs", "ea","lb","ea","cup","ea"),
                                    Ingredient=c('potato, red, fresh, with skin, medium, 2 1/4" to 3 1/4"', 
                                      'bacon, raw, cured', 
                                      'onion, yellow, fresh, chopped', 
                                      'vinegar, distilled',
                                      'sugar, white, granulated, granulated',
                                      'water, generic'))

```


## German Potato Salad 2006
```{r}

German.Potato.Salad.2006=data.frame(Recipe=c(rep("German Potato Salad", 8)),
                                    NDB_No=c(rep(NA, 8)),
                                    Year=c("2006","2006","2006","2006","2006","2006","2006","2006"), 
                                    Amount=c(2.00,4.00,0.25,0.25,0.25,5.00,5.00,0.25),
                                    Measure=c("lb", "pce","cup","cup","cup","cup", "tsp", "cup"),
                                    Ingredient=c('potato, red, fresh, with skin, medium, 2 1/4" to 3 1/4"', 
                                                 'bacon, raw, cured',
                                                 'onion, yellow, fresh, chopped',
                                                 'celery, fresh, diced',
                                                 'pickles, dill, chopped',
                                                 'vinegar, cider',
                                                 'sugar, white, granulated, granulated',
                                                 'water, generic'))

```



```{r}
print(Irish.Stew.1936)
write.table(Irish.Stew.1936,"Irish.Stew.1936.txt")

print(Irish.Stew.2006)
write.table(Irish.Stew.2006,"Irish.Stew.2006.txt")

print(German.Potato.Salad.1936)
write.table(German.Potato.Salad.1936,"German.Potato.Salad.1936.txt")

print(German.Potato.Salad.2006)
write.table(German.Potato.Salad.2006,"German.Potato.Salad.2006.txt")

```

# Part 2 - Checing and matching of ingredients from database USDA.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.
```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```


### Irish strew  ingredient matching to food database in USDA
```{r}
rish.Stew.1936.beef<- food_des.dat[grep("beef.*chunk*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.oil<- food_des.dat[grep("oil, canola", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.flour<- food_des.dat[grep("Flour*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.water<- food_des.dat[grep("generic*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.onion<- food_des.dat[grep ("onion.*yellow*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.potato<- food_des.dat[grep('*potato*' , food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.carrotes<- food_des.dat[grep('carrotes.*' , food_des.dat$Long_Desc, ignore.case = T), ]
Irish.Stew.1936.turnip<- food_des.dat[grep('turnip*' , food_des.dat$Long_Desc, ignore.case = T), ]
Irish.Stew.1936.flour<- food_des.dat[grep("Flour*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.stock<- food_des.dat[grep('(?=.stock.)(?=.*beef*)', food_des.dat$Long_Desc,perl = TRUE),]

Irish.Stew.1936.oil<- food_des.dat[grep("oil, canola", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.onion<- food_des.dat[grep ("onion.*yellow*", food_des.dat$Long_Desc, ignore.case = T),]
rish.Stew.1936.beef<- food_des.dat[grep("lamb.*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.potato<- food_des.dat[grep('*potato*' , food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.flour<- food_des.dat[grep("stock.*chiken*", food_des.dat$Long_Desc, ignore.case = T),]
Irish.Stew.1936.carrotes<- food_des.dat[grep('carrotes.*' , food_des.dat$Long_Desc, ignore.case = T), ]

```

### German Potato salad ingredient matching to food database in USDA
```{r}
potato.1936.carrotes<- food_des.dat[grep('carrotes.*' , food_des.dat$Long_Desc, ignore.case = T), ]
potato.1936.potato.<- food_des.dat[grep("potato.*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.bacon<- food_des.dat[grep("bacon*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.onion.<- food_des.dat[grep ("onion.*yellow*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.water<- food_des.dat[grep( "water*" , food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.sugar<- food_des.dat[grep("sugar*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.generic<- food_des.dat[grep("generic*", food_des.dat$Long_Desc, ignore.case = T),]

potato.1936.potato<- food_des.dat[grep("potato.*red*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.bacon<- food_des.dat[grep("bacon.raw*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.onion<- food_des.dat[grep ("onion.*yellow*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.celery<- food_des.dat[grep('*celery.*' , food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.pickles<- food_des.dat[grep('*pickles.*dill*' , food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.vinegar<- food_des.dat[grep('vinegar.*cider' , food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.sugar<- food_des.dat[grep("sugar*", food_des.dat$Long_Desc, ignore.case = T),]
potato.1936.water<- food_des.dat[grep( "water*" , food_des.dat$Long_Desc, ignore.case = T),]

```


# Part 3. Finding NDB_No and relating tables using NDB_No as a key

During matching in part 2, row numbers of each ingredient in the food.dat database was identified. Then after, using this row numbers I sublet the data of the data in the food.dat database to get NDB_No for each ingredients.

Finally, using NDB_No i able to related and retrieve information from weight and nutrition database in the USDA too.

### Selcting ingreidents from the food data base of USDA using row number identified during matching
```{r}
subset_foodIrish.des1936<- food_des.dat[c(1404,	706,	6000,	4038,	2650,	2709,	2517,	2854,	6000,	1266),c(1,3)]
subset_foodIrish.des2006<- food_des.dat[c(706,	2650,	4622,	2709,	1267,	2717),c(1,3)]
subsetpotato_food.des1936<- food_des.dat[c(2709,	2218,	2650,	315,	5776,	4038),c(1,3)]
subsetpotato_food.des2006<- food_des.dat[c(2709,	2218,	2650,	2530,	3159,	310,	5776,	4038),c(1,3)]

```


### Insering DOB_no into the recipies selcted
```{r}
##Replacing of ingreidents from the data base to the sected ingredients
Irish.Stew.1936$Ingredient=subset_foodIrish.des1936$Long_Desc
Irish.Stew.2006$Ingredient=subset_foodIrish.des2006$Long_Desc
German.Potato.Salad.1936$Ingredient =subsetpotato_food.des1936$Long_Desc
German.Potato.Salad.2006$Ingredient =subsetpotato_food.des2006$Long_Desc

##Inserting NDB_No
Irish.Stew.1936$NDB_No=subset_foodIrish.des1936$NDB_No
Irish.Stew.2006$NDB_No=subset_foodIrish.des2006$NDB_No
German.Potato.Salad.1936$NDB_No =subsetpotato_food.des1936$NDB_No
German.Potato.Salad.2006$NDB_No =subsetpotato_food.des2006$NDB_No
```

# Printing the final tables
```{r}

print(Irish.Stew.1936)
print(Irish.Stew.2006)
print(German.Potato.Salad.1936)
print(German.Potato.Salad.2006)

```

## Write each individual recipe to a tab delimited file. The name of the tab file should be comprised of the recipe file name and the year of the recipe (1936 or 2006, in most cases).
```{r}

write.table(Irish.Stew.1936,file= "Irish.Stew.1936") #Write in .tab file
write.table(Irish.Stew.2006,file="Irish.Stew.2006" )
write.table(German.Potato.Salad.1936, file= "German.Potato.Salad.1936")
write.table(German.Potato.Salad.2006, file= "German.Potato.Salad.1936")

```

## Read in each tab file, and merge the tables with the example Beef Stroganoff tables.

The Reading the example beef Stroganoff tables.
```{r}

BeefStroganoff.2006 <- read.table("BeefStroganoff.2006.tab", header=TRUE)
BeefStroganoff.2006$NDB_No<-c(NA)
BeefStroganoff.2006$Recipe<-c("BeefStroganoff")
BeefStroganoff.2006$Year<-c("2006")
                                                            

BeefStroganoff.1997 <- read.table("BeefStroganoff.1997.tab", header=TRUE)
BeefStroganoff.1997$NDB_No<-c(NA)
BeefStroganoff.1997$Recipe<-c("BeefStroganoff")
BeefStroganoff.1997$Year<-c("1997")




```

### Merging all the selcted recipes to  beeffStroganoff recipes

Merging the recipes  German.Potato.Salad, BeefStroganoff, and Irish.Stew using *reduce function*
```{r}
#merging all the recipes with beeffStroganoff.1997
merged.data=Reduce(function(x, y) merge(x, y, all=TRUE), list(Irish.Stew.1936,
                                                  Irish.Stew.2006,
                                                  German.Potato.Salad.1936,
                                                  German.Potato.Salad.2006,                  
                                                  BeefStroganoff.1997,
                                                  BeefStroganoff.2006))
#Saving the result
write.csv(merged.data,"merged.data.csv")
```

### Comparing with the gram weight database using NDB_No as key


```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat,10)
summary(weight.dat)

```


```{r}
Wt.Irish.1936 = subset(weight.dat, NDB_No %in% c(Irish.Stew.1936$NDB_No))
Wt.Irish.2006 = subset(weight.dat, NDB_No %in% c(Irish.Stew.2006$NDB_No))
Wt.Potato.1936 = subset(weight.dat, NDB_No %in% c(German.Potato.Salad.1936$NDB_No))
Wt.Potato.2006 = subset(weight.dat, NDB_No %in% c(German.Potato.Salad.2006$NDB_No))

#print(Wt.Irish.1936)
#print(Wt.Irish.2006)
#print(Wt.Potato.1936)
#print(Wt.Potato.1936)

```


### Comparing with Nutrion data base

```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat[, c(1,2,5,11,12)])

```

```{r}
nut_data.dat=nut_data.dat[,c(1,2,5,11,12)]

Nt.Irish.1936 = subset(nut_data.dat, NDB_No %in% c(Irish.Stew.1936$NDB_No))
Nt.Irish.2006 = subset(nut_data.dat, NDB_No %in% c(Irish.Stew.2006$NDB_No))
Nt.Potato.1936 = subset(nut_data.dat, NDB_No %in% c(German.Potato.Salad.1936$NDB_No))
Nt.Potato.2006 = subset(nut_data.dat, NDB_No %in% c(German.Potato.Salad.2006$NDB_No))

#print(Nt.Irish.1936)
#print(Nt.Irish.2006)
#print(Nt.Potato.1936)
#print(Nt.Potato.2006)


```

# Part 4. Proof of data to check for errors

Recipies are merged by  Ingredient and to check for inconsistency in the measurement and amounts.

```{r}
#Merging by Ingredient

merge.P<-merge(German.Potato.Salad.1936, German.Potato.Salad.2006, by="Ingredient")
print(merge.P)

#The plot to show the existance of inconsistency

plot(merge.P$Amount.x , merge.P$Amount.y )
print(merge.P)

```

### Comments:
The merging of the recipes using ingadient indicated that there are difrences in some of the gradents interms of ther mesuremen acroos difrent years aswell as with in the some year too. 


```{r}
#Merging by Ingredient

merge.I<-merge(Irish.Stew.1936, Irish.Stew.2006, by="Ingredient")
print(merge.I)

#The plot to show the existance of inconsistency

plot(merge.I$Amount.x , merge.I$Amount.y )
print(merge.I)

```

#Part 5.Reading the Recipes.csv file and appending my recipes.
 
#Reading the example Recipe file

```{r}
#Example recipes
Recipes<- read.csv("Recipes.csv", header = T)

##Recipe file for my recipes
my_recipes<-data.frame(
  Recipe= c('Irish Stew ', 'Irish Stew', 'German Potato Salad', 'German Potato Salad'),
  Year= c(1936,2006,1936,2006),
  MinServings=c(4,4,4,6),
  MaxServings=c(6,6,4,6))

## merging both files
Rec.max.min.merge<-rbind(Recipes, my_recipes)

##writing the merged file to CSV
write.csv(Rec.max.min.merge, "Rec.max.min.merge.csv")

# Table showing the recipes merged to show the  maximum and  minim serving 
print(Rec.max.min.merge)
```







```

