---
title: 'Joy Of Cooking Midterm Project '
author: "Set 39"
date: '21 July,2018'
output:
  html_document: default
  pdf_document: default
---




We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/


```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)
```

```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```

```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)
```


```{r}


############################# STEP 1 ################################


### Creating a data frame for the recipe Apple Dumplings 1936   
### with columns as Amount, Measure, Ingredient.The data frame  
### is named AppleDumpings1936. I have chosen to include all     
### the Ingredients from the pdf

AppleDumplings1936 <- data.frame(
  Amount = c(1.00,2.50,2.00),
  Measure = c("ea","Tbs","tsp"),
  Ingredient = c("apple,fresh,medium,3'' ", "sugar,brown,packed","butter,salted"))

print(AppleDumplings1936)


### Creating a data frame for the recipe Apple Dumplings 2006   
### with columns as Amount, Measure, Ingredient.The data frame  
### is named AppleDumpings2006.  I have chosen to include all    
### the Ingredients from the pdf                                 

AppleDumplings2006 <- data.frame(
  
  Amount = c(6.00,0.50,0.25,1.00),
  Measure = c("ea","cup","cup","lrg"),
  Ingredient = c("apple,fresh,small,2 3/4'' ", "sugar,brown,dark","butter,salted","egg,raw"))

print(AppleDumplings2006)


### Creating a data frame for the recipe Scalloped Oysters 1936   
### with columns as Amount, Measure, Ingredient.The data frame    
### is named ScallopedOysters1936. I have chosen to include all   
### the Ingredients from the pdf                                   

ScallopedOysters1936 <- data.frame(
  
  Amount = c(1.00,6.00,0.50,28.00,0.50),
  Measure = c("pnt","Tbs","cup","ea","cup"),
  Ingredient = c("oysters,eastern,in liquid,12 oz can", "cream,whipping,heavy","bread crumbs,plain,dry,grated","crackers,saltine","butter,salted"))


print(ScallopedOysters1936)

### Creating a data frame for the recipe Scalloped Oysters 1936   
### with columns as Amount, Measure, Ingredient.The data frame    
### is named ScallopedOysters1936. I have chosen to include all   
### the Ingredients from the pdf                                

ScallopedOysters2006 <- data.frame(
  
  Amount = c(1.00,2.00,1.00,0.75,1.00),
  Measure = c("qt","cup","cup","cup","cup"),
  Ingredient = c("oysters,eastern,in liquid,12 oz can","crackers,saltine,crushed", "bread crumbs,plain,dry,grated","butter,salted","cream,whipping,heavy"))


print(ScallopedOysters2006)


```



```{r}


############################# STEP 2 ################################




### Each recipe is written into a tab delimited file using        
### write.table function with '\t' as seperator.The naming        
### convention used for the tab delimited file is name of the                             
### Recipe(Ex:AppleDumplings) followed by Yearname                                  



write.table(AppleDumplings1936,"AppleDumplings1936.tab",sep="\t", row.names = FALSE)

write.table(AppleDumplings2006,"AppleDumplings2006.tab",sep="\t" ,row.names = FALSE)

write.table(ScallopedOysters1936,"ScallopedOysters1936.tab", sep="\t",row.names = FALSE)

write.table(ScallopedOysters2006,"ScallopedOysters2006.tab",sep="\t",row.names = FALSE)



```


```{r}

############################# STEP 3 ################################


### Each tab file generated in the previous is read by using     
### read.table function Apple Dumplings 1936 and 2006 tab files  
### are assigned to t1 , t2 respectively. Scalloped Oysters 1936                         
### and 2006 tab files are assigned to t3,t4 respectively            

t1 <- read.table("AppleDumplings1936.tab", header = TRUE )
t2 <- read.table("AppleDumplings2006.tab", header = TRUE )
t3 <- read.table("ScallopedOysters1936.tab", header = TRUE)
t4 <- read.table("ScallopedOysters2006.tab", header = TRUE)




### The given BeefStroganoff.1997.tab file is read using the  
### function read.table. The name 'BeefStraganoff1997' is       
### assigned to the table. The same is repeated with the        
### BeefStroganoff.2006.tab file                                

BeefStroganoff1997 <- read.table("BeefStroganoff.1997.tab", header = TRUE)
BeefStroganoff2006 <- read.table("BeefStroganoff.2006.tab", header = TRUE)




### Using rbind function to merge tables t1,t2,t3,t4 with    
### the BeefStroganoff tables. The merged data is assigned the  
### name Combined_data. The successfull merge demonstrates that 
### data can be merged into shared format                       

Combined_data <- rbind(t1,t2,t3,t4,BeefStroganoff1997,BeefStroganoff2006)
print(Combined_data)



```


```{r}

############################# STEP 4 ################################

### Proofing thhe data to check for errors                       



### AppleDumplings1936 and AppleDumplins2006 are merged by using
### merge function. This generates common records from both the 
### tables (Natural join)

AppleDumplings_merge <- merge(AppleDumplings1936,AppleDumplings2006,by ="Ingredient")
print(AppleDumplings_merge)



### ScallopedOysterd1936 and ScallopedOysters2006 are merged by  
### using merge function. This generates common records from     
### the tables (Natural join)

ScallopedOysters_merge <- merge(ScallopedOysters1936, ScallopedOysters2006, by ="Ingredient")
print(ScallopedOysters_merge)



### Measure1936 vs Measure2006 of Apple Dumplings is plotted to  
### observe inconsistencies in the units of measurement of       
### ingredients.                                                 
plot(AppleDumplings_merge$Measure.x , AppleDumplings_merge$Measure.y)


### Measure1936 vs Measure2006 of Scalloped Oysters is plotted   
### to observe inconsistencies in the units of measurement of    
### ingredients.                                                 
plot(ScallopedOysters_merge$Measure.x , ScallopedOysters_merge$Measure.y)


### Amount1936 vs Amount2006 of Apple Dumplings is plotted to    
### observe anomalies in values of amount which can also be due  
### wrong entry of data
plot(AppleDumplings_merge$Amount.x , AppleDumplings_merge$Amount.y)



### Amount1936 vs Amount2006 of Scalloped Oysters is plotted to 
### observe anomalies in values of amount which can also be due  
### wrong entry of data

plot(ScallopedOysters_merge$Amount.x , ScallopedOysters_merge$Amount.y)



### Contigency table for the measures of Apple Dumplings.The table 
### helps to determine mismatch in measures                        
table(AppleDumplings_merge$Measure.x , AppleDumplings_merge$Measure.y)



### Contigency table for the measures of Scalloped Oysters.The     
### tables helps to determine mismatch in measures                 

table(ScallopedOysters_merge$Measure.x , ScallopedOysters_merge$Measure.y)



```

###Comments for plots

Apple Dumplings Measures vs Meassures 
Inconsistencies with the units/measures are observed in Apple Dumplings 1936 and 2006. Different measures are used for the same Ingredients in 1936 and 2006. The contigency table also deemonstrates this inconsistency

Apple Dumplings Amount vs AMount   
There is no observed anomaly in the amount values of Apple Dumplings. However, since the measures for the same ingredients changed in 1936 and 2006 there is a visual difference

Scalloped Oysters Measures vs Measures 
Inconsistencies with the units/measure are observed in Scalloped Oysters 1936 and 2006. Different measures are used for the same Ingredients in 1936 and 2006.The contigency
table also deemonstrates this inconsistency


Scalloped Oysters Amount vs AMount   
There is no observed anomaly in the amount values of Apple Dumplings. However, since the measures for the same ingredients changed in 1936 and 2006 there is a visual difference


```{r}

############################# STEP 5 ################################



### The Recipes.csv is read and assigned the name Recipes        

Recipes <- read.csv("Recipes.csv", header = TRUE, stringsAsFactors = FALSE)


### Rows are appended in the Recipes.csv for each Recipe         
### This is done by using indexes of Recipes                     
Recipes[nrow(Recipes)+1,] <- c("Apple Dumplings", 1936,1,1) 
Recipes[nrow(Recipes)+1,] <- c("Apple Dumplings",2006,6,6)
Recipes[nrow(Recipes)+1,] <- c("Scalloped Oysters",1936,6,6)
Recipes[nrow(Recipes)+1,] <- c("Scalloped Oysters",2006,6,6)

print(Recipes)

### The updated Recipes table is written back into Recipes.csv  
write.table(Recipes,"Recipes.csv",sep = ",",row.names = FALSE)


```

```{r}



############################# Performing matches in FOOD_DES ################################


### Below are a series of partial or full matchings for the ingredients  
### of recipes assined to me. They are matched using grep function and   
### as grep returns index numbers, the corresponding matched NDB_No      
### along with Long_Desc for the ingredients are extracted from Food_des 



### Apple Dumplings 1936 Ingredients 
	  # apple,fresh,medium,3" #		
	  # sugar,brown,packed		#
	  # tsp	butter,salted     #

# Matching ' apple,fresh,medium,3" ' in Food_Des           
# First a search for apples is done                        
# Then of all the entries I found 'Apples, raw, with skin' 
# as the apt match
ad36_1<- food_des.dat[grep("^apples[[:punct:]].",food_des.dat$Long_Desc,ignore.case = TRUE, fixed = FALSE),c(1,3)]
ad36_1_s <- ad36_1[1,]
print(ad36_1_s)

# Matching 'sugar,brown,packed' in Food_Des            
# First a search for sugar is done                  
# Then of all the entries I found 'sugar turbinado' 
# as the apt match. The particular record is extracted by index                             
ad36_2 <- food_des.dat[grep("^sugar[[:punct:]]",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
ad36_2_s <- ad36_2[2,]
print(ad36_2_s)



# Matching 'Butter, salted' in Food_Des
# A full match is obtained for 'Butter, salted' in Food_Des
ad36_3 <- food_des.dat[grep("^butter[[:punct:]].salted",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(ad36_3)


# using rbind the NDB_No's and  Ingredients columns of
# AppleDumplings1936 are combined into a new table 
new_AD36 <- rbind(ad36_1_s,ad36_2_s,ad36_3)



### Apple Dumplings 2006 Ingredients 

              # apple,fresh,small,2 3/4''  #
              # sugar,brown,dark           #
              # butter,salted              #
              # egg,raw                    #

# Matching ' apple,fresh,small,2 3/4" ' in Food_Des        
# First a search for apples is done                        
# Then of all the entries I found 'Apples, raw, with skin' 
# as the apt match. The particular record is extracted by index 

ad06_1<- food_des.dat[grep("^apples[[:punct:]].",food_des.dat$Long_Desc,ignore.case = TRUE, fixed = FALSE),c(1,3)]
ad06_1_s <- ad06_1[1,]
print(ad06_1_s)



# Matching 'sugar,brown,dark' in Food_Des               
# First a search for sugar is done                  
# Then of all the entries I found 'sugar turbinado' 
# as the apt match. The particular record is extracted by index 

ad06_2 <- food_des.dat[grep("^sugar[[:punct:]]",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
ad06_2_s <- ad06_2[2,]
print(ad06_2_s)


# Matching 'Butter, salted' in Food_Des  
# A full match is obtained for 'Butter, salted' in Food_Des
ad06_3 <- food_des.dat[grep("^butter[[:punct:]].salted",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(ad06_3)


# Matching 'Egg,raw' in Food_Des
# First a search is done for 'egg'
# Out of all the entries I found Egg, whole, raw, fresh 
# as the apt match. The particular record is extracted by index
ad06_04 <- food_des.dat[grep("^egg[[:punct:]].",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE ),c(1,3)]
ad06_04_s <-  ad06_04[1,]
print(ad06_04_s)

# using rbind the NDB_No's and  Ingredients columns of
# AppleDumplings2006 are combined into a new table 
new_AD06 <- rbind(ad06_1_s,ad06_2_s,ad06_3,ad06_04_s)




# Scalloped Oysters 1936 Ingredients#

    # oysters,eastern,in liquid,12 oz can  #
    #            cream,whipping,heavy      #
    #     bread crumbs,plain,dry,grated    #
    #             crackers,saltine         #   
    #                  butter,salted       #

# Matching 'oysters,eastern,in liquid,12 oz can' in Food_Des
# First a search is done for 'oyster, eastern'
# Out of all entries I found 'Mollusks,oyster,eastern,canned' 
# as the apt match. The particular record is extracted by index
so36_1 <- food_des.dat[grep("oyster[[:punct:]].eastern",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
so36_1_s <- so36_1[4,]
print(so36_1_s)



# Matching 'cream,whipping,heavy' in Food_Des
# First a search is done for 'whipping'
# Out of all the entries I found 'Cream, fluid, heavy whipping'
# as an apt match. The particular record is extracted by index
so36_2 <- food_des.dat[grep("whipping",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
so36_2_s <- so36_2[2,]
print(so36_2_s)


# Matching ' bread crumbs,plain,dry,grated' in Food_Des
# First a search is done for 'bread, crumbs'
# Out of all the entries I found ' Bread, crumbs, dry, grated, plain'
# as an apt match. The particular record is extracted by index
so36_3 <- food_des.dat[grep("bread[[:punct:]].crumbs",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
so36_3_s <- so36_3[1,]
print(so36_3_s)



# Matching 'crackers,  saltine in Food_Des
# First a search is done for 'crackers, saltine'
# Out of all the entries I found 'Crackers, saltines (includes oyster, soda, soup)'
# as an apt match. The particular record is extracted by index.
so36_4 <- food_des.dat[grep("^crackers[[:punct:]].saltine",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
so36_4_s <- so36_4[1,]
print(so36_4_s)


# Matching 'Butter, salted' in Food_Des
# A full match is obtained for 'Butter, salted' in Food_Des
so36_5 <- food_des.dat[grep("^butter[[:punct:]].salted",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(so36_5)


# using rbind the NDB_No's and  Ingredients columns of
# ScallopedOysters1936 are combined into a new table 

new_so36 <- rbind(so36_1_s,so36_2_s,so36_3_s,so36_4_s,so36_5)





#Scalloped Oysters 2006 Ingredients#
         # oysters,eastern,in liquid,12 oz can  #
         #  crackers,saltine,crushed            #
         #   bread crumbs,plain,dry,grated      #
         #        butter,salted                 #
         #        cream,whipping,heavy          #



# Matching 'oysters,eastern,in liquid,12 oz can' in Food_Des
# First a search is done for 'oyster, eastern'
# Out of all entries I found 'Mollusks,oyster,eastern,canned' 
# as the apt match. The particular record is extracted by index
so06_1 <- food_des.dat[grep("oyster[[:punct:]].eastern",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
so06_1_s <- so06_1[4,]
print(so06_1_s)


# Matching 'crackers,  saltine in Food_Des
# First a search is done for 'crackers, saltine'
# Out of all the entries I found 'Crackers, saltines (includes oyster, soda, soup)'
# as an apt match. The particular record is extracted by index.
so06_2 <- food_des.dat[grep("^crackers[[:punct:]].saltine",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
so06_2_s <- so06_2[1,]
print(so06_2_s)



# Matching ' bread crumbs,plain,dry,grated' in Food_Des
# First a search is done for 'bread, crumbs'
# Out of all the entries I found ' Bread, crumbs, dry, grated, plain'
# as an apt match. The particular record is extracted by index
so06_3 <- food_des.dat[grep("bread[[:punct:]].crumbs",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
so06_3_s <- so06_3[1,]
print(so06_3_s)



# Matching 'Butter, salted' in Food_Des
# A full match is obtained for 'Butter, salted' in Food_Des
so06_4 <- food_des.dat[grep("^butter[[:punct:]].salted",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(so06_4)


# Matching 'cream,whipping,heavy' in Food_Des
# First a search is done for 'whipping'
# Out of all the entries I found 'Cream, fluid, heavy whipping'
# as an apt match. The particular record is extracted by index
so06_5 <- food_des.dat[grep("whipping",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
so06_5_s <- so06_5[2,]
print(so06_5_s)

# using rbind the NDB_No's and  Ingredients columns of
# ScallopedOysters2006 are combined into a new table 
new_so06 <- rbind(so06_1_s,so06_2_s,so06_3_s,so06_4,so06_5_s)




# Beef Stroganoff 1997 Ingredients#


         # beef roast, tenderloin, raw, 0" trim       #
         #              butter, salted                #
         #  flour, all purpose, self rising, enriched #
         #     stock, beef, prepared from recipe      #
         #                  sour cream, cultured      #
         #                        mustard, dijon      #

# Matching 'beef roast, tenderloin, o" trim' in Food_Des
# First a search is done for 'tenderloin'
# Out of all the entries I found ' Beef, loin, tenderloin roast, 
# boneless, separable lean only, trimmed to 0" fat, select, raw'
# as an apt match. The particular record is extracted by index

bs97_1_1 <-food_des.dat[grep("tenderloin",food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]
bs97_1 <- bs97_1_1[grep("beef",bs97_1_1$Long_Desc,ignore.case = TRUE,fixed = FALSE), ]
bs97_1_s <- bs97_1[36,] 
print(bs97_1_s)


 
# Matching 'Butter, salted' in Food_Des
# A full match is obtained for 'Butter, salted' in Food_Des
bs97_2 <- food_des.dat[grep("^butter[[:punct:]].salted",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(bs97_2)



# Matching 'flour, all purpose, self rising, enriched' in Food_Des
# First a search is done for 'all-purpose'
# Out of all the entries I found ' Wheat flour, white, all-purpose, 
#self-rising, enriched, as an apt match. 
# The particular record is extracted by index
bs97_3 <- food_des.dat[grep("all.purpose",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
bs97_3_s <- bs97_3[4,]
print(bs97_3_s)


# Matching ' stock, beef, prepared from recipe' in Food_Des
# First a search is done for 'stock'
# Out of all the entries I found 'Soup, stock, beef, home-prepared ', 
# as an apt match. The particular record is extracted by index
bs97_4 <- food_des.dat[grep("stock",food_des.dat$Long_Desc,ignore.case = TRUE, fixed = FALSE),c(1,3)]
bs97_4_s <- bs97_4[2,]
print(bs97_4_s)


# Matching  'sour cream, cultured' in Food_Des
# First a search is done for 'cultured'
# Out of all the entries I found 'Cream, sour, cultured', 
# as an apt match. The particular record is extracted by index
bs97_5 <- food_des.dat[grep("cultured",food_des.dat$Long_Desc, ignore.case = TRUE,  fixed = FALSE),c(1,3)]
bs97_5_s <- bs97_5[2,]
print(bs97_5_s)


# Matching 'Mustard, Dijon' in Food_Des
# A partial match is obtained for 'Mustard, prepared, yellow' in Food_Des
bs97_6 <- food_des.dat[grep("^mustard[[:punct:]]|.dijon",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
print(bs97_6)

# using rbind the NDB_No's and  Ingredients columnns of
# BeefStroganoff2006 are commbined into a new table 
new_bs97 <- rbind(bs97_1_s,bs97_2,bs97_3_s,bs97_4_s,bs97_5_s,bs97_6)



# BeefStroganoff year 2006

           # beef roast, tenderloin, raw, 0" trim  #
           #            oil, canola                #
           #        butter, salted                 #
           # onion, yellow, fresh, medium, 2 1/2"  #
           #         mushrooms, fresh              #
           #  stock, beef, prepared from recipe    #    
           #  sour cream, cultured                 #
           #          mustard, dijon               #


# Matching 'beef roast, tenderloin, o" trim' in Food_Des
# First a search is done for 'tenderloin'
# Out of all the entries I found ' Beef, loin, tenderloin roast, 
# boneless, separable lean only, trimmed to 0" fat, select, raw'
# as an apt match. The particular record is extracted by index
bs06_1_1 <-food_des.dat[grep("tenderloin",food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)]
bs06_1 <- bs06_1_1[grep("beef",bs06_1_1$Long_Desc,ignore.case = TRUE,fixed = FALSE), ]
bs06_1_s <- bs06_1[36,] 
print(bs06_1_s)


# Matching 'oil, canola' in Food_Des
# A full match is obtained for 'Oil, Canola' in Food_Des
bs06_2 <- food_des.dat[grep("^oil[[:punct:]].canola", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(bs06_2)

# Matching 'Butter, salted' in Food_Des
# A full match is obtained for 'Butter, salted' in Food_Des
bs06_3 <- food_des.dat[grep("^butter[[:punct:]].salted",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(bs06_3)



# Matching 'onion, yellow, fresh, medium, 2 1/2' in Food_Des
# A partial match is obtained as 'Onions, yellow, sauteed' in Food_Des
bs06_4 <- food_des.dat[grep("onions[[:punct:]].yellow", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
print(bs06_4)


# Matching 'mushrooms, fresh' in Food_Des
# First a search is done for 'mushrooms'
# Out of all the entries I found 'Mushrooms, white, raw ', 
# as an apt match. The particular record is extracted by index
bs06_5 <- food_des.dat[grep("mushrooms",food_des.dat$Long_Desc, ignore.case = TRUE,fixed = FALSE),c(1,3)]
bs06_5_s <- bs06_5[5,]
print(bs06_5_s)



# Matching ' stock, beef, prepared from recipe' in Food_Des
# First a search is done for 'stock'
# Out of all the entries I found 'Soup, stock, beef, home-prepared ', 
# as an apt match. The particular record is extracted by index
bs06_6 <- food_des.dat[grep("stock",food_des.dat$Long_Desc,ignore.case = TRUE, fixed = FALSE),c(1,3)]
bs06_6_s <- bs06_6[2,]
print(bs06_6_s)


# Matching  'sour cream, cultured' in Food_Des
# First a search is done for 'cultured'
# Out of all the entries I found 'Cream, sour, cultured', 
# as an apt match. The particular record is extracted by index
bs06_7 <- food_des.dat[grep("cultured",food_des.dat$Long_Desc, ignore.case = TRUE,  fixed = FALSE),c(1,3)]
bs06_7_s <- bs06_7[2,]
print(bs06_7_s)


# Matching 'Mustard, Dijon' in Food_Des
# A partial match is obtained for 'Mustard, prepared, yellow' in Food_Des
bs06_8 <- food_des.dat[grep("^mustard[[:punct:]]|.dijon",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
print(bs06_8) 


# using rbind the NDB_No's and  Ingredients columns of 
# BeefStroganoff2006 are commbined into a new table 
new_bs06 <- rbind(bs06_1_s,bs06_2,bs06_3,bs06_4,bs06_5_s,bs06_6_s,bs06_7_s,bs06_8)

```

```{r}



############################## Appending NDB_No column #########################


### The Ingredient and NDB_no column are appended 
### to the tables created in STEP1
### The appending is done in 3 steps 
### First, the Ingredient column in the data table is made empty
### Second, the Long_Desc from the new table created in the previous
### section is assigned as Ingredient to the recipe table
### Third, the NDB_No from the new table is assigned to the NDB_No
### Column created in the recipe table

### The above process is done for AppleDumplings1936, AppleDumplings2006,
### ScallopedOysters1936, ScallopedOysters2006, BeefStroganoff1997,BeefStroganff2006



AppleDumplings1936$Ingredient <- c()
AppleDumplings1936$Ingredient <- new_AD36$Long_Desc
AppleDumplings1936$NDB_No <- new_AD36$NDB_No
print(AppleDumplings1936)


AppleDumplings2006$Ingredient <- c()
AppleDumplings2006$Ingredient <- new_AD06$Long_Desc
AppleDumplings2006$NDB_No <- new_AD06$NDB_No
print(AppleDumplings2006)


ScallopedOysters1936$Ingredient <- c()
ScallopedOysters1936$Ingredient <- new_so36$Long_Desc
ScallopedOysters1936$NDB_No <- new_so36$NDB_No
print(ScallopedOysters1936)


ScallopedOysters2006$Ingredient <- c()
ScallopedOysters2006$Ingredient <- new_so06$Long_Desc
ScallopedOysters2006$NDB_No <- new_so06$NDB_No
print(ScallopedOysters2006)


BeefStroganoff1997$Ingredient <- c()
BeefStroganoff1997$Ingredient <- new_bs97$Long_Desc
BeefStroganoff1997$NDB_No <- new_bs97$NDB_No
print(BeefStroganoff1997)

BeefStroganoff2006$Ingredient <- c()
BeefStroganoff2006$Ingredient <- new_bs06$Long_Desc
BeefStroganoff2006$NDB_No <- new_bs06$NDB_No
print(BeefStroganoff2006)



### All the recipe tables updated above are combined into combined_data1

Combined_data1 <- rbind(AppleDumplings1936,AppleDumplings2006,ScallopedOysters1936,ScallopedOysters2006, BeefStroganoff1997, BeefStroganoff2006)

print(Combined_data1)

### The updated recipes tables are written into tab delimited files

write.table(AppleDumplings1936,"AppleDumplings1936.tab",sep="\t", row.names = FALSE)

write.table(AppleDumplings2006,"AppleDumplings2006.tab",sep="\t" ,row.names = FALSE)

write.table(ScallopedOysters1936,"ScallopedOysters1936.tab", sep="\t",row.names = FALSE)

write.table(ScallopedOysters2006,"ScallopedOysters2006.tab",sep="\t",row.names = FALSE)




```