---
title: "MidTerm_Project"
author: "Set 37"
date: "July 21, 2018"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

# Part 1 - Assign Recipes

```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  1. Creating data tables for your assigned recipes  @@@@#
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  Recipe : Sweet and Sour Brisket @@@@#
#@@@@            Year : 1936           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


Sweet_and_Sour_Brisket_1936 <- data.frame(
  Amount=c(3,2,0.25,1.5,0.25,0.25,1,0.5,1.5,0.25),
  Measure=c("lb","Tbs","cup","cup","cup","cup","Tbs","tsp","tsp","cup"),
  Ingredient=c("beef brisket, flat half, raw, 0'' trim, URMIS 1623",
"Butter, salted","onion, yellow,fresh,chopped","tomatoes,whole,
canned,unsalted","bell pepper,green,fresh,chopped",
"celery,fresh, diced","flour,all purpose,self rising,enriched","sugar,white,
granulated,granulated","vinegar,distilled","cream,whipping,heavy"))
```

### Comment:
I have created a data frame of Sweet and Sour Brisket Recipe of year 1936 with column names as Amount, Measure, Ingredient.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  Recipe : Sweet and Sour Brisket @@@@#
#@@@@            Year : 2006           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


Sweet_and_Sour_Brisket_2006 <- data.frame(
  Amount=c(3.5,3,1,2,0.5,0.5,1,0.5,0.5),
  Measure=c("lb","ea","Tbs","ea","cup","cup","cup","cup","cup"), 
  Ingredient=c("beef brisket, flat half, raw, 0'' trim, URMIS 1623",
               "garlic cloves,fresh","oil, canola",
               "onion,yellow,fresh,large","wine,red",
               "stock,beef,prepared from recipe",
               "chili sauce,tomato","vinegar,cider","sugar,brown,dark"))
```

### Comment:
I have created a data frame of Sweet and Sour Brisket Recipe of year 2006 with column names as Amount, Measure, Ingredient.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@          Recipe : Hot Slaw       @@@@#
#@@@@            Year : 1936           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


Hot_Slaw_1936 <- data.frame(
  Amount=c(3,3,1,2,1),
  Measure=c("cup","Tbs","Tbs","Tbs","ea"),
  Ingredient=c("cabbage,fresh,shredded","vinegar,distilled",
               "sugar,white,granulated,granulated",
               "Butter, salted","apple,fresh,large,3 1/4''"))
```

### Comment:
I have created a data frame of Hot Slaw Recipe of year 1936 with column names as Amount, Measure, Ingredient.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@          Recipe : Hot Slaw       @@@@#
#@@@@            Year : 2006           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


Hot_Slaw_2006 <- data.frame(
  Amount=c(0.5,3,1,3),
  Measure=c("lb","Tbs","Tbs","cup"), 
  Ingredient=c("bacon,raw,cured","vinegar,cider",
               "sugar,brown,packed","cabbage,fresh,shredded"))
```

### Comment:
I have created a data frame of Hot Slaw Recipe of year 2006 with column names as Amount, Measure, Ingredient.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

# Matching Ingredient to the Long Description in the FOOD_DES

# Reading FOOD_DES.txt file
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")


# grep function is used for pattern matching
# rbind function is used for combine the data by rows

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  Recipe : Sweet and Sour Brisket @@@@#
#@@@@            Year : 1936           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

New_Ingredients_SSB_1936 <- rbind(
  
# Matching "beef brisket, flat half, raw, 0'' trim, URMIS 1623" #
# Matched "Beef, brisket, flat half, separable lean only, trimmed to 0" fat, select, cooked, braised"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^beef[[:punct:]].brisket[[:punct:]].flat.half[[:punct:]].separable.lean.only[[:punct:]].trimmed to 0" fat[[:punct:]].select',food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "butter,salad" #
# Matched "Butter, salted"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
'^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "onion, yellow,fresh,chopped" #
# Matched "Onions, yellow, sauteed"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^onions[[:punct:]].yellow',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "tomatoes,whole,canned,unsalted" #
# Matched "Tomatoes, red, ripe, canned, packed in tomato juice, no salt added"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^tomatoes[[:punct:]].red[[:punct:]].ripe[[:punct:]].canned[[:punct:]].packed.in.tomato juice[[:punct:]].no.salt.added',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "bell pepper,green,fresh,chopped" #
# Matched "Peppers, sweet, green, frozen, chopped, unprepared"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^peppers[[:punct:]].sweet[[:punct:]].green[[:punct:]].frozen[[:punct:]].chopped[[:punct:]].unprepared',
food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "celery,fresh, diced" #
# Matched "Celery, raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^celery[[:punct:]].raw',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "flour,all purpose,self rising,enriched" #
# Matched "Wheat flour, white, all-purpose, self-rising, enriched"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^wheat.flour[[:punct:]].white[[:punct:]].all-purpose[[:punct:]].self-rising',
food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "sugar,white,granulated,granulated" #
# Matched "Sugars, granulated"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "vinegar,distilled" #
# Matched "Vinegar, distilled"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
"^vinegar[[:punct:]].distilled",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "cream,whipping,heavy" #
# Matched "Cream, whipped, cream topping, pressurized"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^cream[[:punct:]].whipped",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)]
)

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  Recipe : Sweet and Sour Brisket @@@@#
#@@@@            Year : 2006           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

New_Ingredients_SSB_2006 <- rbind(

# Matching "beef brisket, flat half, raw, 0'' trim, URMIS 1623" #
# Matched "Beef, brisket, flat half, separable lean only, trimmed to 0" fat, select, cooked, braised"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^beef[[:punct:]].brisket[[:punct:]].flat.half[[:punct:]].separable.lean.only[[:punct:]].trimmed to 0" fat[[:punct:]].select',food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "garlic cloves,fresh" #
# Matched "Garlic, raw" #
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep("^garlic[[:punct:]]",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "oil, canola" #
# Matched "Oil, canola"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep("^oil[[:punct:]].canola", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "onion,yellow,fresh,large" #
# Matched "Onions, spring or scallions (includes tops and bulb), raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep("^onions[[:punct:]].spring", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "wine,red" #
# Matched "Alcoholic beverage, wine, table, red"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^Alcoholic beverage[[:punct:]].wine[[:punct:]].table[[:punct:]].red$", 
food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "stock,beef,prepared from recipe" #
# Matched "Soup, stock, beef, home-prepared"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^soup[[:punct:]].stock[[:punct:]].beef", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "chili sauce,tomato" #
# Matched "Sauce, tomato chili sauce, bottled, with salt"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^sauce[[:punct:]].tomato.chili.sauce", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "vinegar,cider" #
# Matched "Vinegar, cider"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep("^vinegar[[:punct:]].cider", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "sugar,brown,dark" #
# Matched "Sugar, turbinado"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^sugar[[:punct:]].turbinado",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
)

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@          Recipe : Hot Slaw       @@@@#
#@@@@            Year : 1936           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

New_Ingredients_HS_1936 <- rbind(
  
# Matching "cabbage,fresh,shredded" #
# Matched "Cabbage, raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^Cabbage[[:punct:]].raw",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)],

# Matching "vinegar,distilled" #
# Matched "Vinegar, distilled"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
"^vinegar[[:punct:]].distilled",food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "sugar,white,granulated,granulated" #
# Matched "Sugars, granulated"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^sugars[[:punct:]].granulated',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "butter,salad" #
# Matched "Butter, salted"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
'^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "apple,fresh,large,3 1/4''" #
# Matched "Apples, frozen, unsweetened, unheated (Includes foods for USDA's Food Distribution Program)"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^apples[[:punct:]].frozen[[:punct:]].unsweetened[[:punct:]].unheated",
food_des.dat$Long_Desc,ignore.case = TRUE, fixed = FALSE),c(1,3)]
)

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@          Recipe : Hot Slaw       @@@@#
#@@@@            Year : 2006           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

New_Ingredients_HS_2006 <- rbind(

# Matching "bacon,raw,cured" #
# Matched "Pork, cured, bacon, unprepared"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^pork[[:punct:]].cured[[:punct:]].bacon[[:punct:]].unprepared",food_des.dat$Long_Desc, 
ignore.case = TRUE, fixed = FALSE),c(1,3)],

# Matching "vinegar,cider" #
# Matched "Vinegar, cider"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
"^vinegar[[:punct:]].cider", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "sugar,brown,packed" #
# Matched "Sugar, turbinado"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^sugar[[:punct:]].turbinado",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)],

# Matching "cabbage,fresh,shredded" #
# Matched "Cabbage, raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^Cabbage[[:punct:]].raw",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
)

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@    Recipe : BeefStroganoff       @@@@#
#@@@@      Year : 1997                 @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


New_Ingredients_BS_1997 <- rbind(
  
# Matching "beef roast, tenderloin, raw, 0" trim" #
# Matched "Beef, loin, tenderloin roast, boneless, separable lean only, trimmed to 0" fat, select, raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^beef[[:punct:]].loin[[:punct:]].tenderloin.roast[[:punct:]].boneless[[:punct:]].separable.lean.only[[:punct:]].trimmed.to.0".fat[[:punct:]].select',food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)],

# Matching "butter,salad" #
# Matched "Butter, salted"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
'^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "flour,all purpose,self rising,enriched" #
# Matched "Wheat flour, white, all-purpose, self-rising, enriched"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^wheat.flour[[:punct:]].white[[:punct:]].all-purpose[[:punct:]].self-rising',
food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "stock, beef, prepared from recipe" #
# Matched "Soup, stock, beef, home-prepared"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^soup[[:punct:]].stock[[:punct:]].beef", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "sour cream, cultured" #
# Matched "Cream, sour, cultured"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^cream[[:punct:]].sour[[:punct:]].cultured",food_des.dat$Long_Desc, 
ignore.case = TRUE,  fixed = FALSE),c(1,3)],

# Matching "mustard, dijon" #
# Matched "Mustard, prepared, yellow"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^mustard[[:punct:]]|.(D|d)ijon",food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)]
)

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@    Recipe : BeefStroganoff       @@@@#
#@@@@      Year : 2006                 @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

New_Ingredients_BS_2006 <- rbind(
  
# Matching "beef roast, tenderloin, raw, 0" trim" #
# Matched "Beef, loin, tenderloin roast, boneless, separable lean only, trimmed to 0" fat, select, raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
'^beef[[:punct:]].loin[[:punct:]].tenderloin.roast[[:punct:]].boneless[[:punct:]].separable.lean.only[[:punct:]].trimmed.to.0".fat[[:punct:]].select',food_des.dat$Long_Desc, ignore.case = TRUE, fixed = FALSE),c(1,3)],

# Matching "oil, canola" #
# Matched "Oil, canola"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
"^oil[[:punct:]].canola", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "butter,salad" #
# Matched "Butter, salted"#
# It is Matched Exactly in the FOOD_DES  #
food_des.dat[grep(
'^butter[[:punct:]].salted',food_des.dat$Long_Desc,ignore.case = TRUE,fixed = FALSE), c(1,3)],

# Matching "onion, yellow, fresh, medium, 2 1/2\" #
# Matched "Onions, yellow, sauteed"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^onions[[:punct:]].yellow", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "mushrooms, fresh" #
# Matched "Mushrooms, white, raw"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^mushrooms[[:punct:]].white[[:punct:]].raw", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "stock, beef, prepared from recipe" #
# Matched "Soup, stock, beef, home-prepared"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^soup[[:punct:]].stock[[:punct:]].beef", food_des.dat$Long_Desc, ignore.case = TRUE),c(1,3)],

# Matching "sour cream, cultured" #
# Matched "Cream, sour, cultured"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep(
"^cream[[:punct:]].sour[[:punct:]].cultured",food_des.dat$Long_Desc, 
ignore.case = TRUE,  fixed = FALSE),c(1,3)],

# Matching "mustard, dijon" #
# Matched "Mustard, prepared, yellow"#
# It is Matched Appropriate or most suitable in the FOOD_DES  #
# It is Partially Matched #
food_des.dat[grep("^mustard[[:punct:]]|.(D|d)ijon",food_des.dat$Long_Desc, 
                  ignore.case = TRUE, fixed = FALSE),c(1,3)]
)


#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
```

### Comment:
In the above chunk i used a grep function which is used for pattern matching. And rbind function is used for combine the data by rows.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@   Printing data tables for your assigned recipes  @@@@#
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  Recipe : Sweet and Sour Brisket @@@@#
#@@@@            Year : 1936           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

Recipes = c("Sweet and Sour Brisket")
Year=c(1936)
Sweet_and_Sour_Brisket1936 <- data.frame(
  Recipes,Year,Sweet_and_Sour_Brisket_1936$Amount,
  Sweet_and_Sour_Brisket_1936$Measure,New_Ingredients_SSB_1936)
row.names(Sweet_and_Sour_Brisket1936) <- c()
colnames(Sweet_and_Sour_Brisket1936) <- c("Recipes","Year","Amount","Measure","NDB_No","Ingredient")
print(Sweet_and_Sour_Brisket1936)   # Printing Sweet_and_Sour_Brisket1936


#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@  Recipe : Sweet and Sour Brisket @@@@#
#@@@@            Year : 2006           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

Recipes = c("Sweet and Sour Brisket")
Year=c(2006)
Sweet_and_Sour_Brisket2006 <- data.frame(
  Recipes,Year,Sweet_and_Sour_Brisket_2006$Amount,
  Sweet_and_Sour_Brisket_2006$Measure,New_Ingredients_SSB_2006)
row.names(Sweet_and_Sour_Brisket2006) <- c()
colnames(Sweet_and_Sour_Brisket2006) <- c("Recipes","Year","Amount","Measure","NDB_No","Ingredient")
print(Sweet_and_Sour_Brisket2006) # Printing Sweet_and_Sour_Brisket2006


#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@          Recipe : Hot Slaw       @@@@#
#@@@@            Year : 1936           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

Recipes = c("Hot Slaw")
Year=c(1936)
Hot_Slaw1936 <- data.frame(Recipes,Year,Hot_Slaw_1936$Amount,
                           Hot_Slaw_1936$Measure,New_Ingredients_HS_1936)
row.names(Hot_Slaw1936) <- c()
colnames(Hot_Slaw1936) <- c("Recipes","Year","Amount","Measure","NDB_No","Ingredient")
print(Hot_Slaw1936)    # Printing Hot_Slaw1936


#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@          Recipe : Hot Slaw       @@@@#
#@@@@            Year : 2006           @@@@# 
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

Recipes = c("Hot Slaw")
Year=c(2006)
Hot_Slaw2006 <- data.frame(Recipes,Year,Hot_Slaw_2006$Amount,
                           Hot_Slaw_2006$Measure,New_Ingredients_HS_2006)
row.names(Hot_Slaw2006) <- c()
colnames(Hot_Slaw2006) <- c("Recipes","Year","Amount","Measure","NDB_No","Ingredient")
print(Hot_Slaw2006)     # Hot_Slaw2006
```

### Comment:
I have printing a Sweet and Sour Brisket & Hot Slaw Recipe of year 1936 & 2006 with column names as Amount, Measure, NDB_No, Ingredient. When i match Ingredient to the Long Description in the FOOD_DES, it shows me a corresponding column NBD_No. Finally printed the Sweet and Sour Brisket & Hot Slaw Recipe of year 1936 & 2006.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@   2. Writing each individual recipe to a tab delimited file     @@@@#
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


write.table(Sweet_and_Sour_Brisket1936,"Sweet_and_Sour_Brisket1936.tab",row.names = FALSE)
write.table(Sweet_and_Sour_Brisket2006,"Sweet_and_Sour_Brisket2006.tab",row.names = FALSE)
write.table(Hot_Slaw1936,"Hot_Slaw1936.tab",row.names = FALSE)
write.table(Hot_Slaw2006,"Hot_Slaw2006.tab",row.names = FALSE)
```

### Comment:
After creating the data tables i write them into a file with .tab extension.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@   3. Reading each tab file, and merging into the tables     @@@@#
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#


# Reading BeefStroganoff Recipies from .tab extension files

BeefStroganoff.1997 = read.table("BeefStroganoff.1997.tab",header = TRUE,stringsAsFactors = FALSE) 
BeefStroganoff.2006 = read.table("BeefStroganoff.2006.tab",header = TRUE,stringsAsFactors = FALSE)

# BeefStroganoff.1997 Recipe
BeefStroganoff.1997$Ingredient <- NULL
Recipes = c("BeefStroganoff")
Year=c(1997)
Combine_Beef97 <- cbind(Recipes,Year,BeefStroganoff.1997,New_Ingredients_BS_1997)
row.names(Combine_Beef97) <- c()
colnames(Combine_Beef97) <- c("Recipes","Year","Amount","Measure","NDB_No","Ingredient")
# Printing BeefStroganoff.1997 Recipe
Combine_Beef97

# BeefStroganoff.2006 Recipe
BeefStroganoff.2006$Ingredient <- NULL
Recipes = c("BeefStroganoff")
Year=c(2006)
Combine_Beef06 <- cbind(Recipes,Year,BeefStroganoff.2006,New_Ingredients_BS_2006)
row.names(Combine_Beef06) <- c()
colnames(Combine_Beef06) <- c("Recipes","Year","Amount","Measure","NDB_No","Ingredient")
# Printing BeefStroganoff.2006 Recipe
Combine_Beef06

# By using rbind function combine the data 
Recipes_1936 = rbind(Sweet_and_Sour_Brisket1936,Hot_Slaw1936)
Recipes_2006 = rbind(Sweet_and_Sour_Brisket2006,Hot_Slaw2006)
Merging_with_BeefStroganoff_1936_1997 = rbind(Recipes_1936,Combine_Beef97)
Merging_with_BeefStroganoff_2006 = rbind(Recipes_2006,Combine_Beef06) 

# Merge all recipes using rbind function
Merge_All <- rbind(Recipes_1936,Recipes_2006,Combine_Beef97,Combine_Beef06)
print(Merge_All)
```

### Comment:
In this chunk i read the each BeefStroganoff Recipe. After reading the BeefStroganoff Recipe i am attaching NDB No when the Ingredient pattern is matched.Then print the BeefStroganoff Recipe.After that by using rbind function combine the different recipe by years.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@   4. Proof the data to check errors     @@@@#
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

# Sweet and Sour Brisket and Hot Slaw Recipes merged by ingredient
# And printed the Sweet and Sour Brisket and Hot Slaw Recipes

Sweet_and_Sour_Brisket_Ingredient <- merge(Sweet_and_Sour_Brisket1936,
                                           Sweet_and_Sour_Brisket2006, by = "Ingredient")
Hot_Slaw_Ingredient <- merge(Hot_Slaw1936,Hot_Slaw2006,by = "Ingredient")



# Plot Measure.x vs Measure.y and see quickly where the measures are not consistent

# Plot for Sweet_and_Sour_Brisket_Ingredient of Measure.x vs Measure.y
plot(Sweet_and_Sour_Brisket_Ingredient$Measure.x,
     Sweet_and_Sour_Brisket_Ingredient$Measure.y,
     las = 3,xlab = "Measure 1936", ylab = "Measure 2006",
     main = "Sweet and Sour Brisket",col = "green")

# Plot for Hot_Slaw_Ingredient of Measure.x vs Measure.y
plot(Hot_Slaw_Ingredient$Measure.x,
     Hot_Slaw_Ingredient$Measure.y,
     las = 3,xlab = "Measure 1936", 
     ylab = "Measure 2006",main = "Hot Slaw",col = "blue")


# Created a contingency table comparing Measure.x and Measure.y to look for mismatches in measures between ingredients

# Contingency table for Sweet_and_Sour_Brisket_Ingredient of Measure.x and Measure.y
table(Sweet_and_Sour_Brisket_Ingredient$Measure.x,Sweet_and_Sour_Brisket_Ingredient$Measure.y)

# Contingency table for Hot_Slaw_Ingredient of Measure.x and Measure.y
table(Hot_Slaw_Ingredient$Measure.x,Hot_Slaw_Ingredient$Measure.y)



# Plot Amount.x and Amount.y to look for unexpected changes in values

# Plot for Sweet_and_Sour_Brisket_Ingredient of Amount.x and Amount.y
plot(Sweet_and_Sour_Brisket_Ingredient$Amount.x,
     Sweet_and_Sour_Brisket_Ingredient$Amount.y,
     las = 3,xlab = "Amount 1936", ylab = "Amount 2006",
     main = "Sweet and Sour Brisket",col = "green")

# Plot for Hot_Slaw_Ingredient of Amount.x and Amount.y
plot(Hot_Slaw_Ingredient$Amount.x,
     Hot_Slaw_Ingredient$Amount.y,
     las = 3,xlab = "Amount 1936", 
     ylab = "Amount 2006",main = "Hot Slaw",col = "blue")
```

### Comment:
The main concept of this chunk is proof of data to check errors. Here by using merge function i merged the recipe by ingredient with different years and printed it.Then plotted Measure.x vs Measure.y and see quickly where the measures are consistent or not. And created a contingency table comparing Measure.x and Measure.y to look for mismatches in measures between ingredients. Then finally i taken another plotting between Amount.x and Amount.y to look for unexpected changes in values.


```{r}
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#
#@@@@   5. Reading & appending rows for your recipes     @@@@#
#@@@@   and then write the table back to Recipes.csv     @@@@#
#@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@#

# My Recipes with their year's,MinServings & MaxServings 
Recipe <- data.frame(
  Recipe=c("Sweet and Sour Brisket","Sweet and Sour Brisket","Hot Slaw","Hot Slaw"),
  Year=c(1936,2006,1936,2006),
  MinServings=c(4,8,6,6),
  MaxServings=c(4,8,6,6))
print(Recipe)    # Print Recipe

# Reading BeefStroganoff_Recipe csv file
BeefStroganoff_Recipe <- read.csv(
  "Recipes.csv",
  header=TRUE)
print(BeefStroganoff_Recipe)  # Printing BeefStroganoff_Recipe 

# Combine BeefStroganoff_Recipe & Recipe using rbind function
Combine <- rbind(BeefStroganoff_Recipe,Recipe)

# Write the combine to Recipes.csv file
write.csv(Combine, file="Recipes.csv",row.names = FALSE)

# Read the Recipes.csv file
Recipe.Table <- read.csv("Recipes.csv",header=TRUE)
print(Recipe.Table)  # Printing Recipe.Table
```

### Comment:
The concept of creating this chunk is for reading & appending rows for your recipes and then write the table back to Recipes.csv. Here first i created a data frame with my recipes and combine with BeefStroganoff_Recipe and write the table back to Recipes.csv and finally printed the table 


# Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.
```{r}
# Reading WEIGHT.txt file
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)   # Summary of weight.dat
```

### Comment:
Here in the above chunk i read the WEIGHT.txt file and printed the head and summary.


```{r}
# Reading FOOD_DES.txt File
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)   # Summary of food_des.dat
```

### Comment:
Here in the above chunk i read the FOOD_DES.txt file and printed the head and summary.


```{r}
# Reading NUT_DATA.txt File
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)   # Summary of nut_data.dat
```

### Comment:
Here in the above chunk i read the NUT_DATA.txt file and printed the head and summary.

