---
title: "Midterm Project"
author: "Set 34"
date: "July 18, 2018"
output: word_document
---

# Part 1 - Create Data Tables for the two recipes assigned.

**In this section of code, we are creating four different data frames. Two tables per recipe, one for each year given, totalling four tables at the end.**
```{r}
ScallopedOysters.1936 <- data.frame(
  Amount=c(1,6,0.5,28,0.5),
  Measure = c("pnt","Tbs","cup","ea","cup"),
  Ingredient = c("oysters, eastern, in liquid, 12 oz can","cream, whipping, heavy","Bread, crumbs, dry, grated, plain","crackers, saltine","butter, salted")
)

ScallopedOysters.2006 <- data.frame(
  Amount=c(1,2,1,0.75,1.00),
  Measure=c("qt","cup","cup","cup","cup"),
  Ingredient=c("oysters, eastern, in liquid, 12 oz can","crackers, saltine, crushed","Bread, crumbs, dry, grated, plain","butter, salted","cream, whipping, heavy")
)

CottagePudding.1936 <- data.frame(
  Amount=c(0.5,0.25,1,1,1.5,0.5),
  Measure = c("cup","cup","lrg","tsp","cup","cup"),
  Ingredient = c("sugar, white, granulated","butter, salted","egg, raw","extract, vanilla","flour, bread, white, enriched","milk, whole, 3.25%")
)

CottagePudding.2006 <- data.frame(
  Amount=c(1.5,1,2,0.5,0.5),
  Measure = c("cup","cup","lrg","cup","cup"),
  Ingredient = c("flour, all purpose, self rising, enriched","sugar, white, granulated","egg, raw","milk, whole, 3.25%","butter, salted")
)
```

# Part 2 - Export to tab file and re-import, then add in columns for recipe name and year. 

**In this section of code we are creating the four different .tab files, one for each table we created above.**
```{r}
write.table(ScallopedOysters.1936, "ScallopedOysters.1936.tab", sep = " ", row.names = FALSE)
write.table(ScallopedOysters.2006, "ScallopedOysters.2006.tab", sep = " ", row.names = FALSE)
write.table(CottagePudding.1936, "CottagePudding.1936.tab", sep = " ", row.names = FALSE)
write.table(CottagePudding.2006, "CottagePudding.2006.tab", sep = " ", row.names = FALSE)
```

**In this section of code, we are reading back in the .tab files, and then we are printing them out to show that they exported and imported correctly.**
```{r}
ScallopedOysters1936 <- read.csv("ScallopedOysters.1936.tab",header=TRUE, sep = " ")
ScallopedOysters2006 <- read.csv("ScallopedOysters.2006.tab",header=TRUE, sep = " ")
CottagePudding1936 <- read.csv("CottagePudding.1936.tab",header=TRUE, sep = " ")
CottagePudding2006 <- read.csv("CottagePudding.2006.tab",header=TRUE, sep = " ")
print(ScallopedOysters1936)
print(ScallopedOysters2006)
print(CottagePudding1936)
print(CottagePudding2006)
```
**In this section of code we are using the recipe list that is in our file explorer folder, and we are using the substring function to pull out different sections of the file name so we can make the Recipe and Year column in the code below.**
```{r}
recipeList <- list.files(path = "./", pattern = '*.tab')
recipeList

scalloped1936 <- substring(recipeList[5], 1, 9)
scalloped2006 <- substring(recipeList[6], 1, 9)
oysters1936 <- substring(recipeList[5],10,16)
oysters2006 <- substring(recipeList[6],10,16)
soyear1936 <- as.numeric(substring(recipeList[5],18,21))
soyear2006 <- as.numeric(substring(recipeList[6],18,21))

cottage1936 <- substring(recipeList[3], 1, 7)
cottage2006 <- substring(recipeList[4], 1, 7)
pudding1936 <- substring(recipeList[3],8,14)
pudding2006 <- substring(recipeList[4],8,14)
cpyear1936 <- as.numeric(substring(recipeList[3],16,19))
cpyear2006 <- as.numeric(substring(recipeList[4],16,19))
```
**In this section of code we are using the variables we created above, along wite cbind() and paste() functions to create the Recipe and Year columns to add to our four tables.**
```{r}
ScallopedOysters1936 <- cbind(Recipe = paste(scalloped1936," ",oysters1936), Year = soyear1936, ScallopedOysters1936)
ScallopedOysters2006 <- cbind(Recipe = paste(scalloped2006," ",oysters2006), Year = soyear2006, ScallopedOysters2006)
CottagePudding1936 <- cbind(Recipe = paste(cottage1936," ",pudding1936), Year = cpyear1936, CottagePudding1936)
CottagePudding2006 <- cbind(Recipe = paste(cottage2006," ",pudding2006), Year = cpyear2006, CottagePudding2006)
```
**In this section of code, we are creating the data frame that contains the Recipe, Year, Minservings, and MaxServings columns. Eventually we will add grams and calories onto that.**
```{r}
recipe <- data.frame(
  Recipe=c("Scalloped Oysters","Scalloped Oysters","Cottage Pudding","Cottage Pudding"),
  Year=c(1936,2006,1936,2006),
  MinServings=c(6,6,6,8),
  MaxServings=c(6,6,6,10)
)
recipe
```

#Part 3 - Import and Merge Beef Stroganoff Tables

**In this section we are reading in the Beef Stroganoff recipe table that is similar to the data frame we created in the code above.**
```{r}
recipes <- read.table("Recipes.csv",header = TRUE, sep = ",")
recipes
```
**In this section of code we are reading in the two Beef Stroganoff recipe .tab files that Dr. Claussen provided for this project.**
```{r}
BeefStroganoff1997 <- read.table("BeefStroganoff.1997.tab",header=TRUE, sep = "")
BeefStroganoff2006 <- read.table("BeefStroganoff.2006.tab",header=TRUE, sep = "")
BeefStroganoff1997
BeefStroganoff2006
```
**In this section of code we are creating the Recipe and Year columns like we did earlier in the code with the cbind() function.**
```{r}
BeefStroganoff1997 <- cbind(Recipe = "Beef Stroganoff", Year = 1997, BeefStroganoff1997)
BeefStroganoff2006 <- cbind(Recipe = "Beef Stroganoff", Year = 2006, BeefStroganoff2006)
BeefStroganoff1997
BeefStroganoff2006
```
**In this section of the code we are merging together all of the data frames into one data frame with all 6 recipes, from 4 different years, with amount, measure, and ingredient.**
```{r}
ScallopedOysters <- merge.data.frame(ScallopedOysters1936,ScallopedOysters2006,all=TRUE)
CottagePudding <- merge.data.frame(CottagePudding1936,CottagePudding2006,all=TRUE)
MyRecipes <- merge.data.frame(ScallopedOysters,CottagePudding,all=TRUE)

BeefStroganoff <- merge.data.frame(BeefStroganoff1997,BeefStroganoff2006,all=TRUE)

AllRecipes <- merge.data.frame(MyRecipes,BeefStroganoff,all=TRUE)
AllRecipes
```
**In this section of code, we merge together the two recipe files for one file, with 6 different recipes with the min and max servings for each one of them.**
```{r}
ServingRecipe <- merge.data.frame(recipe, recipes, all=TRUE)
ServingRecipe
```


# Part 4 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.

**In this section of code we are reading in the weight.txt file.**
```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)
```
**In this section of code we are reading in the food_des.txt file.**
```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```
**In this section of code we are reading in the nut_data.txt file.**
```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)
```
**In this section of code we searched for the correct description of bread crumbs and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(B|b)read, crumbs,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Bread, crumbs, dry, grated, plain')
```
**In this section of code we searched for the correct description of salted butter and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(B|b)utter, salted', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Butter, salted')
```
**In this section of code we searched for the correct description of canned oysters and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('Mollusks, oyster,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Mollusks, oyster, eastern, canned')
```
**In this section of code we searched for the correct description of heavy whipping cream and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('Cream, ', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Cream, fluid, heavy whipping')
```
**In this section of code we searched for the correct description of saltine crackers and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('saltine', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Crackers, saltines (includes oyster, soda, soup)')
```
**In this section of code we searched for the correct description of whole milk and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(M|m)ilk, whole,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D')
```
**In this section of code we searched for the correct description of sugar and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(S|s)ugars,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Sugars, granulated')
```
**In this section of code we searched for the correct description of eggs and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(E|e)gg,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Egg, whole, raw, fresh')
```
**In this section of code we searched for the correct description of flour and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(F|f)lour,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Wheat flour, white, all-purpose, self-rising, enriched')
```
**In this section of code we searched for the correct description of vanilla and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(E|e)xtract,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Vanilla extract, imitation, no alcohol')
```
**In this section of code we searched for the correct description of beef stock and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(S|s)tock,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Soup, stock, beef, home-prepared')
```
**In this section of code we searched for the correct description of mustard and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(M|m)ustard, prepared, yellow', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Mustard, prepared, yellow')
```
**In this section of code we searched for the correct description of sour cream and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(C|c)ream, sour,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Cream, sour, cultured')
```
**In this section of code we searched for the correct description of beef roast and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(B|b)eef, loin, tenderloin roast, boneless, separable lean and fat, trimmed to 0\" fat, all grades, raw', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Beef, loin, tenderloin roast, boneless, separable lean and fat, trimmed to 0\" fat, all grades, raw')
```
**In this section of code we searched for the correct description of mushrooms and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(M|m)ushrooms,', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Mushrooms, white, raw')
```
**In this section of code we searched for the correct description of canola oil and grabbed the NDB no for it and put it in our excel spreadsheet to track the changes.**
```{r}
grep('(O|o)il, canola', food_des.dat$Long_Desc, value = TRUE)
subset(food_des.dat,food_des.dat$Long_Desc == 'Oil, canola')
```
**In this section of code we are creating a vector for our AllRecipes table that will add in the NDB_No**
```{r}
NDB_No <- as.numeric(c(18079,1001,15170,1053,18228,1001,18079,1053,15170,18228,1001,1211,19335,1123,2052,20082,1001,1211,19335,20082,1123,6170,20082,2046,23336,1056,1001,6170,11282,2046,11260,4582,23336,1056,1001))
NDB_No
```
**In this section of code we added on the NDB_No field **
```{r}
AllRecipesWithNDBNo <- cbind(AllRecipes,NDB_No)
AllRecipesWithNDBNo
```
**In this section of code we are using the NDB_No to find the corresponding field descriptions and putting them into a vector and replace the current Ingredient column with correct data.**
```{r}
AllRecipesWithNDBNo$Ingredient <- food_des.dat[match(AllRecipesWithNDBNo$NDB_No,food_des.dat$NDB_No),3]
AllRecipesWithNDBNo
```
**This is our final result.**
```{r}
AllRecipesWithNDBNo
ServingRecipe
```