---
title: "Midterm"
author: "Set 42"
date: "July 21, 2018"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

# Part 1 - Assign Recipes

First, generate a list of all files in the Recipes folder.
```{r,eval=FALSE}
recipes <- list.files(path = "./", pattern = '*.pdf')
print(recipes)
```

Make this reproducible.
```{r,eval=FALSE}
set.seed(10010)
```

*loading stringr and data.table library for later use in file*
```{r}
#PMC - we won't create dependencies; replaced str_trim with trimws
#library(stringr)
#library(data.table)
```




*reading in food_des table for later comparison*
```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
#PMC - this slows typesetting
#print(food_des.dat)
```

*to use the variable long_desc for comparison for code later to receipes ingredient description*
```{r}
food_des.dat$Long_Desc <-as.character(food_des.dat$Long_Desc)
```

*created sauerbrauten 1936 recipe*
```{r}
Sauerbrauten.1936 <- data.frame(Amount =c("3","1","1","1/2","2","1/2","1","2","1"), 
                                Unit=c("lb","cup","cup","cup","Tbs","cup","cup","ea","tsp"), 
                                Ingredient=c("beef roast, chuck arm, raw, choice, 1/8\" trim","vinegar, distilled","water","onion,yellow,fresh,sliced","sugar,white,granulated","flour,all purpose, self rising, enriched","sour cream, cultured","bay leaves","black peppercorns"))
print(Sauerbrauten.1936)
```


*created sauerbrauten 2006 recipe*
```{r}
Sauerbrauten.2006 <- data.frame(Amount =c("3","1/4","1/2","1","1/2","2","1/4","1","1","2","2"), 
                                Unit=c("lb","cup","cup","Tbs","cup","Tbs","cup","cup","tsp","tsp","ea"), 
                                Ingredient=c("beef roast, chuck arm, raw, choice, 1/8\" trim","vinegar, distilled","water","sugar, white,granulated","onion,yellow,fresh,sliced","oil,canola","sugar,brown,packed","sour cream,cultured","black peppercorns","caraway seeds","bay leaves"))
print(Sauerbrauten.2006)
```
*created appledumpling 1937 recipe*
```{r}
AppleDumpling.1937 <- data.frame(Amount =c("6","2 1/2","12","2","2"), 
                                Unit=c("ea","Tbs","tsp","tsp","tsp"), 
                                Ingredient=c("apple,fresh, medium, 3'\'","brown,sugar,packed","butter, unsalted, softened","salt","ground cinnamon"))
print(AppleDumpling.1937)
```

*created apple dumpling 2006 recipe*
```{r}
AppleDumpling.2006 <- data.frame(Amount =c("6","1/2","1","1/4","1/4","1"), 
                                Unit=c("ea","cup","tsp","tsp","cup","ea"), 
                                Ingredient=c("apple,fresh,small,2 3'/'4","sugar,brown,dark","ground cinnamon","salt","unsalted butter,softened","egg,lightly,beaten"))
print(AppleDumpling.2006)
```
*loop through to determine matches for appledumpling 1937, have created comments in code*
```{r, eval=TRUE}
AppleDumpling.1937$Ingredient<-as.character(AppleDumpling.1937$Ingredient)
for(x in AppleDumpling.1937$Ingredient){ 
  #loop over ingredients
    
    #check ingredient in food_des
    if(sum(grepl(x,food_des.dat$Long_Desc)) > 0){
        print('Found a direct match')
        print(paste0('Our Ingredient: ',x))
        print('Ingredient found in FOOD DESC file')
        print(food_des.dat[grepl(x,food_des.dat$Long_Desc),'Long_Desc'])
        } else if (sum(grepl(tolower(x),tolower(food_des.dat$Long_Desc))) > 0){
            #convert ingredients to lowercase as well as food_des and then search
            #to find food_des lower
            print('Found a lowercase match')
            print(paste0('Our Ingredient: ',x))
            print('Ingredient found in FOOD DESC file')
            
            print(food_des.dat[grepl(tolower(x),tolower(food_des.dat$Long_Desc)),'Long_Desc'])
            }
                else{
                    # split ingredient string to search each word separately for partial match
                    
                    print(paste0('Not direct match found for: ',x, 
                     ', hence splitting ingredients into words:'))
                    single_words<-trimws(unlist(strsplit(x,',')))
                    print(single_words)
                    #   this will be in the form of a loop
                    for (sw in single_words){
                        # check if the word is present as it is
                        if(sum(grepl(sw,food_des.dat$Long_Desc)) > 0){
                            print('Found a partial match')
                            print(paste0('Our Partial word: ',sw))
                            print('Ingredient found in FOOD DESC file')
                            print(head(food_des.dat[grepl(sw,food_des.dat$Long_Desc),'Long_Desc']))
                            }else if(sum(grepl(tolower(sw),tolower(food_des.dat$Long_Desc))) > 0){
                                #check is word is present as lower case
                                print('Found a lowercase partial match')
                                print(paste0('Our lowercase partial word: ',sw))
                                print('Ingredient found in FOOD DESC file')
                                print(head(food_des.dat[grepl(tolower(sw),tolower(food_des.dat$Long_Desc)),'Long_Desc']))
                                } else{
                                    print('No partial match found')
                                    } 
                        
                        }
                        
                        }
}

```

*loop through to determine matches for appledumpling 2006, have created comments in code*
```{r, eval=FALSE}
AppleDumpling.2006$Ingredient<-as.character(AppleDumpling.2006$Ingredient)
for(x in AppleDumpling.2006$Ingredient){ 
  #loop over ingredients
    
    #check ingredient in food_des
    if(sum(grepl(x,food_des.dat$Long_Desc)) > 0){
        print('Found a direct match')
        print(paste0('Our Ingredient: ',x))
        print('Ingredient found in FOOD DESC file')
        print(food_des.dat[grepl(x,food_des.dat$Long_Desc),'Long_Desc'])
        } else if (sum(grepl(tolower(x),tolower(food_des.dat$Long_Desc))) > 0){
            #convert ingredients to lowercase as well as food_des and then search
            #to find food_des lower
            print('Found a lowercase match')
            print(paste0('Our Ingredient: ',x))
            print('Ingredient found in FOOD DESC file')
            
            print(food_des.dat[grepl(tolower(x),tolower(food_des.dat$Long_Desc)),'Long_Desc'])
            }
                else{
                    # split ingredient string to search each word separately for partial match
                    
                    print(paste0('Not direct match found for: ',x, 
                     ', hence splitting ingredients into words:'))
                    single_words<-trimws(unlist(strsplit(x,',')))
                    print(single_words)
                    #   this will be in the form of a loop
                    for (sw in single_words){
                        # check if the word is present as it is
                        if(sum(grepl(sw,food_des.dat$Long_Desc)) > 0){
                            print('Found a partial match')
                            print(paste0('Our Partial word: ',sw))
                            print('Ingredient found in FOOD DESC file')
                            print(head(food_des.dat[grepl(sw,food_des.dat$Long_Desc),'Long_Desc']))
                            }else if(sum(grepl(tolower(sw),tolower(food_des.dat$Long_Desc))) > 0){
                                #check is word is present as lower case
                                print('Found a lowercase partial match')
                                print(paste0('Our lowercase partial word: ',sw))
                                print('Ingredient found in FOOD DESC file')
                                print(head(food_des.dat[grepl(tolower(sw),tolower(food_des.dat$Long_Desc)),'Long_Desc']))
                                } else{
                                    print('No partial match found')
                                    } 
                        
                        }
                        
                        }
}

```


*loop through to determine matches for sauerbrauten 1936, have created comments in code*
```{r, eval=TRUE}
Sauerbrauten.1936$Ingredient<-as.character(Sauerbrauten.1936$Ingredient)
for(x in Sauerbrauten.1936$Ingredient){ 
  #loop over ingredients
    
    #check ingredient in food_des
    if(sum(grepl(x,food_des.dat$Long_Desc)) > 0){
        print('Found a direct match')
        print(paste0('Our Ingredient: ',x))
        print('Ingredient found in FOOD DESC file')
        print(food_des.dat[grepl(x,food_des.dat$Long_Desc),'Long_Desc'])
        } else if (sum(grepl(tolower(x),tolower(food_des.dat$Long_Desc))) > 0){
            #convert ingredients to lowercase as well as food_des and then search
            #to find food_des lower
            print('Found a lowercase match')
            print(paste0('Our Ingredient: ',x))
            print('Ingredient found in FOOD DESC file')
            
            print(food_des.dat[grepl(tolower(x),tolower(food_des.dat$Long_Desc)),'Long_Desc'])
            }
                else{
                    # split ingredient string to search each word separately for partial match
                    
                    print(paste0('Not direct match found for: ',x, 
                     ', hence splitting ingredients into words:'))
                    single_words<-trimws(unlist(strsplit(x,',')))
                    print(single_words)
                    #   this will be in the form of a loop
                    for (sw in single_words){
                        # check if the word is present as it is
                        if(sum(grepl(sw,food_des.dat$Long_Desc)) > 0){
                            print('Found a partial match')
                            print(paste0('Our Partial word: ',sw))
                            print('Ingredient found in FOOD DESC file')
                            print(head(food_des.dat[grepl(sw,food_des.dat$Long_Desc),'Long_Desc']))
                            }else if(sum(grepl(tolower(sw),tolower(food_des.dat$Long_Desc))) > 0){
                                #check is word is present as lower case
                                print('Found a lowercase partial match')
                                print(paste0('Our lowercase partial word: ',sw))
                                print('Ingredient found in FOOD DESC file')
                                print(head(food_des.dat[grepl(tolower(sw),tolower(food_des.dat$Long_Desc)),'Long_Desc']))
                                } else{
                                    print('No partial match found')
                                    } 
                        
                        }
                        
                        }
}

```

*loop through to determine matches for sauerbauten 2006, have created comments in code*
```{r, eval=TRUE}
Sauerbrauten.2006$Ingredient<-as.character(Sauerbrauten.2006$Ingredient)
for(x in Sauerbrauten.2006$Ingredient){ 
  #loop over ingredients
    
    #check ingredient in food_des
    if(sum(grepl(x,food_des.dat$Long_Desc)) > 0){
        print('Found a direct match')
        print(paste0('Our Ingredient: ',x))
        print('Ingredient found in FOOD DESC file')
        print(food_des.dat[grepl(x,food_des.dat$Long_Desc),'Long_Desc'])
        } else if (sum(grepl(tolower(x),tolower(food_des.dat$Long_Desc))) > 0){
            #convert ingredients to lowercase as well as food_des and then search
            #to find food_des lower
            print('Found a lowercase match')
            print(paste0('Our Ingredient: ',x))
            print('Ingredient found in FOOD DESC file')
            
            print(food_des.dat[grepl(tolower(x),tolower(food_des.dat$Long_Desc)),'Long_Desc'])
            }
                else{
                    # split ingredient string to search each word separately for partial match
                    
                    print(paste0('Not direct match found for: ',x, 
                     ', hence splitting ingredients into words:'))
                    single_words<-trimws(unlist(strsplit(x,',')))
                    print(single_words)
                    #   this will be in the form of a loop
                    for (sw in single_words){
                        # check if the word is present as it is
                        if(sum(grepl(sw,food_des.dat$Long_Desc)) > 0){
                            print('Found a partial match')
                            print(paste0('Our Partial word: ',sw))
                            print('Ingredient found in FOOD DESC file')
                            print(head(food_des.dat[grepl(sw,food_des.dat$Long_Desc),'Long_Desc']))
                            }else if(sum(grepl(tolower(sw),tolower(food_des.dat$Long_Desc))) > 0){
                                #check is word is present as lower case
                                print('Found a lowercase partial match')
                                print(paste0('Our lowercase partial word: ',sw))
                                print('Ingredient found in FOOD DESC file')
                                print(head(food_des.dat[grepl(tolower(sw),tolower(food_des.dat$Long_Desc)),'Long_Desc']))
                                } else{
                                    print('No partial match found')
                                    } 
                        
                        }
                        
                        }
}

```


```{r}
Sauerbrauten.1936 <- data.frame(Amount =c("3","1","1","1/2","2","1/2","1","2","1"), 
                                Unit=c("lb","cup","cup","cup","Tbs","cup","cup","ea","tsp"), 
                                Ingredient=c("beef roast, chuck arm, raw, choice, 1/8\" trim","vinegar, distilled","water","onion,yellow,fresh,sliced","sugar,white,granulated","flour,all purpose, self rising, enriched","sour cream, cultured","bay leaves","black peppercorns"),
                                NDB_No=c("13808","02053","14555","11282","19335","20481","01074","02004","02030"
))
print(Sauerbrauten.1936)
```


```{r}
Sauerbrauten.2006 <- data.frame(Amount =c("3","1/4","1/2","1","1/2","2","1/4","1","1","2","2"), 
                                Unit=c("lb","cup","cup","Tbs","cup","Tbs","cup","cup","tsp","tsp","ea"), 
                                Ingredient=c("beef roast, chuck arm, raw, choice, 1/8\" trim","vinegar, distilled","water","sugar, white,granulated","onion,yellow,fresh,sliced","oil,canola","sugar,brown,packed","sour cream,cultured","black peppercorns","caraway seeds","bay leaves"),
                                NDB_No=c("13809","02053","14555","19335","11282","04582","19334","01074","02030","02005","02004"
))
print(Sauerbrauten.2006)
```


```{r}
AppleDumpling.1937 <- data.frame(Amount =c("6","2 1/2","12","2","2"), 
                                Unit=c("ea","Tbs","tsp","tsp","tsp"), 
                                Ingredient=c("apple,fresh, medium, 3'\'","brown,sugar,packed","butter, unsalted, softened","salt","ground cinnamon"),
                                NDB_No=c("09004","19334","01145","02047","02010"))
print(AppleDumpling.1937)
```



```{r}
AppleDumpling.2006 <- data.frame(Amount =c("6","1/2","1","1/4","1/4","1"), 
                                Unit=c("ea","cup","tsp","tsp","cup","ea"), 
                                Ingredient=c("apple,fresh,small,2 3'/'4","sugar,brown,dark","ground cinnamon","salt","unsalted butter,softened","egg,lightly,beaten"),
                                NDB_No=c("09004","19334","02010","02047","01145","01125"))
print(AppleDumpling.2006)
```



*write tab files for recipes AppleDumpling 1937 and 2006 and Sauerbrauten 1936 and 2006*
```{r}
write.table(AppleDumpling.1937, "AppleDumpling.1937.tab", sep="")
write.table(AppleDumpling.2006, "AppleDumpling.2006.tab", sep="")
write.table(Sauerbrauten.1936, "Sauerbrauten.1936.tab", sep="")
write.table(Sauerbrauten.2006, "Sauerbrauten.2006.tab", sep="")
```

*reading in the recipe files I just exported*
```{r}
AppleDumpling.1937 <- read.table("AppleDumpling.1937.tab", header=TRUE)
AppleDumpling.2006 <- read.table("AppleDumpling.2006.tab", header=TRUE)
Sauerbrauten.1936 <- read.table("Sauerbrauten.1936.tab", header=TRUE)
Sauerbrauten.2006 <- read.table("Sauerbrauten.2006.tab", header=TRUE)
```


*merging the appledumpling 1937 and 2006*
```{r}
new.AppleDumpling <-merge.data.frame(AppleDumpling.1937, AppleDumpling.2006,by.x="Ingredient",by.y="Ingredient",suffixes=c('.1937','.2006'))
print(new.AppleDumpling)
```

*merging the sauerbrauten 1936 and 2006*
```{r}
new.Sauerbrauten <-merge.data.frame(Sauerbrauten.1936, Sauerbrauten.2006,by.x="Ingredient",by.y="Ingredient",suffixes=c('.1937','.2006'))
print(new.Sauerbrauten)
```
*read in both tab beef stroganoff files*
```{r}
BeefStroganoff.1997 <- read.table("BeefStroganoff.1997.tab", header=TRUE)
print(BeefStroganoff.1997)
BeefStroganoff.2006 <- read.table("BeefStroganoff.2006.tab", header=TRUE)
print(BeefStroganoff.2006)
```

*merge beef stroganoff files*
```{r}
new.BeefStroganoff <-merge.data.frame(BeefStroganoff.1997, BeefStroganoff.2006,by.x="Ingredient",by.y="Ingredient",suffixes=c('.1937','.2006'))
print(new.BeefStroganoff)
```

*Differences in measures apple dumpling*
```{r}
table(new.AppleDumpling$Unit.1937,new.AppleDumpling$Unit.2006) 
plot(new.AppleDumpling$Unit.1937,new.AppleDumpling$Unit.2006)
```

*differences in measures sauerbrauten*
```{r}
table(new.Sauerbrauten$Unit.1937,new.Sauerbrauten$Unit.2006) 
plot(new.Sauerbrauten$Unit.1937,new.Sauerbrauten$Unit.2006)
```

*differences in amount appledumpling*
```{r}
table(new.AppleDumpling$Amount.1937,new.AppleDumpling$Amount.2006) 
plot(new.AppleDumpling$Amount.1937,new.AppleDumpling$Amount.2006)
```

*differences in amount sauerbrauten*
```{r}
table(new.Sauerbrauten$Amount.1937,new.Sauerbrauten$Amount.2006) 
plot(new.Sauerbrauten$Amount.1937,new.Sauerbrauten$Amount.2006)
```




*add my data from my servings for my recipes sauerbrauten and apple dumpling, print and exported new recipe file*
```{r}
recipes <- read.table('Recipes.csv', sep=',',header=TRUE)
print(recipes)
row1 <-data.frame(Recipe="AppleDumpling", Year=1937, MinServings=1, MaxServings=1  )
row2 <-data.frame(Recipe="AppleDumpling", Year=2006, MinServings=6, MaxServings=6)
row3<-data.frame(Recipe="Sauerbrauten", Year=2006, MinServings=6, MaxServings=6)
row4<-data.frame(Recipe="Sauerbrauten", Year=1936, MinServings=12, MaxServings=12)
new.recipes <- rbind.data.frame(recipes,row1,row2,row3,row4)
print(new.recipes)
write.table(new.recipes, 'New Recipes.csv', sep=",")
                
```


