---
title: "Joy Of Cooking Midterm Project Prep"
author: "Set 41"
date: "7/11/2018"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

** Assigned Recipes : **                  
1) Meat Loaf [The PDF contains 2 recipes of MeatLoaf, MeatLoafI (1936 & 2006) and MeatLoafII (1936 & 2006)]  
2) Caramel Custard 

Each recipes has been entered as data tables or data frames and created a ` .tab ` file to a common file format. The file name is composed of the recipe name and year (for example: RecipeName.Year.tab). 

** Data Frame for the Ingredients and NDB_NO reference**
```{r}
NDB_NO<-c(23572,01125,01001,20082, 23562, 11282, 18079, 11935, 01123,11670, 18621, 06559, 10972, 10123,11886, 02055, 06172,  19335, 01077, 02050, 14555)
Ingredient<-c("Beef, ground, 80% lean meat / 20% fat, raw", "Egg, yolk, raw, fresh", "Butter, salted",
              "Wheat flour, white, all-purpose, self-rising, enriched" , "Beef, ground, 90% lean meat
              / 10% fat, raw", "Onions, raw", "Bread, crumbs, dry, grated, plain", "Catsup", "Egg, whole,
              raw, fresh", "Peppers, hot chili, green, raw",  "Nabisco, Nabisco Ritz Crackers", "Soup,
              tomato, canned, prepared with equal volume water, commercial", "Pork, ground, 84% lean 
              / 16% fat, raw", "Pork, cured, bacon, unprepared", "Tomato juice, canned, without salt 
              added",   "Horseradish, prepared", "Soup, stock, chicken, home-prepared", "Sugars, 
              granulated",  "Milk, whole, 3.25% milkfat, with added vitamin D", "Vanilla extract", "Water, bottled, generic")

FOOD_ref<-data.frame(NDB_NO=NDB_NO, Ingredient=Ingredient)
FOOD_ref
```
**Explanation**
Using `grep` function, I choosed the same or the closest matching ingredient from the FOOD_DES data table. The Data frame consists the Ingredients and NDB_NO of that ingredient. The code for grep functions are written below after importing the data from FOOD_DES files in part 2(bottom section of the document)

**Data Frame of Recipe : BeefStroganOff.1997**
```{r}
  NDB_Num = c(13922, 01001, 20082, 06170, 01056, 02046)
  Amount = c(1.5,3.5,1,1,3,1)
  Measure <-c("lb","Tbs","Tbs","cup","Tbs","tsp")
  Ingredient <-c('Beef, tenderloin, roast, separable lean and fat, trimmed to 1/8" fat, choice, cooked, roasted',"Butter, salted","Wheat flour, white, all-purpose, self-rising, enriched","Soup, stock, beef, home-prepared", "Cream, sour, cultured","Mustard, prepared, yellow")

BeefStroganOff_1997<- data.frame(NDB_NO = NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
BeefStroganOff_1997
```
**Data Frame of Recipe : BeefStroganOff.2006**
```{r}
  NDB_Num = c(13922,04582, 01001,11282,11260, 06170, 01056, 02046)
  Amount = c(2,2,3,1,1,1,3,1)
  Measure <-c("lb","Tbs","Tbs","ea","lb","cup","cup","Tbs")
  Ingredient <-c('Beef, tenderloin, roast, separable lean and fat, trimmed to 1/8" fat, choice, cooked, roasted',"Oil, canola","Butter, salted","Onions, raw","Mushrooms, white, raw","Soup, stock, beef, home-prepared", "Cream, sour, cultured","Mustard, prepared, yellow")

BeefStroganOff_2006<- data.frame(NDB_NO = NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
BeefStroganOff_2006
```
**Data Frame of Recipe : MeatLoaf1_1936 **
```{r}
  NDB_Num = c(23572, 01125, 01001, 01001, 20082)
  Amount = c(1.00,1.00,1.00,0.25,2.00)
  Measure <-c("lb","ea","Tbs","cup","Tbs")
  Ingredient <-c("Beef, ground, 80% lean meat / 20% fat, raw","Egg, yolk, raw, fresh","Butter, salted","Butter, salted","Wheat flour, white, all-purpose, self-rising, enriched")

MeatLoaf1_1936<- data.frame(NDB_NO = NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
MeatLoaf1_1936
```
**Data Frame of Recipe : MeatLoaf1_2006**
```{r}
  NDB_Num = c(23572, 23562, 11282, 18079, 11935, 01123)
  Amount = c(12, 12, 1.50, 1, 0.67, 3)
  Measure <-c("oz","oz","cup","cup","cup","lrg")
  Ingredient <-c("Beef, ground, 80% lean meat / 20% fat, raw","Beef, ground, 90% lean meat / 10% fat, raw", "Onions, raw","Bread, crumbs, dry, grated, plain","Catsup", "Egg, whole, raw, fresh")
  
MeatLoaf1_2006<-data.frame( NDB_NO = NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
MeatLoaf1_2006

```

**Data Frame of Recipe : MeatLoaf2_1936**
```{r}
  NDB_Num = c(11670, 11282, 18621, 06559, 23572, 10972, 20082, 10123, 11886, 20082)
  Amount = c(3, 1,24,3.50,1,1,0.25,6,1,2)
  Measure <-c("ea","ea","ea","cup","lb","lb","cup","pce","cup","Tbs")
  Ingredient <-c('Peppers, hot chili, green, raw" ',"Onions, raw","Nabisco, Nabisco Ritz Crackers","tomatoes, whole, canned, with juice","Beef, ground, 80% lean meat / 20% fat, raw","Pork, ground, 84% lean / 16% fat, raw", "Wheat flour, white, all-purpose, self-rising, enriched","Pork, cured, bacon, unprepared","Tomato juice, canned, without salt added","Wheat flour, white, all-purpose, self-rising, enriched")
  
MeatLoaf2_1936<-data.frame(NDB_NO =NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
MeatLoaf2_1936
```

**Data Frame of Recipe : MeatLoaf2_2006**
```{r}
  NDB_Num = c(23562, 02055, 11935, 10123, 18621, 06172, 01123)
  Amount = c(1,1,2,6,24,0.50,1)
  Measure <-c("lb","Tbs","Tbs","pce","ea","cup","lrg")
  Ingredient <-c("Beef, ground, 90% lean meat / 10% fat, raw","Horseradish, prepared", "Catsup", "Pork, cured, bacon, unprepared","Nabisco, Nabisco Ritz Crackers","Soup, stock, chicken, home-prepared","Egg, whole, raw, fresh")
  
MeatLoaf2_2006<-data.frame(NDB_NO =NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
MeatLoaf2_2006
```
**Data Frame of Recipe : CaramelCustard_1936**
```{r}
NDB_Num =c(19335, 01077, 01125, 02050, 14555)
Amount = c(0.50,2,3,0.50,1)
Measure <-c("cup","cup","ea","tsp","Tbs")
Ingredient <-c("Sugars, granulated", "Milk, whole, 3.25% milkfat, with added vitamin D", "Egg, yolk, raw, fresh", "Vanilla extract", "Water, bottled, generic")
  
CaramelCustard_1936<-data.frame(NDB_NO =NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
CaramelCustard_1936
```

**Data Frame of Recipe : CaramelCustard_2006**
```{r}
  NDB_Num <- c(01123, 01077, 19335, 02050, 14555)
  Amount = c(3,2,0.75,0.75,2)
  Measure <-c("lrg","cup","cup","tsp","Tbs")
  Ingredient <-c("Egg, whole, raw, fresh", "Milk, whole, 3.25% milkfat, with added vitamin D", "Sugars, granulated", "extract, vanilla", "Water, bottled, generic")
  
CaramelCustard_2006<-data.frame(NDB_NO =NDB_Num, Amount =Amount, Unit = Measure, Ingredient = Ingredient)
CaramelCustard_2006
```

**Creating ".tab" file for each recipe**
```{r}
write.table(MeatLoaf1_1936, file ="MeatLoaf1.1936.tab",row.names = FALSE)
write.table(MeatLoaf1_2006, file ="MeatLoaf1.2006.tab",row.names = FALSE)
write.table(MeatLoaf2_1936, file ="MeatLoaf2.1936.tab",row.names = FALSE)
write.table(MeatLoaf2_2006, file ="MeatLoaf2.2006.tab",row.names = FALSE)
write.table(CaramelCustard_1936, file = "CaramelCustard.1936.tab",row.names = FALSE)
write.table(CaramelCustard_2006, file = "CaramelCustard.2006.tab",row.names = FALSE)
```

**Getting path**
```{r}
getwd()
```

**Reading data from tab files**
```{r}
MeatLoaf1936_a<- read.table(file = "MeatLoaf1.1936.tab",  sep="\t", header=TRUE)
MeatLoaf1936_a

MeatLoaf2006_a<- read.table(file = "MeatLoaf1.2006.tab",  sep="\t", header=TRUE)
MeatLoaf2006_a

MeatLoaf1936_b<- read.table(file = "MeatLoaf2.1936.tab",  sep="\t", header=TRUE)
MeatLoaf1936_b

MeatLoaf2006_b<- read.table(file = "MeatLoaf2.2006.tab",  sep="\t", header=TRUE)
MeatLoaf2006_b

CaramelCustard1<- read.table(file = "CaramelCustard.1936.tab",  sep="\t", header=TRUE)
CaramelCustard1

CaramelCustard2<- read.table(file = "CaramelCustard.2006.tab",  sep="\t", header=TRUE)
CaramelCustard2

```

**Merging recipes for both years with ingredient to plot graphs for checking if there are any errors**
## Merging MeatLoaf1 (Year 1936 and 2006)
```{r}
Merged_MeatLoaf1 <-merge(MeatLoaf1_1936, MeatLoaf1_2006, by="Ingredient", all=TRUE)
Merged_MeatLoaf1
```
##Merging MeatLoaf2 (Year 1936 and 2006)
```{r}
Merged_MeatLoaf2 <-merge(MeatLoaf2_1936, MeatLoaf2_2006, by="Ingredient", all=TRUE)
Merged_MeatLoaf2
```
## Merging CaramelCustard (Year 1936 and 2006)
```{r}
Merged_CaramelCustard <-merge(CaramelCustard_1936, CaramelCustard_2006, by="Ingredient", all=TRUE)
Merged_CaramelCustard
```

**Plotting graph for Merged_MeatLoaf1 `amount` of 1936 and 2006**
```{r}
plot(Merged_MeatLoaf1$Ingredient,Merged_MeatLoaf1$Amount.x,
main="MeatLoaf1: Amount of ingredients in Year 1936", xlab="Ingredient", ylab="Amount", col="green")
```
```{r}
plot(Merged_MeatLoaf1$Ingredient,Merged_MeatLoaf1$Amount.y,
main="MeatLoaf1: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="blue")
```

**Comparing both above**
```{r}
plot(Merged_MeatLoaf1$Ingredient,Merged_MeatLoaf1$Amount.y,pch=19,
main="MeatLoaf1 : Amount 1936 vs 2006 based on ingredient", xlab="Ingredient", ylab="Amount", col="blue")
points(Merged_MeatLoaf1$Ingredient,Merged_MeatLoaf1$Amount.x,main="MeatLoaf1 : Amount 1936 vs 
       2006 based on ingredient",pch=19, xlab="Ingredient", ylab="Amount", col="blue")
legend("topright", legend=c("Amount 2006"), col=c("black"), lty=1, cex=0.8)
legend("topleft", legend=c("Amount 1936"), col=c( "blue"),pch = 15, cex=0.8)

```
The above plot where symbol for 1936 is blue points and symbol for 2006 is a bar.
The graph verifies that the data taken  is right and there are no errors in the input of data.


**Plotting graph between Merged_MeatLoaf2 amount of 1936 and 2006**
```{r}
plot(Merged_MeatLoaf2$Ingredient,Merged_MeatLoaf2$Amount.x,
main="MeatLoaf2: Amount of ingredients in Year 1936", xlab="Ingredient", ylab="Amount", col="green")
```
```{r}
plot(Merged_MeatLoaf2$Ingredient,Merged_MeatLoaf2$Amount.y,
main="MeatLoaf2: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="blue")
```

**Comparing both above**
```{r}
plot(Merged_MeatLoaf2$Ingredient,Merged_MeatLoaf2$Amount.y,pch=19,
main="MeatLoaf2 : Amount 1936 vs 2006 based on ingredient", xlab="Ingredient", ylab="Amount", col="blue")
points(Merged_MeatLoaf2$Ingredient,Merged_MeatLoaf2$Amount.x,main="MeatLoaf2 : Amount 1936 vs 2006 based on ingredient",pch=19, xlab="Ingredient", ylab="Amount", col="blue")
legend("topright", legend=c("Amount 2006"), col=c("black"), lty=1, cex=0.8)
legend("right", legend=c("Amount 1936"), col=c( "blue"),pch = 15, cex=0.8)

```
The above plot where symbol for 1936 is blue points and symbol for 2006 is a bar verifies that I have taken right data and there are no errors in the data intake.


**Plotting graph between CaramelCustard amount of 1936 and 2006**
```{r}
plot(Merged_CaramelCustard$Ingredient,Merged_CaramelCustard$Amount.x,
main="CaramelCustard: Amount of ingredients in Year 1936", xlab="Ingredient", ylab="Amount", col="red")
```
```{r}
plot(Merged_CaramelCustard$Ingredient,Merged_CaramelCustard$Amount.y,
main="CaramelCustard: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="blue")
```

**Comparing both above**
```{r}
plot(Merged_CaramelCustard$Ingredient,Merged_CaramelCustard$Amount.y,
main="CaramelCustard : Amount 1936 vs 2006 based on ingredient",pch=19, xlab="Ingredient", ylab="Amount", col="blue")
points(Merged_CaramelCustard$Amount.x,Merged_CaramelCustard$Ingredient,
main="CaramelCustard : Amount 1936 vs 2006 based on ingredient",pch=19, xlab="Ingredient", ylab="Amount", col="blue")
legend("right", legend=c("Amount 1936"), col=c("black"), lty=1, cex=0.8)
legend("topright", legend=c("Amount 2006"), col=c( "blue"),pch = 15, cex=0.8)

```
The above plot where symbol for 1936 is blue points and symbol for 2006 is a bar verifies that I have taken right data and there are no errors in the data intake.

The above graphs show verifies that I have taken right values and there is no error in data intake. **

**Contingency Table**
Follwing is the contingency table comparing Measure 1936 and 2006 to look for mismatches in measures between Ingredients:
```{r}
table(MeatLoaf1_1936$Ingredient,MeatLoaf1_1936$Unit)
plot(MeatLoaf1_1936$Ingredient,MeatLoaf1_1936$Amount, las=2,main="MeatLoaf1: Amount of ingredients in Year 1936", xlab="Ingredient", ylab="Amount", col="green")

table(MeatLoaf1_2006$Ingredient,MeatLoaf1_2006$Unit)
plot(MeatLoaf1_2006$Ingredient,MeatLoaf1_2006$Amount, las=2, main="MeatLoaf1: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="red")

table(MeatLoaf2_1936$Ingredient,MeatLoaf2_1936$Unit)
plot(MeatLoaf2_1936$Ingredient,MeatLoaf2_1936$Amount, las=2, main="MeatLoaf2: Amount of ingredients in Year 1936", xlab="Ingredient", ylab="Amount", col="blue")

table(MeatLoaf2_2006$Ingredient,MeatLoaf2_2006$Unit)
plot(MeatLoaf2_2006$Ingredient,MeatLoaf2_2006$Amount, las=2, main="MeatLoaf2: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="blue")

table(CaramelCustard_1936$Ingredient,CaramelCustard_1936$Unit)
plot(CaramelCustard_1936$Ingredient,CaramelCustard_1936$Amount, las=2, main="CaramelCustard: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="blue")

table(CaramelCustard_2006$Ingredient,CaramelCustard_2006$Unit)
plot(CaramelCustard_2006$Ingredient,CaramelCustard_2006$Amount, las=2, main="CaramelCusturd: Amount of ingredients in Year 2006", xlab="Ingredient", ylab="Amount", col="blue")
```
** It is clear from the contingency table that there is no mismatch in measure and Amount with Ingredients.**

**Minimum and Maximum Servings of each recipe**
```{r}
MinServings = c(6,8,10,4,4,5)
MaxServings = c(6,8,10,4,4,5)
servings <- data.frame (MinServings, MaxServings)
write.csv(servings, file = "Servings.csv")
```

Based on PDF  of recipe we find the maximum and minimum servings and created a `.csv` file.

**joining all my recipes together in one file called "MyRecipe" **
Demonstrating that the data can be merged with into a shared table by merging with the example (Beef Stroganoff) recipes.
```{r}
MyRecipe <- rbind(MeatLoaf1_1936, MeatLoaf1_2006, MeatLoaf2_1936, MeatLoaf2_2006, CaramelCustard_1936, CaramelCustard_2006)
MyRecipe
```

**Getting the `RecipeName` and `Year` from file name**
```{r}
ML1_1936<-lapply(strsplit('MeatLoaf1.1936', split=".", fixed = TRUE),trimws)
ML1_1936
ML1_2006<-lapply(strsplit('MeatLoaf1.2006', split=".", fixed = TRUE),trimws)
ML1_2006
ML2_1936<-lapply(strsplit('MeatLoaf2.1936', split=".", fixed = TRUE),trimws)
ML2_1936
ML2_2006<-lapply(strsplit('MeatLoaf2.2006', split=".", fixed = TRUE),trimws)
ML2_2006
CC_1936<-lapply(strsplit('CaramelCusturd.1936', split=".", fixed = TRUE),trimws)
CC_1936
CC_2006<-lapply(strsplit('CaramelCusturd.2006', split=".", fixed = TRUE),trimws)
CC_2006

```
**Creating data frame consists Recipe and Year, extracted from the File name**
```{r}
AddRecipe <-data.frame(c(ML1_1936,ML1_2006,ML2_1936,ML2_2006,CC_1936,CC_2006),stringsAsFactors=FALSE)
rownames(AddRecipe)<- c("Recipe", "Year")
colnames(AddRecipe)<-NULL
AddRecipeName <-(t(AddRecipe))
AddRecipeName

```
**generating csv file**
```{r}
write.csv(AddRecipeName, file = "AddRecipe.csv", row.names = FALSE)
```
**joining Recipe name and year with Servings based on example of Beef Stroganoff**
```{r}
Add_Recipe_servings <- cbind(AddRecipeName, servings)
Add_Recipe_servings
```

**Generating .csv file of Add_Recipe, that consists Recipe name, year, minserving and max serving**
```{r}
write.csv(Add_Recipe_servings, file = "Add_Recipe_servings.csv", row.names = FALSE)
```

**Combining the data of sample file for Beef Stroganoff with my recipe data with servings**
```{r}
# PMC - should be reading commas
Recipe1 <-read.csv(file = "Recipes.csv", header=TRUE)
Recipe2 <- read.csv(file = "Add_Recipe_servings.csv", header=TRUE)
My_Recipe<-rbind(Recipe1, Recipe2)
My_Recipe
```
**Output CSV for combined data**
```{r}
write.csv(My_Recipe, file = "Recipes.csv", row.names = FALSE)
```
**Joining my all recipe name and year together**
```{r}
Recipe_1 = matrix(rep(AddRecipeName[1,], length(MeatLoaf1_1936$Amount)),ncol=2, byrow = T)
Recipe_2 = matrix(rep(AddRecipeName[2,], length(MeatLoaf1_2006$Amount)),ncol=2, byrow = T)
Recipe_3 = matrix(rep(AddRecipeName[3,], length(MeatLoaf2_1936$Amount)),ncol=2, byrow = T)
Recipe_4 = matrix(rep(AddRecipeName[4,], length(MeatLoaf2_2006$Amount)),ncol=2, byrow = T)
Recipe_5 = matrix(rep(AddRecipeName[5,], length(CaramelCustard_1936$Amount)),ncol=2, byrow = T)
Recipe_6 = matrix(rep(AddRecipeName[6,], length(CaramelCustard_2006$Amount)),ncol=2, byrow = T)
MyAllRecipe<-rbind(Recipe_1,Recipe_2,Recipe_3,Recipe_4,Recipe_5,Recipe_6)
colnames(MyAllRecipe)<- c("Recipe", "Year")
MyAllRecipe
```
** joing data frame MyAllRecipe and MyRecipe to get single table having `name`,`Year`,`Amount`,`Unit`,`Ingredient`**
```{r}
Final <- cbind(MyAllRecipe,MyRecipe)
names(Final)[names(Final) == "Unit"] = "Measure"
Final
```
**Creating CSV file of Final Table.**
```{r}
write.csv(Final, file = "Final_Recipe.csv", row.names = FALSE)
```

## ** Codes for merging with example beef recipe BeefStroganOff.1997 and BeefSroganOff.2006**

```{r}
write.table(BeefStroganOff_1997, file ="Beef_StroganOff.1997.tab",row.names = FALSE)
write.table(BeefStroganOff_2006, file ="Beef_StroganOff.2006.tab",row.names = FALSE)
```
```{r}
Beef_1997<- read.table(file = "BeefStroganoff.1997.tab",  sep="\t", header=TRUE)
Beef_1997

Beef_2006<- read.table(file = "BeefStroganoff.2006.tab",  sep="\t", header=TRUE)
Beef_2006
```
**Getting the `RecipeName` and `Year` from file name**
```{r}
BSO_1997<-lapply(strsplit('BeefStroganOff.1997', split=".", fixed = TRUE),trimws)
BSO_1997
BSO_2006<-lapply(strsplit('BeefStroganOff.2006', split=".", fixed = TRUE),trimws)
BSO_2006
```
**Creating data frame consists Recipe and Year, extracted from the File name**
```{r}
Add_beef <-data.frame(c(BSO_1997,BSO_2006),stringsAsFactors=FALSE)
rownames(Add_beef)<- c("Recipe", "Year")
colnames(Add_beef)<-NULL
Add_BSO <-(t(Add_beef))
Add_BSO

```
```{r}
write.csv(Add_BSO, file = "AddBeef.csv", row.names = FALSE)
```
```{r}
Beef_rep <- rbind(BeefStroganOff_1997, BeefStroganOff_2006)
Beef_rep
```
```{r}
Rec_1 = matrix(rep(Add_BSO[1,], length(BeefStroganOff_1997$Amount)),ncol=2, byrow = T)
Rec_2 = matrix(rep(Add_BSO[2,], length(BeefStroganOff_2006$Amount)),ncol=2, byrow = T)
BeefRecipe<-rbind(Rec_1,Rec_2)
colnames(BeefRecipe)<- c("Recipe", "Year")
BeefRecipe

```
** Again Merging "My all recipe" with "BeefStroganOff" example recipe. ** 
#Demonstrates that the data can be merged with into a shared table by merging with the example (Beef Stroganoff) recipes.**
```{r}
Final_Beef <- cbind(BeefRecipe,Beef_rep)
names(Final_Beef)[names(Final_Beef) == "Unit"] = "Measure"
Final_Beef
```
```{r}
MyRecipeAndBeef <- rbind(Final_Beef,Final)
MyRecipeAndBeef
```
**Creating CSV file of after merging My recipe (Final table) and BeefStroganOff (Example recipe)**
```{r}
write.csv(MyRecipeAndBeef, file = "Final_All_Recipes.csv", row.names = FALSE)
```

** Path of my working directory**
```{r}
getwd()
```
# Part 1 - Assign Recipes

First, generate a list of all files in the Recipes folder.
```{r,eval=FALSE}
recipes <- list.files(path = "C:/Users/Divya/Desktop/STAT-700/Mid term project/Recipes", pattern = '*.pdf')
recipes
```

Make this reproducible.
```{r,eval=FALSE}
set.seed(10010)
```

Create two randomly sampled list of recipe files and cycle this list through 80 rows

```{r,eval=FALSE}
recipe.pairs <- matrix(c(sample(recipes),sample(recipes)),ncol=2,nrow=106)
recipe.pairs
```

Are there any pairs across rows?

```{r,eval=FALSE}
sum(recipe.pairs[,1]==recipe.pairs[,2])
```

Write to a table

```{r,eval=FALSE}
write.csv(recipe.pairs, file="Recipe.Table.csv",row.names = FALSE)
```

Read the table in and print:

```{r,eval=FALSE}
Recipe.Table <- read.csv("Recipe.Table.csv",header=FALSE,skip=1)
print(Recipe.Table)
```
# Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.
```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)
```

** FOOD_DES from the USDA Nutrition Database.**
```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")

names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```

**Searching Ingredients in FOOD_DES.dat file**
** Ingregients search for Meat Loaf, BeefStroganOff and Caramel Custard**
```{r}
#grep(pattern="[tT]enderloin,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 13922
#grep(pattern="[bB]eef, ground", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 23572
#grep(pattern="[sS]our,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 01056
#grep(pattern="[mM]sustard,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 02046
#grep(pattern="[oO]il,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 04582
#grep(pattern="[mM]ushrooms,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 11282
#grep(pattern="^[eE]gg, yolk,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) # NDB_No 01125
#grep(pattern="^[bB]utter, salted", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 01001
#grep(pattern="*[fF]lour,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 20082
#grep(pattern="^[oO]nion", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 11282
#grep(pattern="^[bB]read, crumbs", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 18079
#grep(pattern="^[cC]atsup", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 11935
#grep(pattern="^[eE]gg", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 01123
#grep(pattern="*[pP]epper", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 11670
#grep(pattern="[cC]rackers", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 18621
#grep(pattern="[tT]omato,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 06559
#grep(pattern="^[pP]ork, ground,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 10972
#grep(pattern="[bB]acon,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 10123
#grep(pattern="[jJ]uice,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 11886
#grep(pattern="[hH]orseradish,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 02055
#grep(pattern="[sS]tock,", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 06172
#grep(pattern="[sS]ugars, granulated", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 19335
#grep(pattern="[mM]ilk, whole, 3.25%", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 01077
#grep(pattern=" [eE]xtract", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 02050
#grep(pattern="[gG]eneric", food_des.dat$Long_Desc, ignore.case = TRUE, value=TRUE) #NDB_No 14555
```

## Uisng the grep function I matched the recipe ingredient with FOOD_DES.dat and select the same if it is there and if not the i choosed the best possible match available in ingredients. Through this I found the NBD_NO from weight table and manually entered the the `NDB_NO`, `Amount`, `Unit`, `Ingredients` and created a data frame to perfor the required instruction.


**To Check John Becker's listed ingredients (in the PDF) are consistent with the USDA database**
```{r}
subset(weight.dat, weight.dat$Msre_Desc %in% MyRecipe$Unit)
```

```{r}
GrWeight_Sum <- sum((weight.dat$Gm_Wgt[weight.dat$Msre_Desc %in% MyRecipe$Unit])* MyRecipe$unit, is.na=TRUE)
GrWeight_Sum
```
** NUT_DATA from the USDA Nutrition Database.**
```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)
```

## Goal of the Mid Term:

1) Create a data table from vectors of the 2 recipes that you were assigned. Columns are to include NDB_No ,  Amount, Unit, Ingredient.

2) Export to tab file and re-import, adding columns for recipe name and year. 

3) Import beef stroganoff recipes and merge these tables with the tables from your 2 recipes. This shows that other students can create your data tables and merge with theirs for the Final Project.

4) to check for errors plotting Measure.1936 vs Measure.2006 and see quickly where the measures are not consistent.

5) Reading the Recipes.csv, append rows for my recipes, then write the table back to Recipes.csv.

6) I have also appended my recipes to BeefStroganOff_1997.tab and BeefStroganOff_2006.tab
