---
title: "Joy Of Cooking Midterm Project"
author: "Set 46"
date: "7/21/2018"
output:
  word_document: default
  html_document: default
---

# Check Nutrient Data

Run this to make sure your system can read the required nutrient database tables.

```{r}
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)
summary(weight.dat)
```

```{r}
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```

```{r}
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)
summary(nut_data.dat)
```

You are required to enter ingredients using the text listed in FOOD_DES.txt. In some cases, the ingredient in the PDF files are identical to the ingredients lists in FOOD_DES.txt, aside from case. The Beef Stroganoff recipes list "butter, salted" , in FOOD_DES, this is listed as "Butter, salted". In other cases, there are not exact matches. You will need to perform partial matching in FOOD_DES and pick the best match. For example, using grep('(D|d)ijon' ... , I don't find any matches, so I use grep('(M|m)ustard', to find the best match. I would enter "Mustard, prepared, yellow" into the Beef Stroganoff recipe, instead of the listed ingredient "mustard, dijon".

Print the WEIGHT.dat table featuring the NDB_No and Msre_Desc variables.

Print the FOOD_DES.dat table featuring the NDB_No and Long_Desc variables.

```{r}
data.frame(weight.dat$NDB_No,weight.dat$Msre_Desc)
data.frame(food_des.dat$NDB_No,food_des.dat$Long_Desc)
```

Use grep to find the ingredients lists in FOOD_DES.txt to replace the ingredients in the created data tables.

Use grep to find the NDB_No in FOOD_DES.txt that corresponds with the found ingredient for the created data tables.

Use grep to find the Msre_Desc that corresponds with the found NDB_No

```{r}
food_des.dat$Long_Desc[(grep('(S|s)tock', food_des.dat$Long_Desc))]
#stock, beef, prepared from recipe = Soup, stock, beef, home-prepare
food_des.dat$NDB_No[(grep('Soup, stock, beef, home-prepared', food_des.dat$Long_Desc))]
#NDB_No = 6170

food_des.dat$Long_Desc[(grep('(M|m)ustard', food_des.dat$Long_Desc))]
#mustard, dijon = Mustard, prepared, yellow
food_des.dat$NDB_No[(grep('Mustard, prepared, yellow', food_des.dat$Long_Desc))]
#NDB_No = 2046

food_des.dat$Long_Desc[(grep('(B|b)utter, (S|s)alted', food_des.dat$Long_Desc))]
#butter, salted = Butter, salted
food_des.dat$NDB_No[(grep('Butter, salted', food_des.dat$Long_Desc))]
#NDB_No = 1001

food_des.dat$Long_Desc[(grep('(M|m)ushrooms', food_des.dat$Long_Desc))]
#mushrooms, fresh = Mushrooms, canned, drained solids
food_des.dat$NDB_No[(grep('Mushrooms, canned, drained solids', food_des.dat$Long_Desc))]
#NDB_No = 11264

food_des.dat$Long_Desc[(grep('(S|s)our (C|c)ream', food_des.dat$Long_Desc))]
#sour cream, cultured = Sour cream, imitation, cultured
food_des.dat$NDB_No[(grep('Sour cream, imitation, cultured', food_des.dat$Long_Desc))]
#NDB_No = 1074

food_des.dat$Long_Desc[(grep('(O|o)il, (C|c)anola', food_des.dat$Long_Desc))]
#oil, canola = Oil, canola
food_des.dat$NDB_No[(grep('Oil, canola', food_des.dat$Long_Desc))]
#NDB_No = 4582

food_des.dat$Long_Desc[(grep('(A|a)pple', food_des.dat$Long_Desc))]
#apple, fresh, large, 3 1/4 = Apples, raw, with skin (Includes foods for USDA's Food Distribution Program)
food_des.dat$NDB_No[(grep('Apples, raw, with skin', food_des.dat$Long_Desc))]
#NDB_No = 9003

#Note: 'Apples, raw, with skin (Includes foods for USDA's Food Distribution Program)' is the proper name in FOOD_DES.dat table, but function used to find NDB_No results in integer(0). Find NDB_No using only 'Apples, raw, with skin', but use proper name when replacing Ingredient names.

food_des.dat$Long_Desc[(grep('(V|v)inegar', food_des.dat$Long_Desc))]
#vinegar, distilled = Vinegar, distilled
food_des.dat$NDB_No[(grep('Vinegar, distilled', food_des.dat$Long_Desc))]
#NDB_No = 2053

#vinegar, cider = Vinegar, cider
food_des.dat$NDB_No[(grep('Vinegar, cider', food_des.dat$Long_Desc))]
#NDB_No = 2048

food_des.dat$Long_Desc[(grep('(S|s)ugar', food_des.dat$Long_Desc))]
#sugar, white, granulated, granulated = Sugars, granulated
food_des.dat$NDB_No[(grep('Sugars, granulated', food_des.dat$Long_Desc))]
#NDB_No = 19335

#sugar, brown, packed = Sugars, brown
food_des.dat$NDB_No[(grep('Sugars, brown', food_des.dat$Long_Desc))]
#NDB_No = 19334

food_des.dat$Long_Desc[(grep('(B|b)acon', food_des.dat$Long_Desc))]
#bacon, raw, cured = Bacon, pre-sliced, reduced/low sodium, unprepared
food_des.dat$NDB_No[(grep('Bacon, pre-sliced, reduced/low sodium, unprepared', food_des.dat$Long_Desc))]
#NDB_No = 10994

food_des.dat$Long_Desc[(grep('(C|c)abbage', food_des.dat$Long_Desc))]
#cabbage, fresh, shredded = Cabbage, raw
food_des.dat$NDB_No[(grep('Cabbage, raw', food_des.dat$Long_Desc))]
#NDB_No = 11109

food_des.dat$Long_Desc[(grep('(W|W)ater', food_des.dat$Long_Desc))]
#water, generic = Water, bottled, generic
food_des.dat$NDB_No[(grep('Water, bottled, generic', food_des.dat$Long_Desc))]
#NDB_No = 14555

food_des.dat$Long_Desc[(grep('(G|g)elatin', food_des.dat$Long_Desc))]
#gelatin, dry, unsweetened = Gelatins, dry powder, unsweetened
food_des.dat$NDB_No[(grep('Gelatins, dry powder, unsweetened', food_des.dat$Long_Desc))]
#NDB_No = 19177

#gelatin mix, lemon, Jell-O = Gelatin desserts, dry mix, reduced calorie, with aspartame, added phosphorus, potassium, sodium, vitamin C
food_des.dat$NDB_No[(grep('Gelatin desserts, dry mix, reduced calorie, with aspartame, added phosphorus, potassium, sodium, vitamin C', food_des.dat$Long_Desc))]
#NDB_No = 19703

food_des.dat$Long_Desc[(grep('(J|j)uice', food_des.dat$Long_Desc))]
#juice, lemon, fresh = Lemon juice, raw
food_des.dat$NDB_No[(grep('Lemon juice, raw', food_des.dat$Long_Desc))]
#NDB_No = 9152

food_des.dat$Long_Desc[(grep('(H|h)orseradish', food_des.dat$Long_Desc))]
#horseradish, grated, in vinegar = Horseradish, prepared
food_des.dat$NDB_No[(grep('Horseradish, prepared', food_des.dat$Long_Desc))]
#NDB_No = 2055

food_des.dat$Long_Desc[(grep('(C|c)ream, (W|w)hip', food_des.dat$Long_Desc))]
#cream, whipping, heavy = Cream, whipped, cream topping, pressurized
food_des.dat$NDB_No[(grep('Cream, whipped, cream topping, pressurized', food_des.dat$Long_Desc))]
#NDB_No = 1054

food_des.dat$Long_Desc[(grep('(O|o)nion', food_des.dat$Long_Desc))]
#onion, yellow, fresh, medium, 2 1/2" = Onions, yellow, sauteed
food_des.dat$NDB_No[(grep('Onions, yellow, sauteed', food_des.dat$Long_Desc))]
#NDB_No = 11286

#onion, yellow, fresh, chopped = Onions, yellow, sauteed
food_des.dat$NDB_No[(grep('Onions, yellow, sauteed', food_des.dat$Long_Desc))]
#NDB_No = 11286

food_des.dat$Long_Desc[(grep('(F|f)lour', food_des.dat$Long_Desc))]
#flour, all purpose, self rising, enriched = Wheat flour, white, all-purpose, self-rising, enriched
food_des.dat$NDB_No[(grep('Wheat flour, white, all-purpose, self-rising, enriched', food_des.dat$Long_Desc))]
#NDB_No = 20082

food_des.dat$Long_Desc[(grep('(A|a)vocado', food_des.dat$Long_Desc))]
#avocado, California, fresh = Avocados, raw, California
food_des.dat$NDB_No[(grep('Avocados, raw, California', food_des.dat$Long_Desc))]
#NDB_No = 9038

food_des.dat$Long_Desc[(grep('(C|c)elery', food_des.dat$Long_Desc))]
#celery stalk, fresh, medium, 7.5"to 8" = Celery, raw
food_des.dat$NDB_No[(grep('Celery, raw', food_des.dat$Long_Desc))]
#NDB_No = 11143

food_des.dat$Long_Desc[(grep('(B|b)eef, (T|t)enderloin', food_des.dat$Long_Desc))]
#beef roast, tenderloin, raw, 0" trim = Beef, tenderloin, separable lean and fat, trimmed to 1/8" fat, prime, raw
food_des.dat$NDB_No[(grep('Beef, tenderloin, separable lean and fat, trimmed to 1/8" fat, prime, raw', food_des.dat$Long_Desc))]
#NDB_No = 13926


food_des.dat$Long_Desc[(grep('(T|t)omato', food_des.dat$Long_Desc))]
#tomatoes, whole, canned, with juice = Tomatoes, red, ripe, canned, packed in tomato juice, no salt added
food_des.dat$NDB_No[(grep('Tomatoes, red, ripe, canned, packed in tomato juice, no salt added', food_des.dat$Long_Desc))]
#NDB_No = 11885

#juice, tomato, with salt = Tomato juice, canned, with salt added
food_des.dat$NDB_No[(grep('Tomato juice, canned, with salt added', food_des.dat$Long_Desc))]
#NDB_No = 11540

#juice, tomato, canned, unsalted = Tomato juice, canned, without salt added
food_des.dat$NDB_No[(grep('Tomato juice, canned, without salt added', food_des.dat$Long_Desc))]
#NDB_No = 11886

#tomato puree, canned = Tomato products, canned, puree, without salt added
food_des.dat$NDB_No[(grep('Tomato products, canned, puree, without salt added', food_des.dat$Long_Desc))]
#NDB_No = 11547

#soup, tomato, canned, condensed = Soup, tomato, canned, condensed, reduced sodium
food_des.dat$NDB_No[(grep('Soup, tomato, canned, condensed, reduced sodium', food_des.dat$Long_Desc))]
#NDB_No = 6956

```

Replace the Ingredient names listed for the Beef Stroganoff, Hot Slaw and Tomato Aspic data to the corresponding Ingredient names above from the FOOD_DES.dat table.

Replace Measure / Unit names 'TBS' and 'ea' listed for the Beef Stroganoff, Hot Slaw and Tomato Aspic data to the corresponding Ingredient names above from the WEIGHT.dat table:
- 'TBS' = 'tbsp' (both represent the unit of measure Tablespoon)
- 'ea' = 'peices' (ea represents unit of measure eaches or peices)


# Part 1 - Creating Data Tables

Create data tables for the assigned recipes, with columns Amount, Unit, Ingredient and NDB_No.

Write the code to create the data tables for Beef Stroganoff, 1997 and Beef Stroganoff, 2006.

```{r}
Beef_Stroganoff_1997.dat <- data.frame(
  Amount = c(1.5, 3.5, 1.0, 1.0, 3.0, 1.0),
  Measure = c('lb', 'tbsp', 'tbsp', 'cup', 'tbsp', 'tsp'),
  Ingredient = c('Beef, tenderloin, separable lean and fat, trimmed to 1/8" fat, prime, raw','Butter, salted','Wheat flour, white, all-purpose, self-rising, enriched','Soup, stock, beef, home-prepared','Sour cream, imitation, cultured','Mustard, prepared, yellow'),
  NDB_No = c(13926, 1001, 20082, 6170, 1074, 2046)
)
Beef_Stroganoff_1997.dat

Beef_Stroganoff_2006.dat <- data.frame(
  Amount = c(2.0, 2.0, 3.0, 1.0, 1.0, 1.0, 3.0, 1.0),
  Measure = c('lb', 'tbsp', 'tbsp', 'pieces', 'lb', 'cup', 'cup', 'tbsp'),
  Ingredient = c('Beef, tenderloin, separable lean and fat, trimmed to 1/8" fat, prime, raw','Oil, canola','Butter, salted','Onions, yellow, sauteed','Mushrooms, canned, drained solids','Soup, stock, beef, home-prepared','Sour cream, imitation, cultured','Mustard, prepared, yellow'),
  NDB_No = c(13926, 4582, 1001, 11286, 11264, 6170, 1074, 2046)
)
Beef_Stroganoff_2006.dat
```

Write the code to create the data tables for Hot Slaw, 1936 and Hot Slaw, 2006.

```{r}
Hot_Slaw_1936.dat <- data.frame(
  Amount = c(3.00, 3.00, 1.00, 2.00, 1.00),
  Unit = c('cup', 'tbsp', 'tbsp', 'tbsp', 'pieces'),
  Ingredient = c('Cabbage, raw', 'Vinegar, distilled', 'Sugars, granulated', 'Butter, salted', "Apples, raw, with skin (Includes foods for USDA's Food Distribution Program)"),
  NDB_No = c(11109, 2053, 19335, 1001, 9003)
)
Hot_Slaw_1936.dat

Hot_Slaw_2006.dat <- data.frame(
  Amount = c(0.50, 3.00, 1.00, 3.00),
  Unit = c('lb', 'tbsp', 'tbsp', 'cup'),
  Ingredient = c('Bacon, pre-sliced, reduced/low sodium, unprepared', 'Vinegar, cider', 'Sugars, brown', 'Cabbage, raw'),
  NDB_No = c(10994, 2048, 19334, 11109)
)
Hot_Slaw_2006.dat
```

Write the code to create the data tables for Tomato Aspic i, 1936 and Tomato Aspic i, 2006.

Write the code to create the data tables for Tomato Aspic ii, 1936 and Tomato Aspic ii, 2006.

```{r}
Tomato_Aspic_i_1936.dat <- data.frame(
  Amount = c(3.50, 1.50, 3.00, 2.00, 4.00, 2.00, 0.50, 0.50, 3.00, 2.00),
  Unit = c('cup', 'tsp', 'tbsp', 'tbsp', 'pieces', 'tbsp', 'cup', 'cup', 'tbsp', 'tbsp'),
  Ingredient = c('Tomatoes, red, ripe, canned, packed in tomato juice, no salt added', 
                 'Sugars, granulated', 
                 'Onions, yellow, sauteed', 
                 'Lemon juice, raw', 
                 'Celery, raw', 
                 'Gelatins, dry powder, unsweetened', 
                 'Water, bottled, generic', 
                 'Cream, whipped, cream topping, pressurized', 
                 'Lemon juice, raw', 
                 'Horseradish, prepared'),
  NDB_No = c(11885, 19335, 11286, 9152, 11143, 19177, 14555, 1054, 9152, 2055)
)
  
Tomato_Aspic_i_1936.dat

Tomato_Aspic_ii_1936.dat <- data.frame(
  Amount = c(2.00, 4.00),
  Unit = c('tbsp', 'cup'),
  Ingredient = c('Gelatins, dry powder, unsweetened', 
                 'Tomato juice, canned, with salt added'),
  NDB_No = c(19177, 11540)
)
Tomato_Aspic_ii_1936.dat

Tomato_Aspic_i_2006.dat <- data.frame(
  Amount = c(4.00, 0.50, 0.50, 2.00, 2.00, 2.00, 4.50, 0.50, 1.00),
  Unit = c('cup', 'cup', 'cup', 'pieces', 'tbsp', 'tsp', 'tsp', 'cup', 'pieces'),
  Ingredient = c('Tomato juice, canned, without salt added', 
                 'Tomato products, canned, puree, without salt added', 
                 'Onions, yellow, sauteed', 
                 'Celery, raw', 
                 'Lemon juice, raw', 
                 'Sugars, granulated', 
                 'Gelatins, dry powder, unsweetened', 
                 'Water, bottled, generic', 
                 'Avocados, raw, California'),
  NDB_No = c(11886, 11547, 11286, 11143, 9152, 19335, 19177, 14555, 9038)
)
Tomato_Aspic_i_2006.dat

Tomato_Aspic_ii_2006.dat <- data.frame(
  Amount = c(2.25, 10.75, 2.00, 3.00),
  Unit = c('tsp', 'oz', 'cup', 'oz'),
  Ingredient = c('Gelatins, dry powder, unsweetened', 
                 'Soup, tomato, canned, condensed, reduced sodium', 
                 'Tomato juice, canned, without salt added', 
                 'Gelatin desserts, dry mix, reduced calorie, with aspartame, added phosphorus, potassium, sodium, vitamin C'),
  NDB_No = c(19177, 6956, 11886, 19703)
)
Tomato_Aspic_ii_2006.dat
```


# Part 2 - Creating Tab Delimited Files

Write each individual recipe to a tab delimited file. The name of the tab file should be comprised of the recipe file name and the year of the recipe (1936 or 2006, in most cases).

Create tab delimited files for:
- Beef Stroganoff, 1997
- Beef Stroganoff, 2006
- Hot Slaw, 1936
- Hot Slaw, 2006
- Tomato Aspic i, 1936
- Tomato Aspic i, 2006
- Tomato Aspic ii, 1936
- Tomato Aspic ii, 2006

```{r}
write.table(Beef_Stroganoff_1997.dat, file = "BeefStroganoff.1997.tab",  sep = "\t")

write.table(Beef_Stroganoff_2006.dat, file = "BeefStroganoff.2006.tab", sep = "\t")


write.table(Hot_Slaw_1936.dat, file = "HotSlaw.1936.tab", sep = "\t")

write.table(Hot_Slaw_2006.dat, file = "HotSlaw.2006.tab", sep = "\t")


write.table(Tomato_Aspic_i_1936.dat, file = "TomatoAspici.1936.tab", sep = "\t")

write.table(Tomato_Aspic_i_2006.dat, file = "TomatoAspici.2006.tab", sep = "\t")

write.table(Tomato_Aspic_ii_1936.dat, file = "TomatoAspicii.1936.tab", sep = "\t")

write.table(Tomato_Aspic_ii_2006.dat, file = "TomatoAspicii.2006.tab", sep = "\t")
```


# Part 3 - Merging Tables

Read in each tab file and merge the tables.

Read in the tab demilimated files while adding Recipe and Year columns for:
- Beef Stroganoff, 1997
- Beef Stroganoff, 2006
- Hot Slaw, 1936
- Hot Slaw, 2006
- Tomato Aspic i, 1936
- Tomato Aspic i, 2006
- Tomato Aspic ii, 1936
- Tomato Aspic ii, 2006

```{r}
BeefStroganoff.1997.Table <- read.table("BeefStroganoff.1997.tab", header=TRUE)
BeefStroganoff.1997.Table["Recipe"] <- "Beef Stroganoff"
BeefStroganoff.1997.Table["Year"] <- 1997
BeefStroganoff.1997.Table

BeefStroganoff.2006.Table <- read.table("BeefStroganoff.2006.tab", header=TRUE)
BeefStroganoff.2006.Table["Recipe"] <- "Beef Stroganoff"
BeefStroganoff.2006.Table["Year"] <- 2006
BeefStroganoff.2006.Table


HotSlaw.1936.Table <- read.table("HotSlaw.1936.tab", header=TRUE)
HotSlaw.1936.Table["Recipe"] <- "Hot Slaw"
HotSlaw.1936.Table["Year"] <- 1936
HotSlaw.1936.Table

HotSlaw.2006.Table <- read.table("HotSlaw.2006.tab", header=TRUE)
HotSlaw.2006.Table["Recipe"] <- "Hot Slaw"
HotSlaw.2006.Table["Year"] <- 2006
HotSlaw.2006.Table


TomatoAspici.1936.Table <- read.table("TomatoAspici.1936.tab", header=TRUE)
TomatoAspici.1936.Table["Recipe"] <- "Tomato Aspic i"
TomatoAspici.1936.Table["Year"] <- 1936
TomatoAspici.1936.Table

TomatoAspici.2006.Table <- read.table("TomatoAspici.2006.tab", header=TRUE)
TomatoAspici.2006.Table["Recipe"] <- "Tomato Aspic i"
TomatoAspici.2006.Table["Year"] <- 2006
TomatoAspici.2006.Table

TomatoAspicii.1936.Table <- read.table("TomatoAspicii.1936.tab", header=TRUE)
TomatoAspicii.1936.Table["Recipe"] <- "Tomato Aspic ii"
TomatoAspicii.1936.Table["Year"] <- 1936
TomatoAspicii.1936.Table

TomatoAspicii.2006.Table <- read.table("TomatoAspicii.2006.tab", header=TRUE)
TomatoAspicii.2006.Table["Recipe"] <- "Tomato Aspic ii"
TomatoAspicii.2006.Table["Year"] <- 2006
TomatoAspicii.2006.Table
```

Merge BeefStroganoff.1997.Table and BeefStroganoff.2006.Table to create BeefStroganoff_Merged.
Merge HotSlaw.1936.Table and HotSlaw.2006.Table to create HotSlaw_Merged.
Merge TomatoAspici.1936.Table and TomatoAspici.2006.Table to create TomatoAspici_Merged.
Merge TomatoAspicii.1936.Table and TomatoAspicii.2006.Table to create TomatoAspicii_Merged.
Merge TomatoAspici_Merged and TomatoAspicii_.2006.Table_Merged to create TomatoAspic_Merged.
Merge HotSlaw_Merged and TomatoAspic_Merged to create Merged_Table_1.
Merge Merged_Table_1 and BeefStroganoff_Merged to create Merged_Table_2.

```{r}
BeefStroganoff_Merged <- merge(BeefStroganoff.1997.Table,BeefStroganoff.2006.Table, by=c("Recipe","Year","Amount","Measure","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
BeefStroganoff_Merged

HotSlaw_Merged <- merge(HotSlaw.1936.Table,HotSlaw.2006.Table, by=c("Recipe","Year","Amount","Unit","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
HotSlaw_Merged

TomatoAspici_Merged <- merge(TomatoAspici.1936.Table,TomatoAspici.2006.Table, by=c("Recipe","Year","Amount","Unit","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
TomatoAspici_Merged

TomatoAspicii_Merged <- merge(TomatoAspicii.1936.Table,TomatoAspicii.2006.Table, by=c("Recipe","Year","Amount","Unit","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
TomatoAspicii_Merged

TomatoAspic_Merged <- merge(TomatoAspici_Merged,TomatoAspicii_Merged, c("Recipe","Year","Amount","Unit","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
TomatoAspic_Merged

HotSlaw_TomatoAspic_Merged <- merge(HotSlaw_Merged,TomatoAspic_Merged, c("Recipe","Year","Amount","Unit","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
HotSlaw_TomatoAspic_Merged

BeefStroganoff_HotSlaw_TomatoAspic_Merged <- merge(HotSlaw_TomatoAspic_Merged,BeefStroganoff_Merged, by.x = c("Recipe","Year","Amount","Unit","Ingredient","NDB_No"), by.y = c("Recipe","Year","Amount","Measure","Ingredient","NDB_No"), all = TRUE, sort=TRUE)
BeefStroganoff_HotSlaw_TomatoAspic_Merged
```


# Part 4 - Checking for Errors

Proof the data to check for errors.

Create tables to find the errors between years in the created merged tables for Measure / Units and Amounts.
Sort by Ingredient.

```{r}
#BeefStroganoff_Merged
BeefStroganoff_Merged_1 <- merge(BeefStroganoff.1997.Table,BeefStroganoff.2006.Table, by=c("Ingredient"), all = TRUE, sort=TRUE)

table(BeefStroganoff_Merged_1$Measure.x,BeefStroganoff_Merged_1$Measure.y,BeefStroganoff_Merged_1$Ingredient)

table(BeefStroganoff_Merged_1$Amount.x,BeefStroganoff_Merged_1$Amount.y,BeefStroganoff_Merged_1$Ingredient)

#HotSlaw_Merged
HotSlaw_Merged_1 <- merge(HotSlaw.1936.Table,HotSlaw.2006.Table, by=c("Ingredient"), all = TRUE, sort=TRUE)

table(HotSlaw_Merged_1$Amount.x,HotSlaw_Merged_1$Amount.y,HotSlaw_Merged_1$Ingredient)

table(HotSlaw_Merged_1$Amount.x,HotSlaw_Merged_1$Amount.y,HotSlaw_Merged_1$Ingredient)

#TomatoAspici_Merged
TomatoAspici_Merged_1 <- merge(TomatoAspici.1936.Table,TomatoAspici.2006.Table, by=c("Ingredient"), all = TRUE, sort=TRUE)

table(TomatoAspici_Merged_1$Amount.x,TomatoAspici_Merged_1$Amount.y,TomatoAspici_Merged_1$Ingredient)

table(TomatoAspici_Merged_1$Amount.x,TomatoAspici_Merged_1$Amount.y,TomatoAspici_Merged_1$Ingredient)

#TomatoAspicii_Merged
TomatoAspicii_Merged_1 <- merge(TomatoAspicii.1936.Table,TomatoAspicii.2006.Table, by=c("Ingredient"), all = TRUE, sort=TRUE)

table(TomatoAspicii_Merged_1$Amount.x,TomatoAspicii_Merged_1$Amount.y,TomatoAspicii_Merged_1$Ingredient)

table(TomatoAspicii_Merged_1$Amount.x,TomatoAspicii_Merged_1$Amount.y,TomatoAspicii_Merged_1$Ingredient)
```

Noted Errors:

Beef Stroganoff:
- The ingredient 'Mustard, prepared, yellow' changes from unit of tsp (teaspoon) in 1997 to tbsp (tablespoon) in 2006.
- The ingredient 'Sour cream, imitation, cultured' changes from unit of tbsp (tablespoon) in 1997 to cup in 2006.
- The ingredient 'Wheat flour, white, all-purpose, self-rising, enriched' changes from tbsp (tablespoon) to no units in 2006.
- The ingredients 'Mushrooms, canned, drained solids', 'Oil, canola' and 'Onions, yellow, sauteed' have no units in 1997.

- The ingredient 'Beef, tenderloin, separable lean and fat, trimmed to 1/8" fat, prime, raw' changes from an amount of 1.5 in 1997 to an amount of 2 in 2006.
- The ingredient 'Butter, salted' changes from an amount of 3.5 in 1997 to an amount of 3 in 2006.

Hot Slaw:
- The ingredients 'Bacon, pre-sliced, reduced/low sodium, unprepared', 'Sugars, brown' and 'Vinegar, cider' have no units in 1936.
- The ingredients 'Apples, raw, with skin (Includes foods for USDA's Food Distribution Program)', 'Butter, salted', 'Sugars, granulated' and 'Vinegar, distilled ' have no units in 2006.

Tomato Aspic i:
- The ingredients 'Avocados, raw, California', 'Tomato juice, canned, without salt added' and 'Tomato products, canned, puree, without salt added' have no units in 1936.
- The ingredients 'Cream, whipped, cream topping, pressurized', 'Horseradish, prepared' and 'Tomatoes, red, ripe, canned, packed in tomato juice, no salt added' have no units in 2006.

- The ingredient 'Celery, raw' changes from an amount of 4 in 1936 to an amount of 2 in 2006.
- The ingredient 'Gelatins, dry powder, unsweetened' changes from an amount of 2 in 1936 to an amount of 4.5 in 2006.
- One of the ingredients, 'Lemon juice, raw', appears twice in 1936 and once in 2006. One changes from an amount of 3 in 1936 to an amount of 0 in 2006.
- The ingredient 'Onions, yellow, sauteed' changes from an amount of 3 in 1936 to an amount of 0.5 in 2006.
- The ingredient 'Sugars, granulated' changes from an amount of 1.5 in 1936 to an amount of 2 in 2006.
 
Tomato Aspic ii:
- The ingredient 'Gelatins, dry powder, unsweetened' changes from unit of tbsp (tablespoon) in 1997 to tsp (teaspoon) in 2006.
- The ingredients 'Gelatin desserts, dry mix, reduced calorie, with aspartame, added phosphorus, potassium, sodium, vitamin C', 'Tomato juice, canned, without salt added' and 'Soup, tomato, canned, condensed, reduced sodium' have no units in 1936.
- The ingredient 'Tomato juice, canned, with salt added' has no unit in 2006.

- The ingredient 'Gelatins, dry powder, unsweetened' changes from an amount of 21.5 in 1936 to an amount of 2.25 in 2006.

 
# Part 5 - Create Recipe Table

Read the Recipes.csv, append rows for your recipes, then write the table back to Recipes.csv. We would be able to determine the file names for your recipes by concatentating Recipe and Year, removing white space if necessary.

Read in the csv file Recipes.csv to Recipes.

Merge BeefStroganoff_HotSlaw_TomatoAspic_Merged and Recipes to create Recipe_Table.

Add Gram and Calories columns to Recipe_Table.

```{r}
Recipes <- read.csv("Recipes.csv")
Recipes

Recipe_Table <- merge(BeefStroganoff_HotSlaw_TomatoAspic_Merged,Recipes, by=c("Recipe", "Year"), all = TRUE, sort=TRUE)
Recipe_Table

Recipe_Table['Grams'] <- NA
Recipe_Table['Calories'] <- NA
Recipe_Table
```

Create csv file from Recipe_Table back into "Recipes.csv".

```{r}
#PMC - this doesn't match the original file column format
#write.csv(Recipe_Table, file = "Recipes.csv", row.names=FALSE)
```

