---
title: "STAT 700 Midterm Project"
author: "Set 47"
date: "7/20/2018"
output:
  pdf_document: default
  html_document: default
---

The Midtrem Project is used to validate the calories per recipe data provided by John Becker. The Chocolate Bavarian Cream and the sugar Drop Cookies are two recipes that I need to validate.

# Step 1: Create data tables for assigned recipes

```{r}
# Create two data frame (1936 and 2006) for chocolate Bavarian Cream 
Chocolate_Bavarian_Cream_36 <- data.frame(Amount = c(1, 0.5, 2, 0.33, 1.5, 1, 2), Measure=c("Tbs", "cup", "cup", "cup", "tsp", "cup", "oz"),
                             Ingredient=c("gelatin, dry, unsweetened","water, generic","milk, whole, 3.25%","sugar, white, granulated, granulated","extract, vanilla","cream, whipping, heavy","chocolate, dark, 45-59% cacao solids"))

Chocolate_Bavarian_Cream_06 <- data.frame(Amount = c(3, 1.5, 2.25, 4, 0.25, 1, 1), Measure=c("Tbs", "cup", "tsp", "oz", "cup", "tsp", "cup"),
                             Ingredient=c("milk, whole, 3.25%","milk, whole, 3.25%","gelatin, dry, unsweetened","chocolate, dark, 45-59% cacao solids","sugar, white, granulated, granulated","extract, vanilla","cream, whipping, heavy"))

# Create two data frame (1936 and 2006) for sugar drop cookies 
Drop_Sugar_Cookies_36 <- data.frame(Amount = c(1, 0.5, 1, 1, 1.67, 0.33, 40.00, 0.25), Measure=c("cup", "cup", "lrg", "tsp", "cup", "cup", "ea", "cup"),
                             Ingredient=c("sugar, white, granulated, granulated","butter, salted","egg, raw","extract, vanilla","flour, bread, white, enriched","sour cream, cultured", "pecans, halves", "sugar, white, granulated, granulated"))

Drop_Sugar_Cookies_06 <- data.frame(Amount = c(2.5, 1, 0.75, 2, 1, 0.25), Measure=c("cup", "cup", "cup", "lrg", "tsp", "cup"),
                             Ingredient=c("flour, all purpose, self rising, enriched", "sugar, white, granulated, granulated","oil, canola","egg, raw","extract, vanilla", "sugar, white, granulated, granulated"))
```

Comments: These four dateframes are created manually.

# Step 2. Wirte each individual recipe to a tab delimited file.

```{r}
# Write all data frames into tab delimited file
write.table(Chocolate_Bavarian_Cream_36,"Chocolate_Bavarian_Cream.1936.tab", sep="\t",row.names = FALSE)
write.table(Chocolate_Bavarian_Cream_06,"Chocolate_Bavarian_Cream.2006.tab", sep="\t",row.names = FALSE)
write.table(Drop_Sugar_Cookies_36,"Drop_Sugar_Cookies.1936.tab", sep="\t",row.names = FALSE)
write.table(Drop_Sugar_Cookies_06,"Drop_Sugar_Cookies.2006.tab", sep="\t",row.names = FALSE)
```

# Step 3. Read in each tab file, and merger the tables with the example Beef Stroganoff tables.

```{r}
# read tab files
files <- list.files(pattern="*.tab")
Beef1 <- read.table(files[1],header=TRUE) 
Beef2 <- read.table(files[2],header=TRUE) 
Chocolate1 <- read.table(files[3],header=TRUE)
Chocolate2 <- read.table(files[4],header=TRUE) 
Sugar1 <- read.table(files[5],header=TRUE) 
Sugar2 <- read.table(files[6],header=TRUE) 

# Merge the same recipe of 1936, 2006 by ingredient
Beef_merge <- merge(Beef1,Beef2,by="Ingredient",all=TRUE)
Chocolate_merge <- merge(Chocolate1,Chocolate2,by="Ingredient",all=TRUE)
Sugar_merge <- merge(Sugar1,Sugar2,by="Ingredient",all=TRUE)
```

Comment: The two recipes are merged by ingredience so that the Measure.1997 and measure.2006 for beef recipes can be ploted to see where the measures are not consistent. 

# Step 4. Proof the data to check for erros. 

```{r}
# Check the consistancy or inconsistancy for measures between two different years
table(Beef_merge$Measure.x,Beef_merge$Measure.y)
table(Chocolate_merge$Measure.x,Chocolate_merge$Measure.y)
table(Sugar_merge$Measure.x,Sugar_merge$Measure.y)
```

Comments: In these contingency tables, the mismatches in measures between ingredients are displayed. 

# Step 5. Read FOOD_DES.txt. Perform partial matching in FOOD_DES and pick the best match 

```{r}
# read FOOD_DES.txt
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc","ComName","ManufacName","Survey","Ref_desc",
                         "Refuse","SciName","N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")

# list all unique ingredient in all tables
unique_ingredient <- unique(c(as.vector(Beef_merge$Ingredient),as.vector(Chocolate_merge$Ingredient),as.vector(Sugar_merge$Ingredient)))
print(unique_ingredient)
```

Seache some keywords using a function for these unique ingredient above.

```{r}
# seach the key word matched in FOOD_DES file
match <- function(x){
  food_des.dat[grep(x, food_des.dat$Long_Desc), c(1,3)]
}


# Type-in the key words abvoe
match("Water, bottled, generic")

```

Modify the name and find the NDB number to rename all of Ingredient and code.

```{r}
# create a new tab
newtab <- data.frame()
for (i in 1:length(files)){
  file_name <- unlist(strsplit(files[i],".",fixed=TRUE)) # load all files
  tab <- read.table(files[i],header=TRUE) # read each table
  tab <- cbind(Recipe=c(rep(file_name[1],nrow(tab))),Year=c(rep(file_name[2],nrow(tab))),tab) # use cbind to combine all tabs together
  newtab <- as.data.frame(rbind(newtab,tab)) #create a data frame
}

# Follow the Intredient order in the new tab and give them names and NDBs based on the serach reustls

Ingredient <- c('Beef, loin, tenderloin roast, boneless, separable lean and fat, trimmed to 0" fat, all grades, raw',
                'Butter, salted',
                'Wheat flour, white, all-purpose, self-rising, bleached',
                'Soup, stock, beef, home-prepared',
                'Cream, sour, cultured',
                'Mustard, prepared, yellow',
                'Beef, loin, tenderloin roast, boneless, separable lean and fat, trimmed to 0" fat, all grades, raw',
                'Oil, canola',
                'Butter, salted',
                'Onions, raw',
                'Mushrooms, white, raw',
                'Soup, stock, beef, home-prepared',
                'Cream, sour, cultured',
                'Mustard, prepared, yellow',
                'Gelatins, dry powder, unsweetened',
                'Water, bottled, generic',
                'Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D',
                'Sugars, granulated',
                'Vanilla extract',
                'Cream, fluid, heavy whipping',
                'Candies, chocolate, dark, NFS (45-59% cacao solids 90%; 60-69% cacao solids 5%; 70-85% cacao solids 5%)',
                'Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D',
                'Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D',
                'Gelatins, dry powder, unsweetened',
                'Candies, chocolate, dark, NFS (45-59% cacao solids 90%; 60-69% cacao solids 5%; 70-85% cacao solids 5%)',
                'Sugars, granulated',
                'Vanilla extract',
                'Cream, fluid, heavy whipping',
                'Sugars, granulated',
                'Butter, salted',
                'Egg, whole, raw, fresh',
                'Vanilla extract',
                'Wheat flour, white, bread, enriched',
                'Cream, sour, cultured',
                'Nuts, pecans',
                'Sugars, granulated',
                'Wheat flour, white, all-purpose, self-rising, bleached',
                'Sugars, granulated',
                'Oil, canola',
                'Egg, whole, raw, fresh',
                'Vanilla extract',
                'Sugars, granulated')
NDB <- c('23336',
         '1001',
         '20082',
         '6170',
         '1056',
         '2046',
         '23336',
         '4582',
         '1001',
         '11282',
         '11260',
         '6170',
         '1056',
         '2046',
         '19177',
         '14555',
         '1211',
         '19335',
         '2050',
         '1053',
         '19905',
         '1211',
         '1211',
         '19177',
         '19905',
         '19335',
         '2050',
         '1053',
         '19335',
         '1001',
         '1123',
         '2050',
         '20083',
         '1056',
         '12142',
         '19335',
         '20082',
         '19335',
         '4582',
         '1123',
         '2050',
         '19335')
mat <- data.frame(Ingredient=Ingredient,NDB_No=NDB)
finalRecipe <- cbind(newtab[,1:4],mat)

# modified table is wrote into a tab file
write.table(finalRecipe,"final.tab", sep="\t",row.names = FALSE)
```

Comments: The total rows in teh final recipe have 42 rows. Each intredient and NDB was modified row by row manually.

# Step 6. Create a Recipe table with servings information

```{r}
# read the Recipes table
Beef.recipe <- read.csv("Recipes.csv",header=TRUE)

# create a data frame for my designed two recipes with min and max servings infromation
recipe <- data.frame(Recipe=c("Chocolate_Bavarian_Cream","Chocolate_Bavarian_Cream","Drop_Sugar_Cookies","Drop_Sugar_Cookies"),Year=c("1936","2006","1936","2006"),
                     MinServings=c("8","8","40","60"),MaxServings=c("8","8","40","60"))   # min and max servinhs based on pdf file

# combine above two dataframes together
all.recipe <- rbind(Beef.recipe,recipe)

#write a csv
write.csv(all.recipe,"Recipe.csv",row.names = FALSE)
```

