---
title: "Midterm Project"
author: "Set 4"
date: "7/21/2018"
output:
  pdf_document:
    toc: 1 
    df_print: kable
---
###Notes: I used the kable function for pdf knit to print more legible data frame tables.

*import WEIGHT and FOOD_DES tables downloaded from the USDA website*
```{r}
PathToWEIGHT = "WEIGHT.txt"
USDA_WEIGHT <- read.table(PathToWEIGHT, 
                          sep="^", 
                          quote="~",
                          header=FALSE, 
                          stringsAsFactors = FALSE)

names(USDA_WEIGHT) <- c("NDB_No", "Seq", "Amount", "Msre_Desc",
                        "Gm_Wgt", "Num_Data_pts", "Std_Dev")

PathToFOOD_DES = "FOOD_DES.txt"
USDA_FOOD_DES <- read.table(PathToFOOD_DES, 
                            sep="^", 
                            quote="~",
                            header=FALSE, 
                            stringsAsFactors = FALSE)

names(USDA_FOOD_DES) <- c("NDB_No", "FdGrp_Cd", "Long_Desc", "Shrt_Desc","Com_Name",
                          "ManufacName", "Survey", "Ref_Desc", "Refuse", "Sci_Name", 
                          "N_Factor", "Pro_Factor_", "Fat_Factor_", "CHO_Factor")

```

#Cottage Pudding 1936

*Create vectors for data frame.*
```{r}
Amount_CottPudd.1936 <- c(0.5,0.25,1,1,1.5,2,0.25,0.5)
Measure_CottPudd.1936 <- c("cup","cup" ,"large","tsp","cup","tsp","tsp", "cup")
Ingredient_CottPudd.1936 <- c("Sugars, granulated",
                              "Butter, salted",
                              "Egg, whole, raw, fresh","Vanilla extract",
                              "Wheat flour, white, bread, enriched",
                              "Leavening agents, baking powder, double-acting, straight phosphate",
                              "Salt, table",
                              "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D")
NDB_No.1936 <- {}
Gm_Wgt.1936 <- {}

#for loop to populate the ingredient index from USDA FOOD_DES table
for (i in 1:length(Ingredient_CottPudd.1936)) {
  x <- USDA_FOOD_DES[Ingredient_CottPudd.1936[i] == USDA_FOOD_DES$Long_Desc, "NDB_No"]
  NDB_No.1936 <- c(NDB_No.1936, x)
}

#for loop to populate the weight of ingredients from USDA WEIGHT table 
for (i in 1:length(NDB_No.1936)) {
  y <- USDA_WEIGHT[NDB_No.1936[i] == USDA_WEIGHT$NDB_No 
                   & Measure_CottPudd.1936[i] == USDA_WEIGHT$Msre_Desc, 
                   "Gm_Wgt"]
  Gm_Wgt.1936 <- c(Gm_Wgt.1936, y[1])
}

```

*Verify lengths of all vectors are equivalent.*
```{r}
length(Amount_CottPudd.1936)
length(Measure_CottPudd.1936)
length(Ingredient_CottPudd.1936)
length(NDB_No.1936)
length(Gm_Wgt.1936)
```
*Create data frame for Cottage Pudding 1936*
```{r}
#column Gm_Wgt was not included as this is not required for the midterm
CottagePudding.1936.dat <- data.frame(
  Amount = Amount_CottPudd.1936,
  Measure = Measure_CottPudd.1936,
  Ingredient = Ingredient_CottPudd.1936,
  NDB_No = NDB_No.1936
)
CottagePudding.1936.dat
```
*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
#match each ingredient in the USDA FOOD_Des table
CottagePudding.1936.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc

#for loop to match each measure (based on NDB_No) in the USDA WEIGHT table
for (i in 1:length(CottagePudding.1936.dat$Measure)) {
  idx <- USDA_WEIGHT[CottagePudding.1936.dat$NDB_No[i] == USDA_WEIGHT$NDB_No, "Msre_Desc"]
  print(CottagePudding.1936.dat$Measure[i] %in% idx)
}

```
*All ingredient and measure values return true and therefore are present in the USDA database.*

#Cottage Pudding 2006

*Create vectors for data frame.*
```{r}
Amount_CottPudd.2006 <- c(1.5,1,1.5,0.5,2,0.5,1,0.5,1)
Measure_CottPudd.2006 <- c("cup","cup","tsp","tsp","large","cup","tsp","cup","cup")
Ingredient_CottPudd.2006 <- c("Wheat flour, white, all-purpose, enriched, bleached",
                              "Sugars, granulated",
                              "Leavening agents, baking powder, double-acting, straight phosphate",
                              "Salt, table",
                              "Egg, whole, raw, fresh",
                              "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D",
                              "Vanilla extract",
                              "Butter, salted",
                              "Marmalade, orange")
NDB_No.2006 <- {}
Gm_Wgt.2006 <- {}


#for loop to populate the ingredient index from USDA FOOD_DES table
for (i in 1:length(Ingredient_CottPudd.2006)) {
  x <- USDA_FOOD_DES[Ingredient_CottPudd.2006[i] == USDA_FOOD_DES$Long_Desc, "NDB_No"]
  NDB_No.2006 <- c(NDB_No.2006, x)
}

#for loop to populate the weight of ingredients from USDA WEIGHT table 
for (i in 1:length(NDB_No.2006)) {
  y <- USDA_WEIGHT[NDB_No.2006[i] == USDA_WEIGHT$NDB_No 
                   & Measure_CottPudd.2006[i] == USDA_WEIGHT$Msre_Desc, 
                   "Gm_Wgt"]
  Gm_Wgt.2006 <- c(Gm_Wgt.2006, y[1])
}

```

*Verify lengths of vectors are equivalent.*
```{r}
length(Amount_CottPudd.2006)
length(Measure_CottPudd.2006)
length(Ingredient_CottPudd.2006)
length(NDB_No.2006)
length(Gm_Wgt.2006)
```
*Create data frame for Cottage Pudding 2006*
```{r}
#column Gm_Wgt was not included as this is not required for the midterm
CottagePudding.2006.dat <- data.frame(
  Amount = Amount_CottPudd.2006,
  Measure = Measure_CottPudd.2006,
  Ingredient = Ingredient_CottPudd.2006,
  NDB_No = NDB_No.2006
)
CottagePudding.2006.dat
```
*Notes: the recipe from 2006 makes vanilla extract and marmalade optional. I have included both. Furthermore, it calls for unsalted butter. John Becker (as I did) used salted butter in his data because unsalted butter cannot be found in the USDA database. Finally, the recipe uses either cake flour or all-purpose flour. I chose to use all-purpose flour as this is more readily available in stores.*

*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
#match each ingredient in the USDA FOOD_Des table
CottagePudding.2006.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc

#for loop to match each measure (based on NDB_No) in the USDA WEIGHT table
for (i in 1:length(CottagePudding.2006.dat$Measure)) {
  idx <- USDA_WEIGHT[CottagePudding.2006.dat$NDB_No[i] == USDA_WEIGHT$NDB_No, "Msre_Desc"]
  print(CottagePudding.2006.dat$Measure[i] %in% idx)
}

```
*All ingredient and measure values return true and therefore are present in the USDA database.*

*Before exporting, we will check for errors and proofread. Start with by merging the 1936 and 2006 tables.*
```{r}
merge_CottPudd <- merge(CottagePudding.1936.dat[,c("Ingredient","Amount","Measure")],
                        CottagePudding.2006.dat[,c("Ingredient","Amount","Measure")], 
                        by="Ingredient",
                        all=TRUE)
merge_CottPudd
```
*Marmalade was added and the type of flour changed in the 2006 recipe.*

*Create a contingency table for Measure.*
```{r}
table(merge_CottPudd$Measure.x,merge_CottPudd$Measure.y)
```
*Measure units are the same for both years.*

*A graphical representation of the measures from 1936 and 2006 confirms this finding.*
```{r}
plot(merge_CottPudd$Measure.x,merge_CottPudd$Measure.y, xlab="1936", ylab="2006")
```
*Create a contingency table for Amount*
```{r}
table(merge_CottPudd$Amount.x,merge_CottPudd$Amount.y)
```
*There are increases in the amounts but nothing significant. After checking the recipes, these amounts did in fact increase.*

*A graphical representation of the amounts from 1936 and 2006 confirms this finding.*
```{r}
plot(merge_CottPudd$Amount.x,merge_CottPudd$Amount.y, xlab="1936", ylab="2006", col="blue")
```
*Since we are confident that these tables are error free, we output both tables to tab file, using* recipe.year.tab *as the naming convention.*
```{r}
write.table(CottagePudding.1936.dat, "CottagePudding.1936.tab") 
write.table(CottagePudding.2006.dat, "CottagePudding.2006.tab")
```

#Chantilly Potatoes 1936

*Create vectors for data frame.*
```{r}
Amount_ChantPot.1936 <- c(6,3,1,0.33333,0.5,0.5)
Measure_ChantPot.1936 <- c("medium (2-1/4\" to 3-1/4\" dia.)","tbsp","tsp","cup",
                           "cup, fluid (yields 2 cups whipped)","cup, shredded")
Ingredient_ChantPot.1936 <- c("Potatoes, boiled, cooked without skin, flesh, without salt", 
                              "Butter, salted",
                              "Salt, table",
                              "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D",
                              "Cream, fluid, heavy whipping",
                              "Cheese, cheddar (Includes foods for USDA's Food Distribution Program)")
NDB_No.1936 <- {}
Gm_Wgt.1936 <- {}

#for loop to populate the ingredient index from USDA FOOD_DES table
for (i in 1:length(Ingredient_ChantPot.1936)) {
  x <- USDA_FOOD_DES[Ingredient_ChantPot.1936[i] == USDA_FOOD_DES$Long_Desc, "NDB_No"]
  NDB_No.1936 <- c(NDB_No.1936, x)
}

#for loop to populate the weight of ingredients from USDA WEIGHT table 
for (i in 1:length(NDB_No.1936)) {
  y <- USDA_WEIGHT[NDB_No.1936[i] == USDA_WEIGHT$NDB_No 
                   & Measure_ChantPot.1936[i] == USDA_WEIGHT$Msre_Desc, 
                   "Gm_Wgt"]
  Gm_Wgt.1936 <- c(Gm_Wgt.1936, y[1])
}

```

*Verify lengths of vectors are equivalent.*
```{r}
length(Amount_ChantPot.1936)
length(Measure_ChantPot.1936)
length(Ingredient_ChantPot.1936)
length(NDB_No.1936)
length(Gm_Wgt.1936)
```
*Create data frame for Chantilly Potatoes 1936*
```{r}
#column Gm_Wgt was not included as this is not required for the midterm
ChantillyPotatoes.1936.dat <- data.frame(
  Amount = Amount_ChantPot.1936,
  Measure = Measure_ChantPot.1936,
  Ingredient = Ingredient_ChantPot.1936,
  NDB_No = NDB_No.1936
)
ChantillyPotatoes.1936.dat
```
*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
#match each ingredient in the USDA FOOD_Des table
ChantillyPotatoes.1936.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc

#for loop to match each measure (based on NDB_No) in the USDA WEIGHT table
for (i in 1:length(ChantillyPotatoes.1936.dat$Measure)) {
  idx <- USDA_WEIGHT[ChantillyPotatoes.1936.dat$NDB_No[i] == USDA_WEIGHT$NDB_No, "Msre_Desc"]
  print(ChantillyPotatoes.1936.dat$Measure[i] %in% idx)
}
```
*All ingredient and measure values return true and therefore are present in the USDA database.*

#Chantilly Potatoes 2006

*Create vectors for data frame.*
```{r}
Amount_ChantPot.2006 <- c(6,3,1,0.33333,0.5,0.5)
Measure_ChantPot.2006 <- c("medium (2-1/4\" to 3-1/4\" dia.)","tbsp", "tsp","cup",
                           "cup, fluid (yields 2 cups whipped)","cup")
Ingredient_ChantPot.2006 <- c("Potatoes, boiled, cooked without skin, flesh, without salt",
                              "Butter, salted",
                              "Salt, table",
                              "Milk, whole, 3.25% milkfat, without added vitamin A and vitamin D",
                              "Cream, fluid, heavy whipping",
                              "Cheese, parmesan, grated")
NDB_No.2006 <- {}
Gm_Wgt.2006 <- {}


#for loop to populate the ingredient index from USDA FOOD_DES table
for (i in 1:length(Ingredient_ChantPot.2006)) {
  x <- USDA_FOOD_DES[Ingredient_ChantPot.2006[i] == USDA_FOOD_DES$Long_Desc, "NDB_No"]
  NDB_No.2006 <- c(NDB_No.2006, x)
}

#for loop to populate the weight of ingredients from USDA WEIGHT table 
for (i in 1:length(NDB_No.2006)) {
  y <- USDA_WEIGHT[NDB_No.2006[i] == USDA_WEIGHT$NDB_No 
                   & Measure_ChantPot.2006[i] == USDA_WEIGHT$Msre_Desc, 
                   "Gm_Wgt"]
  Gm_Wgt.2006 <- c(Gm_Wgt.2006, y[1])
}

```

*Verify lengths of vectors are equivalent.*
```{r}
length(Amount_ChantPot.2006)
length(Measure_ChantPot.2006)
length(Ingredient_ChantPot.2006)
length(NDB_No.2006)
length(Gm_Wgt.2006)
```
*Create data frame for Chantilly Potatoes 2006*
```{r}
#column Gm_Wgt was not included as this is not required for the midterm
ChantillyPotatoes.2006.dat <- data.frame(
  Amount = Amount_ChantPot.2006,
  Measure = Measure_ChantPot.2006,
  Ingredient = Ingredient_ChantPot.2006,
  NDB_No = NDB_No.2006
)
ChantillyPotatoes.2006.dat
```
*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
#match each ingredient in the USDA FOOD_Des table
ChantillyPotatoes.2006.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc

#for loop to match each measure (based on NDB_No) in the USDA WEIGHT table
for (i in 1:length(ChantillyPotatoes.2006.dat$Measure)) {
  idx <- USDA_WEIGHT[ChantillyPotatoes.2006.dat$NDB_No[i] == USDA_WEIGHT$NDB_No, "Msre_Desc"]
  print(ChantillyPotatoes.2006.dat$Measure[i] %in% idx)
}

```
*All ingredient and measure values return true and therefore are present in the USDA database.*

*Before exporting, we will check for errors and proofread. Start with by merging the 1936 and 2006 tables.*
```{r}
merge_ChantPot <- merge(ChantillyPotatoes.1936.dat[,c("Ingredient","Amount","Measure")],
                        ChantillyPotatoes.2006.dat[,c("Ingredient","Amount","Measure")], 
                        by="Ingredient",
                        all=TRUE)
merge_ChantPot
```
*The only difference in ingredients is the type of cheese.*

*Create a contingency table for Measure.*
```{r}
table(merge_ChantPot$Measure.x,merge_ChantPot$Measure.y)
```
*Measure units are the same for both years.*

*A graphical representation of the measures between 1936 and 2006 confirms this finding.*
```{r}
plot(merge_CottPudd$Measure.x,merge_CottPudd$Measure.y, xlab="1936", ylab="2006")
```
*Create a contingency table for Amount.*
```{r}
table(merge_CottPudd$Amount.x,merge_CottPudd$Amount.y)
```
*There are increases and decreases in the amounts but nothing significant. After checking the recipes, these amounts did in fact change*

*A graphical representation of the amounts from 1936 and 2006 confirms this finding.*
```{r}
plot(merge_CottPudd$Amount.x,merge_CottPudd$Amount.y, xlab="1936", ylab="2006", col="blue")
```
*Since we are confident that these tables are error free, we output both tables to tab file, using* recipe.year.tab *as naming convention.*
```{r}
write.table(ChantillyPotatoes.1936.dat, "ChantillyPotatoes.1936.tab") 
write.table(ChantillyPotatoes.2006.dat, "ChantillyPotatoes.2006.tab") 
```

#Errors in the Beef Stroganoff files

*Import files.*
```{r}
PathToBeefStroganoff.1997 <- "BeefStroganoff.1997.tab"
BeefStroganoff.1997.dat <- read.table(PathToBeefStroganoff.1997, 
                                      sep="", 
                                      header=TRUE, 
                                      stringsAsFactors = TRUE)
BeefStroganoff.1997.dat

PathToBeefStroganoff.2006 <- "BeefStroganoff.2006.tab"
BeefStroganoff.2006.dat <- read.table(PathToBeefStroganoff.2006, 
                                      sep="", 
                                      header=TRUE, 
                                      stringsAsFactors = TRUE)
BeefStroganoff.2006.dat
```
*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
BeefStroganoff.1997.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc
BeefStroganoff.2006.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc
BeefStroganoff.1997.dat$Measure %in% USDA_WEIGHT$Msre_Desc
BeefStroganoff.2006.dat$Measure %in% USDA_WEIGHT$Msre_Desc
```
*All values in the Ingredient column return false so we must rephrase the ingredients to match the USDA database.*   
  
*Some of the Measure units also return false.*  
  
*When checking the specific Amounts in the recipe itself, there are some that do not match up with what is stated in the Joy of Cooking.*  
  
*Therefore, corrections in all 3 columns need to be made. Rather than replace data into above data frames, we will create new data frames with correct values.*

#Beef Stroganoff 1997 - Corrected

*Create vectors for data frame.*
```{r}
Amount_BeefStrog.1997 <- c(1.5,0.0625,24,2,1,0.1875,1,1)
Measure_BeefStrog.1997 <- c("tbsp" ,"cup","oz","tbsp" ,"medium (2-1/2\" dia)",
                            "cup","tsp or 1 packet","cup")
Ingredient_BeefStrog.1997 <- c("Butter, salted",
                               "Wheat flour, white, all-purpose, enriched, bleached",
                               "Beef, loin, tenderloin roast, boneless, separable lean and fat, trimmed to 0\" fat, all grades, raw",
                               "Butter, salted",
                               "Onions, raw",
                               "Cream, sour, cultured",
                               "Mustard, prepared, yellow",
                               "Soup, stock, beef, home-prepared")

NDB_No.1997 <- {}
Gm_Wgt.1997 <- {}

#for loop to populate the ingredient index from USDA FOOD_DES table
for (i in 1:length(Ingredient_BeefStrog.1997)) {
  x <- USDA_FOOD_DES[Ingredient_BeefStrog.1997[i] == USDA_FOOD_DES$Long_Desc, "NDB_No"]
  NDB_No.1997 <- c(NDB_No.1997, x)
}

#for loop to populate the weight of ingredients from USDA WEIGHT table 
for (i in 1:length(NDB_No.1997)) {
  y <- USDA_WEIGHT[NDB_No.1997[i] == USDA_WEIGHT$NDB_No 
                   & Measure_BeefStrog.1997[i] == USDA_WEIGHT$Msre_Desc, 
                   "Gm_Wgt"]
  Gm_Wgt.1997 <- c(Gm_Wgt.1997, y[1])
}
```

*Verify lengths of vectors are equivalent.*
```{r}
length(Amount_BeefStrog.1997)
length(Measure_BeefStrog.1997)
length(Ingredient_BeefStrog.1997)
length(NDB_No.1997)
length(Gm_Wgt.1997)
```
*Create data frame for Beef Stroganoff 1997*
```{r}
#column Gm_Wgt was not included as this is not required for the midterm
BeefStroganoff.1997.dat<-data.frame(
  Amount = Amount_BeefStrog.1997,
  Measure = Measure_BeefStrog.1997,
  Ingredient = Ingredient_BeefStrog.1997,
  NDB_No = NDB_No.1997
)
BeefStroganoff.1997.dat
```
*Notes: some of the measures were converted (lb for beef roast to oz, and tablespoon to cup for sour cream) in order to match USDA database.*

*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
#match each ingredient in the USDA FOOD_Des table
BeefStroganoff.1997.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc

#for loop to match each measure (based on NDB_No) in the USDA WEIGHT table
for (i in 1:length(BeefStroganoff.1997.dat$Measure)) {
  idx <- USDA_WEIGHT[BeefStroganoff.1997.dat$NDB_No[i] == USDA_WEIGHT$NDB_No, "Msre_Desc"]
  print(BeefStroganoff.1997.dat$Measure[i] %in% idx)
}
```
*All ingredient and measure values return true and therefore are present in the USDA database.*

#Beef Stroganoff 2006 - Corrected

*Create vectors for data frame.*
```{r}
Amount_BeefStrog.2006 <- c(32,2,3,1,5,2,1,1)
Measure_BeefStrog.2006 <- c("oz","tbsp","tbsp","medium (2-1/2\" dia)",
                            "cup, pieces or slices","cup","cup" ,"tsp or 1 packet")
Ingredient_BeefStrog.2006 <- c("Beef, loin, tenderloin roast, boneless, separable lean and fat, trimmed to 0\" fat, all grades, raw",
                               "Oil, canola",
                               "Butter, salted",
                               "Onions, raw","Mushrooms, white, raw",
                               "Soup, stock, beef, home-prepared",
                               "Cream, sour, cultured",
                               "Mustard, prepared, yellow")
NDB_No.2006 <- {}
Gm_Wgt.2006 <- {}

#for loop to populate the ingredient index from USDA FOOD_DES table
for (i in 1:length(Ingredient_BeefStrog.2006)) {
  x <- USDA_FOOD_DES[Ingredient_BeefStrog.2006[i] == USDA_FOOD_DES$Long_Desc, "NDB_No"]
  NDB_No.2006 <- c(NDB_No.2006, x)
}

#for loop to populate the weight of ingredients from USDA WEIGHT table 
for (i in 1:length(NDB_No.2006)) {
  y <- USDA_WEIGHT[NDB_No.2006[i] == USDA_WEIGHT$NDB_No 
                   & Measure_BeefStrog.2006[i] == USDA_WEIGHT$Msre_Desc, 
                   "Gm_Wgt"]
  Gm_Wgt.2006 <- c(Gm_Wgt.2006, y[1])
}
```

*Verify lengths of vectors are equivalent.*
```{r}
length(Amount_BeefStrog.2006)
length(Measure_BeefStrog.2006)
length(Ingredient_BeefStrog.2006)
length(NDB_No.2006)
length(Gm_Wgt.2006)
```
*Create data frame for Beef Stroganoff 2006*
```{r}
#column Gm_Wgt was not included as this is not required for the midterm
BeefStroganoff.2006.dat<-data.frame(
  Amount = Amount_BeefStrog.2006,
  Measure = Measure_BeefStrog.2006,
  Ingredient = Ingredient_BeefStrog.2006,
  NDB_No = NDB_No.2006
)
BeefStroganoff.2006.dat
```
*Notes: some of the measures were converted (lb for beef roast to oz, and tablespoon to cup for sour cream) in order to match USDA database.*

*To verify units and ingredients match the USDA database, search Msre_Desc column and Long_Desc columns.*
```{r}
#match each ingredient in the USDA FOOD_Des table
BeefStroganoff.2006.dat$Ingredient %in% USDA_FOOD_DES$Long_Desc

#for loop to match each measure (based on NDB_No) in the USDA WEIGHT table
for (i in 1:length(BeefStroganoff.2006.dat$Measure)) {
  idx <- USDA_WEIGHT[BeefStroganoff.2006.dat$NDB_No[i] == USDA_WEIGHT$NDB_No, "Msre_Desc"]
  print(BeefStroganoff.2006.dat$Measure[i] %in% idx)
}
```
*All ingredient and measure values return true and therefore are present in the USDA database.*

*Once again, we will check for errors and proofread. Start with by merging the 1997 and 2006 tables.*
```{r}
merge_BeefStrog <- merge(BeefStroganoff.1997.dat[,c("Ingredient","Amount","Measure")],
                         BeefStroganoff.2006.dat[,c("Ingredient","Amount","Measure")],
                         by="Ingredient",
                         all=TRUE)
merge_BeefStrog
```
*From the "NA" values, we see some ingredients were either added or dropped.*

*Create a contingency table for Measure.*
```{r}
table(merge_BeefStrog$Measure.x,merge_BeefStrog$Measure.y)
```
*Measure units are the same for both years. Ingredients that were dropped or added are also evident in this table.*

*A graphical representation of the measures from 1997 and 2006 confirms this finding.*
```{r}
plot(merge_BeefStrog$Measure.x,merge_BeefStrog$Measure.y, xlab="1997", ylab="2006")
```
*Create a contingency table for Amount.*
```{r}
table(merge_BeefStrog$Amount.x,merge_BeefStrog$Amount.y)
```
*There are increases in the amounts but nothing significant. After checking the recipes, these amounts did in fact change*

*A graphical representation of the amounts from 1997 and 2006 confirms this finding.*
```{r}
plot(merge_CottPudd$Amount.x,merge_CottPudd$Amount.y, xlab="1997", ylab="2006", col="blue")
```
*Since we are confident that these tables are error free, we output both tables to tab file, using* recipe.year.tab *as naming convention.*
```{r}
#PMC - other RMD can't be run if this file is changed.
#write.table(BeefStroganoff.1997.dat, "BeefStroganoff.1997.tab") 
#write.table(BeefStroganoff.2006.dat, "BeefStroganoff.2006.tab") 
```

#Merging Data

*Import all 6 tables and add columns for recipe name and year.*
```{r}
#vectorize a list of filenames
filename <- c("BeefStroganoff.1997.tab",
              "BeefStroganoff.2006.tab",
              "CottagePudding.1936.tab",
              "CottagePudding.2006.tab",
              "ChantillyPotatoes.1936.tab",
              "ChantillyPotatoes.2006.tab")

table.dat <- data.frame()
Recipe <- {}
Year <- {}

#for loop to accept each element of the filename vector
for (i in 1:length(filename)) {
  PathToFile <- filename[i]
  
  #create a running table of the data from each file
  y <- read.table(PathToFile, sep="", header=TRUE, stringsAsFactors = FALSE)
  table.dat <- rbind(table.dat,y)
  
  #parse the filename 
  x <- sapply(strsplit(filename[i], split="\\."),FUN=as.character)
  
  #create a running table for Recipe and Year
  Recipe <- cbind(c(Recipe,rep(x[1],length(y$Ingredient))))
  Year <- cbind(c(Year,rep(x[2],length(y$Ingredient))))
}

#combine into a data frame
JoyOfCooking.dat <- data.frame(Recipe=Recipe,Year=Year,table.dat, stringsAsFactors = FALSE)
JoyOfCooking.dat
```

#Second table - calories per serving and calories per gram

*Create vectors for data frame.*
```{r}
Recipe2 <- c("Beef Stroganoff","Beef Stroganoff",
             "Chanitilly_Potatoes","Chanitilly_Potatoes",
             "Cottage_Pudding","Cottage_Pudding")

Year2 <- c(1997,2006,
           1936,2006,
           1936,2006)

Min_Serving_Size <- c(4,4,6,6,6,8)
Max_Serving_Size <- c(6,6,6,8,6,10)

#these were taken from the recipe filess and will be used later
Gram_Per_Serving <- c(169.82,371.4,217.06,133.4,89.88,72.47)
Calories_Per_serving <- c(250.6,416.1,310,170.3,282.9,247.6)
Calories_Per_Gram <- c(1.48,1.12,1.28,1.28,3.15,3.42)

```

*Create data frame*
```{r}
Servings.dat <- data.frame(
  Recipe = Recipe2,
  Year = Year2,
  "Min Servings" = Min_Serving_Size,
  "Max Servings" = Max_Serving_Size,
  
  #will be filled in for the final
  Grams = NA,
  Calories = NA
)
Servings.dat
```
*Values were double checked in the recipe files for accuracy.*

*Output table to tab file.*
```{r}
write.table(Servings.dat, "Servings.tab")
```

*Import back in for verification.*
```{r}
PathToServings = "Servings.tab"
Servings.imported <- read.table(PathToServings, sep="", header=TRUE, stringsAsFactors = FALSE)
Servings.imported
```
*Re-import is successful.*