---
title: "Midterm Project Project"
author: "Set 53"
date: "7/21/2018"
output:
  pdf_document: default
  html_document: default
---

#Create Data table for Herring Salad and Sukiyaki recipes: Step - 1


```{r}
HerringSalad.1936.dat<-data.frame (Amount=c(12.00,1.00,6 ,2 ,1.5 ,0.50 ,0.50 ,2.00 ,
                                            0.50 ,3.00 ,1.00 ,1.00 ,2.00 ),
Measure=c('lb', 'cup','oz','lrg','cup','cup','cup','ea','cup','cup','cup','cup','Tbs'),
Ingredient=c('herring, atlantic, raw, fillet','wine, red',
'veal, cooked','egg,raw','beets, pickled, caanned, sliced, in liquid',
'onion, yellow, fresh, chopped','pickles, dill, chopped',
'celery stalk, fresh, medium, 7.5\" to 8\" ','potato, cooked, peeled, diced',
'apple, refesh, sliced','almonds, dry roastd, whole, unsalted',
'sugar,  white, granulated, granulated',' horseradish, prepared'))

HerringSalad.1936.dat





```

```{r}
sukiyaki.2006.dat<-data.frame (
Amount=c(2.00,0.50,6.00,2.00,2.00,2.00,0.50,.025,0.50,1.00,3.00,4.00),
Measure=c('lb','cup','ea','cup','cup','cup','cup','cup','cup','tsp','Tbs','cup' ),
Ingredient=c('beef roast, tenderloin, raw, 0\" trim','onion, yellow, fresh, sliced',
'green onion, fresh, tops & bulb, medium,4 1/8\" ','mushrooms, fresh, sliced',
'cabbage, napa, fresh','bamboo shoots, canned, sliced, drained',
'tofu, regular, with calcium sulfate, 1/4 block','soy sauce, from soy & wheat',
'stock, chicken, prepared from recipe','sugar, white, granulated, granulated',
'oil, canola','rice, white, cooked, long grain, unriched')
)

sukiyaki.2006.dat
```

```{r}

sukiyaki.1936.dat<-data.frame (
Amount=c(0.25,2.50,2.50,2.00,20.00,2.00,2.00,0.50,0.25,2.00,5.00),
Measure=c('lb','lb','cup','cup','ea','ea','ea','lb','cup','Tbs','cup'),
Ingredient=c(
'suet, beef, raw','beef roast, tenderloin, raw, 0\" trim',
'stock, beef, prepared from recipe','bamboo shoots, canned, sliced, drained',
'green onion, fresh, tops & bulb, medium,4 1/8\" ',
'bell pepper, green, fresh, medium, 2 1/2\" ',
'celery stalk, fresh, medium, 7.5\" to 8\" ',
'mushrooms,fresh',
'soy sauce, from soy & wheat',
'sugar, white, granulated, granulated',
'rice, white, cooked, long grain, unriched'))

sukiyaki.1936.dat

```

```{r}
HerringSalad.2006.dat<-data.frame(
Amount=c(36.00,3.00,1.50,1.00,0.50,0.50,0.50,2,2.00,1.00,1.00,2.00),
Measure=c('oz','cup','cup','cup','cup','cup','cup','lrg','ea','cup','cup','Tbs'),
Ingredient=c('herring, atlantic, raw, fillet','apple, refesh, sliced',
'beets, pickled, caanned, sliced, in liquid',
'almonds, dry roastd, whole, unsalted',
'onion, yellow, fresh, chopped','pickles, dill, chopped',
'potato, cooked, peeled, diced','egg,raw',
'celery stalk, fresh, medium, 7.5\" to 8\" ',
'wine, red','sugar,  white, granulated, granulated','horseradish, prepared')
)

HerringSalad.2006.dat
```

```{r}

```


_My recipes with NDB_No for both herring salad and sukiyaki select from FOOD_DES.txt whereever not found exact match closest Long_Desc was chosen_



```{r}
my.HerringSalad.1936.dat<-data.frame (Amount=c(12.00,1.00,6 ,2 ,1.5 ,0.50 ,0.50 ,
                                               2.00 ,0.50 ,3.00 ,1.00 ,1.00 ,2.00 ),
Measure=c(
'lb', 
'cup',
'oz',
'lrg',
'cup',
'cup',
'cup',
'ea',
'cup',
'cup',
'cup',
'cup',
'Tbs'),
Ingredient=c(
"Fish, herring, Atlantic, raw" ,
"Vinegar, red wine",
"Veal, sirloin, separable lean and fat, cooked, braised",
"Egg, whole, raw, fresh",
"Beets, pickled, canned, solids and liquids",
"Onions, yellow, sauteed",
"Pickles, cucumber, dill or kosher dill",
"Celery, raw",
"Sweet potato, cooked, boiled, without skin",
"Apples, raw, without skin",
"Nuts, almonds",
"Sugars, granulated",
"Horseradish, prepared"),
NDB_No=c(15039,2068,17135,1123,11609,11286,11937,11143,11510,9004,12061,19335,2055)
)
my.HerringSalad.1936.dat




```

```{r}
my.sukiyaki.2006.dat<-data.frame (
Amount=c(2.00,0.50,6.00,2.00,2.00,2.00,0.50,.025,0.50,1.00,3.00,4.00),
Measure=c('lb','cup','ea','cup','cup','cup','cup','cup','cup','tsp','Tbs','cup' ),
Ingredient=c(
"Beef, round, tip round, roast, separable lean only, trimmed to 0\" fat, all grades, raw" ,
"Onions, yellow, sauteed",
"Onions, young green, tops only",
"Mushrooms, white, raw",
"Cabbage, raw",
"Bamboo shoots, canned, drained solids",
"Tofu, raw, regular, prepared with calcium sulfate",
"Soy sauce made from soy and wheat (shoyu)",
"Soup, stock, chicken, home-prepared" ,
"Sugars, granulated" ,
"Oil, canola",
"Rice, white, long-grain, parboiled, unenriched, cooked"),
NDB_No=c(23651,11286,11292,11260,11109,11028,16427,16123,6172,19335,4582,20447)
)
my.sukiyaki.2006.dat

```

```{r}

my.sukiyaki.1936.dat<-data.frame (
Amount=c(0.25,2.50,2.50,2.00,20.00,2.00,2.00,0.50,0.25,2.00,5.00),
Measure=c('lb','lb','cup','cup','ea','ea','ea','lb','cup','Tbs','cup'),
Ingredient=c(
"Beef, variety meats and by-products, suet, raw",
"Beef, round, tip round, roast, separable lean only, trimmed to 0\" fat, all grades, raw" ,
"Soup, stock, beef, home-prepared",
"Bamboo shoots, canned, drained solids",
"Onions, young green, tops only" ,
"Peppers, sweet, green, raw" ,
"Celery, raw",
"Mushrooms, white, raw" ,
"Soy sauce made from soy and wheat (shoyu)" , 
"Sugars, granulated" ,
"Rice, white, long-grain, parboiled, unenriched, cooked"),
NDB_N0=c(13335,23651,6170,11028,11292,11333,11143,11260,16123,19335,20447)
)

my.sukiyaki.1936.dat

```

```{r}

my.HerringSalad.2006.dat<-data.frame(
Amount=c(36.00,3.00,1.50,1.00,0.50,0.50,0.50,2,2.00,1.00,1.00,2.00),
Measure=c(
'oz',
'cup',
'cup',
'cup',
'cup',
'cup',
'cup',
'lrg',
'ea',
'cup',
'cup',
'Tbs'),
Ingredient=c(
"Fish, herring, Atlantic, raw" ,
"Apples, raw, without skin",
"Beets, pickled, canned, solids and liquids",
"Nuts, almonds",
"Onions, yellow, sauteed",
"Pickles, cucumber, dill or kosher dill",
"Sweet potato, cooked, boiled, without skin",
"Egg, whole, raw, fresh",
"Celery, raw",
"Vinegar, red wine",
"Sugars, granulated",
"Horseradish, prepared"),
NDB_No=c(15039,9004,11609,12061,11286,11937,11510,1123,11143,2068,19335,2055)
)

my.HerringSalad.2006.dat

```


#Write each individual recipe to a tab delimited file : Step - 2 

```{r}
#Adding columns Recipe and Year from the file name, initialized with 0
#Populating Recipe name and Year

recipeyear<-strsplit("HerringSalad.1936.dat","\\.")[[1]]
HerringSalad.1936.dat["Recipe"]<-0
HerringSalad.1936.dat["Year"]<-0
HerringSalad.1936.dat<-HerringSalad.1936.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
HerringSalad.1936.dat$Recipe <-recipeyear[1]
HerringSalad.1936.dat$Year<- recipeyear[2] 

recipeyear<-strsplit("HerringSalad.2006.dat","\\.")[[1]]
HerringSalad.2006.dat["Recipe"]<-0
HerringSalad.2006.dat["Year"]<-0
HerringSalad.2006.dat<-HerringSalad.2006.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
HerringSalad.2006.dat$Recipe <-recipeyear[1]
HerringSalad.2006.dat$Year<- recipeyear[2] 

recipeyear<-strsplit("sukiyaki.1936.dat","\\.")[[1]]
sukiyaki.1936.dat["Recipe"]<-0
sukiyaki.1936.dat["Year"]<-0
sukiyaki.1936.dat<-sukiyaki.1936.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
sukiyaki.1936.dat$Recipe <-recipeyear[1]
sukiyaki.1936.dat$Year<- recipeyear[2] 

recipeyear<-strsplit("sukiyaki.2006.dat","\\.")[[1]]
sukiyaki.2006.dat["Recipe"]<-0
sukiyaki.2006.dat["Year"]<-0
sukiyaki.2006.dat<-sukiyaki.2006.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
sukiyaki.2006.dat$Recipe <-recipeyear[1]
sukiyaki.2006.dat$Year<- recipeyear[2] 

write.table(HerringSalad.1936.dat,"HerringSalad.1936.tab",row.names = FALSE)
write.table(HerringSalad.2006.dat,"HerringSalad.2006.tab",row.names = FALSE)
write.table(sukiyaki.1936.dat,"sukiyaki.1936.tab",row.names = FALSE)
write.table(sukiyaki.2006.dat,"sukiyaki.2006.tab",row.names = FALSE)

```

```{r}
recipeyear<-strsplit("my.HerringSalad.1936.dat","\\.")[[1]]
my.HerringSalad.1936.dat["Recipe"]<-0
my.HerringSalad.1936.dat["Year"]<-0
my.HerringSalad.1936.dat<-my.HerringSalad.1936.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
my.HerringSalad.1936.dat$Recipe <-recipeyear[1]
my.HerringSalad.1936.dat$Year<- recipeyear[2] 

recipeyear<-strsplit("my.HerringSalad.2006.dat","\\.")[[1]]
my.HerringSalad.2006.dat["Recipe"]<-0
my.HerringSalad.2006.dat["Year"]<-0
my.HerringSalad.2006.dat<-my.HerringSalad.2006.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
my.HerringSalad.2006.dat$Recipe <-recipeyear[1]
my.HerringSalad.2006.dat$Year<- recipeyear[2] 

recipeyear<-strsplit("my.sukiyaki.1936.dat","\\.")[[1]]
my.sukiyaki.1936.dat["Recipe"]<-0
my.sukiyaki.1936.dat["Year"]<-0
my.sukiyaki.1936.dat<-my.sukiyaki.1936.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
my.sukiyaki.1936.dat$Recipe <-recipeyear[1]
my.sukiyaki.1936.dat$Year<- recipeyear[2] 

recipeyear<-strsplit("my.sukiyaki.2006.dat","\\.")[[1]]
my.sukiyaki.2006.dat["Recipe"]<-0
my.sukiyaki.2006.dat["Year"]<-0
my.sukiyaki.2006.dat<-sukiyaki.2006.dat[c('Recipe','Year','Amount','Measure','Ingredient')]
my.sukiyaki.2006.dat$Recipe <-recipeyear[1]
my.sukiyaki.2006.dat$Year<- recipeyear[2] 

write.table(my.HerringSalad.1936.dat,"my.HerringSalad.1936.tab",row.names = FALSE)
write.table(my.HerringSalad.2006.dat,"my.HerringSalad.2006.tab",row.names = FALSE)
write.table(my.sukiyaki.1936.dat,"my.sukiyaki.1936.tab",row.names = FALSE)
write.table(my.sukiyaki.2006.dat,"my.sukiyaki.2006.tab",row.names = FALSE)


```


# Read in each tab files and merge the tables : Step - 3

```{r}
PATHofHerringSalad.1936="my.HerringSalad.1936.tab"
PATHofHerringSalad.2006="my.HerringSalad.2006.tab"
PATHofsukiyaki.1936="my.sukiyaki.1936.tab"
PATHofsukiyaki.2006="my.sukiyaki.2006.tab"
PATHofBGS.1997="BeefStroganoff.1997.tab"
PATHofBGS.2006="BeefStroganoff.2006.tab"

PATHofHerringSalad.1936="HerringSalad.1936.tab"
PATHofHerringSalad.2006="HerringSalad.2006.tab"
PATHofsukiyaki.1936="sukiyaki.1936.tab"
PATHofsukiyaki.2006="sukiyaki.2006.tab"


BeefStroganoff.1997<-read.table(file=PATHofBGS.1997,header = TRUE)
BeefStroganoff.2006<-read.table(file=PATHofBGS.2006,header = TRUE)

HerringSalad.1936<-read.table(PATHofHerringSalad.1936,header = TRUE)
HerringSalad.2006<-read.table(PATHofHerringSalad.2006,header = TRUE)

sukiyaki.1936<-read.table(PATHofsukiyaki.1936,header = TRUE)
sukiyaki.2006<-read.table(PATHofsukiyaki.2006,header = TRUE)

my.HerringSalad.1936<-read.table(PATHofHerringSalad.1936,header = TRUE)
my.HerringSalad.2006<-read.table(PATHofHerringSalad.2006,header = TRUE)

my.sukiyaki.1936<-read.table(PATHofsukiyaki.1936,header = TRUE)
my.sukiyaki.2006<-read.table(PATHofsukiyaki.2006,header = TRUE)

#BeefStroganoff.1997
#sukiyaki.2006


```



# Merge with Beef Stroganoff i.e. stacking on the top of each other_

```{r}
#Stacking on the top of each other recipes
rbind(BeefStroganoff.1997,BeefStroganoff.2006,HerringSalad.1936[3:5],
      HerringSalad.2006[3:5],sukiyaki.1936[3:5],sukiyaki.2006[3:5])

#Merging recipes by Ingredient
merge(BeefStroganoff.1997,sukiyaki.1936,by="Ingredient")
merge(BeefStroganoff.2006,sukiyaki.2006,by="Ingredient")
merge(BeefStroganoff.1997,HerringSalad.1936,by="Ingredient")
merge(BeefStroganoff.2006,HerringSalad.2006,by="Ingredient")


```

#The data to check for errors proof reading 
```{r}
#Proof reading for Beef Stroganoff
merged_file<-merge(BeefStroganoff.1997,BeefStroganoff.2006,by="Ingredient")
merged_file
plot(merged_file$Amount.x,merged_file$Amount.y)
plot(merged_file$Measure.x,merged_file$Measure.y)
```


```{r}

#Finding the change in amount and measurements in Herring salad between 2006 and 1936 recipes
merged_file<-merge(my.HerringSalad.2006,my.HerringSalad.1936,by="Ingredient")
merged_file
plot(merged_file$Amount.x,merged_file$Amount.y)
plot(merged_file$Measure.x,merged_file$Measure.y)

```




```{r}
#Finding the change in amount and measurements in sukiyaki between 2006 and 1936 recipes
merged_file<-merge(my.sukiyaki.2006,my.sukiyaki.1936,by="Ingredient")
merged_file
plot(merged_file$Amount.x,merged_file$Amount.y)
plot(merged_file$Measure.x,merged_file$Measure.y)
```




_Reading recipe files contingency table_

```{r}
#Tables showing measurements in 1936 and 2006

table(sukiyaki.1936$Ingredient,sukiyaki.1936$Measure)
table(sukiyaki.2006$Ingredient,sukiyaki.2006$Measure)

table(HerringSalad.1936$Ingredient,HerringSalad.1936$Measure)
table(HerringSalad.2006$Ingredient,HerringSalad.2006$Measure)
   

```

_Ingredients changed since 1936 to 2006 in Sukiyaki _
```{r}
table(my.sukiyaki.1936$Ingredient,my.sukiyaki.1936$Measure)
table(my.sukiyaki.2006$Ingredient,my.sukiyaki.2006$Measure)

table(my.HerringSalad.1936$Ingredient,my.HerringSalad.1936$Measure)
table(my.HerringSalad.2006$Ingredient,my.HerringSalad.2006$Measure)
   
```


Added ingredients for sukiyaki on 2006 recipe:
 "onion, yellow, fresh, sliced"
 "cabbage, napa, fresh"
 "tofu, regular, with calcium sulfate, 1/4 block"
 "oil, canola"
and removed ingredients from 1936 recipe:
"suet, beef, raw"
"stock, beef, prepared from recipe"
"bell pepper, green, fresh, medium, 2 1/2\" "

_Ingredients changed since 1936 to 2006 in Herring Salad _
Only change in Herring salad since 1936 to 2006 is the removal of 
"veal, cooked"
from 2006 recipe ingredient.

```{r}
#Reading Recipe.csv and write back
PATH="Recipes.csv"
Recipes.csv<-read.csv(PATH,header = TRUE)
#PMC need to write back to CSV
write.csv(Recipes.csv,"Recipes.csv",row.names = FALSE)


```

#Reading FOOD_DES txt files
```{r}
PATH="FOOD_DES.txt"
food_des.dat<-read.table(PATH,header=FALSE, sep="^",quote="~")
names(food_des.dat)<-c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc","ComName",
                       "ManufacName","Survey","Ref_desc","Refuse","SciName",
                       "N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")


dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)

#listOfIngredient<-as.character(food_des.dat$Long_Desc[grep("Mushrooms, white, raw", 
#food_des.dat$Long_Desc)])
#listOfIngredient<-as.character(food_des.dat$Long_Desc[grep("Mushrooms*|mushroom*", 
#food_des.dat$Long_Desc)])
#as.character(listOfIngredient[grepl("Mushrooms, white, raw", listOfIngredient)])
#listOfIngredient
#for ( i in 1:nrow(food_des.dat)){
  #print (food_des.dat$NDB_No[i])
 # print (as.character(food_des.dat$Long_Desc[i]))
#}
#as.character(food_des.dat$Long_Desc[grep(as.character("^pickles*|Pickles*"), 
#food_des.dat$Long_Desc)])


```

#Reading from WEIGHT.txt 
```{r}
PATH="WEIGHT.txt"
weight.dat<-read.table(PATH,header=FALSE, sep="^",quote="~")
names(weight.dat)<-c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
#weight.dat

dim(food_des.dat)
head(food_des.dat)
summary(food_des.dat)
```


```{r}

#Capitalize first letter of the word
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
```
```{r}
#Separate factor into single word
getWord<-function(x){
  sub("^([[:alpha:]]*).*", "\\1", x)
}

```
```{r}

#Get NDB_No from the long description of FOOD_DES.txt
get.NDB_No<-function(X,food_des.dat){
  for (j in 1:length(x)){
    for (i in 1:nrow(food_des.dat)){
      if (food_des.dat$Long_Desc[i]==x[j]){
        print (food_des.dat$NDB_No[i])
        print (x[j])
      }
    }
  }
}
```


```{r}
y<-c("Beef, round, tip round, roast, separable lean only, trimmed to 0\" fat, all grades, raw" ,
"Onions, yellow, sauteed",
"Onions, young green, tops only",
"Mushrooms, white, raw",
"Cabbage, raw",
"Bamboo shoots, canned, drained solids",
"Tofu, raw, regular, prepared with calcium sulfate",
"Soy sauce made from soy and wheat (shoyu)",
"Soup, stock, chicken, home-prepared" ,
"Sugars, granulated" ,
"Oil, canola",
"Rice, white, long-grain, parboiled, unenriched, cooked")

x=c(
"Fish, herring, Atlantic, raw" ,
"Apples, raw, without skin",
"Beets, pickled, canned, solids and liquids",
"Nuts, almonds",
"Onions, yellow, sauteed",
"Pickles, cucumber, dill or kosher dill",
"Sweet potato, cooked, boiled, without skin",
"Egg, whole, raw, fresh",
"Celery, raw",
"Vinegar, red wine",
"Sugars, granulated",
"Horseradish, prepared")

```

