---
title: "Midterm"
author: "Set 57"
date: "7/20/2018"
output:
pdf_document: default
html_document: default
---

```{r}

# Add column of NBD_No to merged table. NBD_No can be used to cross-check with weight.dat table

CreamedEggswithAsparagus.1936.dat <- data.frame(
    Amount=c(5,1,4,4,2,0.5,2), 
    Measure=c('lrg', 'lb', 'Tbs', 'Tbs', 'cup', 'cup', 'Tbs'), 
    Ingredient=c('egg, raw', 
                'asparagus spears, canned, 5" long, drained',
                'butter, salted',
                'flour, all purpose, self rising, enriched',
                'creaming, whipping, heavy', 
                'bread crumbs, plain, dry, grated', 
                'butter, salted'),
    NDB_No = c(1123, 11015, 1001, 20082, 1053, 18079, 1001)
)

# Make sure to change directory to location where you want the file
write.table(CreamedEggswithAsparagus.1936.dat,
"CreamedEggswithAsparagus.1936.tab", sep=" ", row.names = FALSE)

CreamedEggswithAsparagus.2006.dat <- data.frame(
    Amount=c(1,1,0.25,4,2,2,1,1,0.5,0.75,1,0.75), 
    Measure=c('Tbs', 'Tbs', 'cup', 'lrg', 'Tbs', 'Tbs', 'cup', 'tsp', 'tsp', 'cup', 'Tbs', 'cup'), 
    Ingredient=c('butter, salted',
                'oil, olive', 
                'onion, yellow, fresh, chopped', 
                'egg, raw','butter, salted', 
                'flour, all purpose, self rising, enriched', 
                'milk, whole, 3.25%',
                'mustard, country dijon', 
                'garlic, minced', 
                'bread crumbs, plain, dry, grated', 
                'butter, salted', 
                'asparagus tips, fresh'),
    NDB_No = c(1001, 4053, 11282, 1123, 1001, 20082, 1077, 2046, 11215, 18079, 1001, 11015)
)

# Make sure to change directory to location where you want the file
write.table(CreamedEggswithAsparagus.2006.dat, "CreamedEggswithAsparagus.2006.tab", sep=" ", row.names = FALSE)

StickyToffeePudding.1936.dat <- data.frame(
    Amount=c(1, 0.25, 1, 1, 3, 1, 2, 0.5), 
    Measure=c('cup', 'cup', 'cup', 'cup', 'lrg', 'tsp', 'Tbs', 'Tbs'), 
    Ingredient=c('pecans, chopped', 
                'walntus, black, dried, chopped', 
                'dates, california, chopped', 
                'sugar, white, granulated, granulated', 
                'egg, raw',
                'extract, vanilla', 
                'milk, whole, 3.25%', 
                'flour, bread, white, enriched'),
    NDB_No = c(12142,12154, 9087, 19335, 1123, 2050, 1077, 20083)
)

# Make sure to change directory to location where you want the file
write.table(StickyToffeePudding.1936.dat, "StickyToffeePudding.1936.tab", sep=" ", row.names = FALSE)

StickyToffeePudding.2006.dat <- data.frame(
    Amount=c(1.5, 2, 1.25, 6, 3, 1.5), 
    Measure=c('cup', 'cup', 'cup', 'Tbs', 'lrg', 'tsp'), 
    Ingredient=c('dates, california, chopped',
                'flour, all purpose, self rising, enriched',
                'sugar, brown, packed', 
                'butter, salted', 
                'egg, raw', 
                'extract, vanilla'),
    NDB_No = c(9087, 20082, 19334, 1001, 1123, 2046)
)

write.table(StickyToffeePudding.1936.dat, "StickyToffeePudding.2006.tab", sep=" ", row.names = FALSE)

recipe <- data.frame(Recipe = c('Creamed Eggs with Asparagus', 'Creamed Eggs with Asparagus', 'Sticky Toffee Pudding', 'Sticky Toffee Pudding'), Year= c(1936,2006,1936,2006)
                     , MinServings = c(6,4,8,8), MaxServings = c(6,4,8,8))

BeefStroganoff.1997.dat <- data.frame(
  Amount = c(1.5, 3.5, 1.0, 1.0, 3.0, 1.0),
  Measure = c('lb','Tbs','Tbs','cup','Tbs','tsp'),
  Ingredient = c('beef roast, tenderloin, raw, 0" trim',
                 'butter, salted',
                 'flour, all purpose, self rising, enriched',
                 'stock, beef, prepared from recipe',
                 'sour cream, cultured',
                 'mustard, dijon'),
  NDB_No = c(23336, 1001, 20082, 6170, 1074, 2046)
)

# Replace the file with the updated version of the BeefStroganoff tables with the included NBD_No added 
write.table(BeefStroganoff.1997.dat, "BeefStroganoff.1997.tab", sep=" ", row.names = FALSE)

BeefStroganoff.2006.dat <- data.frame(
  Amount = c(2.0, 2.0, 3.0, 1.0, 1.0, 1.0, 3.0, 1.0),
  Measure = c('lb','Tbs','Tbs', 'ea', 'lb', 'cup','cup','Tbs'),
  Ingredient = c('beef roast, tenderloin, raw, 0" trim',
                 'oil, canola',
                 'butter, salted',
                 'onion, yellow, fresh, medium, 2 1/2',
                 'mushrooms, fresh',
                 'stock, beef, prepared from recipe',
                 'sour cream, cultured',
                 'mustard, dijon'),
  NDB_No = c(23336,4582, 1001, 11282, 11260, 6170, 1074, 2046)
)

# Replace the file with the updated version of the BeefStroganoff tables with the included NBD_No added
write.table(BeefStroganoff.2006.dat, "BeefStroganoff.2006.tab", sep=" ", row.names = FALSE)

```

This next piece of code runs a for loop of all the recipe.tab files and adds a column for the Recipe and the Year. Each file is then merged with the previous files. 

```{r}

# Make sure directory has been changed so that path is recognized
path = "./"
# Initialize empty list of recipes 
mergedRecipes.dat=c()
# Creates list of all files in the folder ending in .tab
file.names <- dir(path, pattern =".tab")
# Loop over files in list
for(i in 1:length(file.names))
{
  # Read in each file into a table
  recipeTable <- read.table(paste(path, file.names[i], sep="//"),header=TRUE)
  # use regex to sub out everything after the first period with "", which results in the title of the recipe
  recipeTable$Recipe<-gsub("\\..*","", file.names[i])
  # use regex to sub out everyting but the numbers with "", which should result in the year of the recipe
  recipeTable$Year <- gsub("[^0-9]", "", file.names[i])
  # merge the table with the previous table
  mergedRecipes.dat <- rbind(mergedRecipes.dat, recipeTable)
}

mergedRecipes.dat = mergedRecipes.dat[c("Recipe", "Year", "Ingredient", "NDB_No", "Amount", "Measure" )]

print(mergedRecipes.dat)

```

This part of the code reads in the 'WEIGHT.txt' and 'FOOD_DES.txt' files and renames the columns with the appropriate column names. 
```{r}
# Read in WEIGHT.txt file
weight.dat <- read.table("WEIGHT.txt", header=FALSE, sep="^", quote="~")
# Create appropriate column names 
names(weight.dat) <- c('NDB_No', 'Seq', 'Amount', 'Msre_Desc', 'Gm_Wgt', 'Num_Data_Pts', 'Std_Dev')

# Read in FOOD_DES.txt
food_des.dat <- read.table("FOOD_DES.txt", header=FALSE, sep="^", quote="~")
# Create appropriate column names
names(food_des.dat) <- c('NDB_No', 'FdGrp_Cd', 'Long_Desc', 'Shrt_Desc', 'ComName', 'ManufacName', 'Survey', 'Ref_desc', 'Refuse', 'SciName', 'N-Factor', 'Pro_Factor', 'Fat_Factor', 'CHO_Factor')

# Merge the recipes table with the 'food_des.dat' table 
mergedFood = merge(mergedRecipes.dat, food_des.dat, by.x = "NDB_No", by.y = "NDB_No")
mergedWeight = merge(mergedRecipes.dat, weight.dat, by.x = "NDB_No", by.y = "NDB_No")
 
# Shows that the mergedRecipes table can be cross-referenced with the food_des table 
mergedFood
mergedWeight
```
*By examining the mergedWeight table, it's clear that there are a number of Measures for each ingredient, and at times the measure used in the recipe is not a measure used in the NDB. For final analysis, it will be necessary to convert the measures used in the recipe. *


