---
title: "Midterm project"
author: "Set 64"
date: "6/25/2018"
output:
  word_document: default
  html_document: default
---

# Part 1 - Assign Recipes

First, generate a list of all files in the Recipes folder.
```{r,echo=FALSE}
#Downloaded the pdf files from the midterm project and imported with length we get the number of recieps
recipes <- list.files(path = "~/Downloads/reciepe", pattern = '*.pdf')
length(recipes)

```



Make this reproducible.
```{r,echo=FALSE}
#setting the seed
set.seed(10010)
```



Create two randomly sampled list of recipe files and cycle this list through 80 rows

```{r,echo=FALSE}
#randomly sampled the reciepss
recipe.pairs <- matrix(c(sample(recipes),sample(recipes)),ncol=2,nrow=108)
recipe.pairs
```

Are there any pairs across rows?

```{r,echo=FALSE}
#checking number of pairs across
sum(recipe.pairs[,1]==recipe.pairs[,2])
```

Write to a table

```{r,echo=FALSE}
#creating the samples into the csv files and exporting
write.csv(recipe.pairs, file="~/Downloads/Recipe.Table.csv",row.names = FALSE)
```

Read the table in and print:

```{r,echo=FALSE}
#reading the table exported and printing it
Recipe.Table <- read.csv("~/Downloads/Recipe.Table.csv",header=FALSE,skip=1)
print(Recipe.Table)
```


# Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.
```{r}

#using the nutirnt database to get the information of the foods and their nutrient vaues
weight.dat <- read.table("WEIGHT.txt",header=FALSE,sep="^",quote="~")
names(weight.dat) <- c("NDB_No","Seq","Amount","Msre_Desc","Gm_Wgt","Num_Data_Pts","Std_Dev")
head(weight.dat)

```

```{r}
#summary statistics of the nutrient data 
summary(weight.dat)
```

```{r}
#food _des data is imported is having the information of the foods and the nutrient values of the foods 
food_des.dat <- read.table("FOOD_DES.txt",header=FALSE,sep="^",quote="~")
names(food_des.dat) <- c("NDB_No","FdGrp_Cd","Long_Desc","Shrt_Desc",
"ComName","ManufacName","Survey","Ref_desc","Refuse","SciName",
"N_Factor","Pro_Factor","Fat_Factor","CHO_Factor")
dim(food_des.dat)

```

```{r}
#header gives to information of the first few rows of the table 
head(food_des.dat)
```

```{r}
#summary statustics

summary(food_des.dat)
```


```{r}
#nutrients data of the food have the nutrient values downoaded from the database 
nut_data.dat <- read.table("NUT_DATA.txt",header=FALSE,sep="^",quote="~")
names(nut_data.dat) <- c("NDB_No","Nutr_No","Nutr_Val","Num_Data_Pts","Std_Error",
"Src_Cd","Deriv_Cd","Ref_NDB_No","Add_Nutr_Mark","Num_Studies",
"Min","Max","DF","Low_EB","Up_EB","Stat_cmt","AddMod_Date")
head(nut_data.dat)

```

```{r}
summary(nut_data.dat)
```





```{r}
#Creating for the reciepes given to me swiss steak 1936 and scaulopped cauliflower,
#creating table for the swiss steak 1936
SwissSteak.1936.dat <- data.frame(
    Amount = c(2.00, 0.25, 0.25, 0.50, 2.00 , 3.00),
    Measure = c('lb','cup','cup','cup','cup','Tbs'),
    Ingredient = c('beefsteak, bottom round, raw, 1/8 trim , URMIS 1466','flour, allpurpose, selfrising,enriched', 'fat, bacon, cooked, rendered', 'onion, yellow, fresh,chopped', 'juice, tomato, canned , unsalted', 'flour, allpurpose, self rising,enriched')
)
SwissSteak.1936.dat

```
```{r}
##Creating for the reciepes given to me swiss steak2006
SwissSteak.2006.dat <- data.frame(
    Amount = c(2.00, 0.25,2.00, 0.5, 0.33, 0.33 , 0.33,1.00,1.00),
    Measure = c('lb','cup','Tbs','cup','cup','cup','cup', 'cup' ,'Tbs'),
    Ingredient = c('beef steak, bottom round, raw, 1/8 trim , URMIS 1466','flour, allpurpose, selfrising, enriched','oil ,canola', 'onion, yellow, fresh,chopped','carrots,fresh, chopped','bell-pepper, green, fresh,chopped', 'celery,fresh,diced','stock, beef, prepared from reciepe','flour, all purpose , self rising , enriched ')
)
SwissSteak.2006.dat

```



```{r}
##Creating for the reciepes given to me scalloped cauliflower
ScallopedCauliflower.1936.dat <- data.frame(
    Amount = c(1.00, 2.00, 1.50, 1.00, 0.25, 2.00),
    Measure = c('ea','Tbs','Tbs','cup','cup','Tbs'),
    Ingredient = c('cauliflower, fresh, medium head, 5 to 6','butter,salted','flour, allpurpose,self rising, enriched', 'milk, whole, 3.25%', 'bread crumbs, plain, dry, grated', 'butter, salted'  )
)
ScallopedCauliflower.1936.dat

```




```{r}
##Creating for the reciepes given to me scalloped cauliflower
ScallopedCauliflower.2006.dat <- data.frame(
    Amount = c(1.00, 4.00, 4.00, 2.00, 0.50,0.33),
    Measure = c('ea','Tbs','Tbs','cup','cup','cup'),
    Ingredient = c('cauliflower, fresh, medium head, 5 to 6','butter,salted','flour, allpurpose,self rising, enriched', 'milk, whole, 3.25%', 'bread crumbs, plain, dry, grated', 'cheese, parmesan, grated'  )
)
ScallopedCauliflower.2006.dat

```

```{r}
# grep to find out the closest of the ingredients and trying to replace with standrad ingredients but i was unable to do i get an error
grep('(S|s)teak', food_des.dat$Long_Desc, value = TRUE)

```

```{r}
# grep to find out the closest of the ingredients for the scualopped cauliflower 
grep('(C|c)auliflower', food_des.dat$Long_Desc, value = TRUE)

```

```{r}
#exporting the dataset to the tab swisssteak1936
write.table(SwissSteak.1936.dat, "SwissSteak.1936.tab", row.names=FALSE, sep=' \t')
```


```{r}
#exporting the dataset to the tab swisssteak2006
write.table(SwissSteak.2006.dat, "SwissSteak.2006.tab", row.names=FALSE, sep=' \t')
```


```{r}
#exporting the dataset to the tab ScallopedCauliflower1936
write.table(ScallopedCauliflower.1936.dat, "ScallopedCauliflower.1936.tab", row.names=FALSE, sep=' \t')
```


```{r}
#exporting the dataset to the tab ScallopedCauliflower2006
write.table(ScallopedCauliflower.2006.dat, "ScallopedCauliflower.2006.tab", row.names=FALSE, sep=' \t')
```

```{r}
#Reading the tab files exported to bring back and add to the reciepes files 
SwissSteak1936 <- read.table("SwissSteak.1936.tab", header = TRUE, sep = "\t")
SwissSteak1936
```

```{r}
#Reading the tab files exported to bring back and add to the reciepes files 
SwissSteak2006 <- read.table("SwissSteak.2006.tab", header = TRUE, sep = "\t")
SwissSteak2006
```
```{r}
#Reading the tab files exported to bring back and add to the reciepes files 
ScallopedCauliflower1936 <- read.table("ScallopedCauliflower.1936.tab", header = TRUE, sep = "\t")
ScallopedCauliflower1936
```



```{r}
#Reading the tab files exported to bring back and add to the reciepes files 
ScallopedCauliflower2006<- read.table("ScallopedCauliflower.2006.tab", header = TRUE, sep = "\t")
ScallopedCauliflower2006
```



```{r}
#Reading the tab files exported to bring back and add to the reciepes files 
ScallopedCauliflower2006<- read.table("ScallopedCauliflower.2006.tab", header = TRUE, sep = "\t")
ScallopedCauliflower2006
```


```{r}
#adding year and reciepe to the tab file for scalloped cauliflower
ScallopedCauliflower1936$Recipe <- "ScallopedCauliflower"
ScallopedCauliflower1936$Year <- 1936
ScallopedCauliflower1936
```
```{r}
#adding year and reciepe to the tab file for scalloped cauliflowe 2006
ScallopedCauliflower2006$Recipe <- "ScallopedCauliflower"
ScallopedCauliflower2006$Year <- 2006
ScallopedCauliflower2006

```

```{r}
#adding year and reciepe to the tab file for swiss steak 1936
SwissSteak1936$Recipe <- "SwissSteak"
SwissSteak1936$Year <- 1936
SwissSteak1936

```


```{r}
#adding year and reciepe to the tab file for swiss steak 2006
SwissSteak2006$Recipe <- "SwissSteak"
SwissSteak2006$Year <- 2006
SwissSteak2006

```

```{r}
#importing the beef strongoff reciepe into the R
BeefStroganoff2006<- read.table("BeefStroganoff.2006.tab", header = TRUE, sep = '\t')
BeefStroganoff2006
```
```{r}
#FROM THE RECIEPES TABLE PULLING INFORMATION ABOUT THE BEEF STRONGOFF
Recipes <- read.table("Recipes.csv", header = TRUE, sep = ',')
Recipes
```

```{r}
#Creating a swiss steak dataframe with minimum and maximum servings 
Swisssteak<- data.frame(
Recipe = c('SwissSteak','SwissSteak'),
Year = c('1936','2006'),
MinServings = c('4','4'),
MaxServings = c('4','4'))
Swisssteak
```
```{r}
#binding swiss steak with beef strongoff
Recipes<- rbind(Recipes, Swisssteak )
Recipes
```

```{r}

#creating a dataframe scalloped cauliflower with the maximum and minimum servings 
ScallopedCauliflower<- data.frame(
Recipe = c('ScallopedCauliflower','ScallopedCauliflower'),
Year = c('1936','2006'),
MinServings = c('4','4'),
MaxServings = c('4','4'))
ScallopedCauliflower
```
```{r}
#Sbinding the BeefStroganoff,swisssteak and scalloped caulifliwer
Recipes <- rbind(Recipes, ScallopedCauliflower )
Recipes
```
```{r}
#Comparison of amount of ingredients SwissSteak.1936 vs 2006
plot(SwissSteak.2006.dat$Amount, col="blue", main="Comparison of amount of ingredients SwissSteak.1936 vs 2006", xlab = "Ingredient", ylab = "Amount")
points(SwissSteak.1936.dat$Amount, col="green")

```


```{r}
#Comparison of amount of ingredients ScallopedCauliflower 1936 vs 2006
plot(ScallopedCauliflower.2006.dat$Amount, col="blue", main="Comparison of amount of ingredients ScallopedCauliflowern1936 vs 2006", xlab = "Ingredient ", ylab = "Amount")
points(ScallopedCauliflower.1936.dat$Amount, col="green")

```







