---
title: "Midterm Project"
author: "Set 67"
date: "July 21, 2018"
output: word_document
---

# Instructions

First, generate a list of all files in the Recipes folder.
```{r,eval=FALSE}
setwd("./")
recipes <- list.files(path = "./", pattern = '*.pdf')
```

Make this reproducible.
```{r,eval=FALSE}
set.seed(10010)
```

Create two randomly sampled list of recipe files and cycle this list through 80 rows

```{r,eval=FALSE}
recipe.pairs <- matrix(c(sample(recipes), sample(recipes)), ncol=2, nrow=80)
```

Are there any pairs across rows?

```{r,eval=FALSE}
sum(recipe.pairs[,1]==recipe.pairs[,2])
```

Write to a table

```{r,eval=FALSE}
write.csv(recipe.pairs, file="Recipe.Table.csv", row.names = FALSE)
```

Read the table in and print:

```{r,eval=FALSE}
Recipe.Table <- read.csv("Recipe.Table.csv", header=FALSE, skip=1)
print(Recipe.Table)
```

# Part 2 - Check Nutrient Data.

We will be using the USDA food nutrient database, found at https://www.ars.usda.gov/northeast-area/beltsville-md-bhnrc/beltsville-human-nutrition-research-center/nutrient-data-laboratory/docs/usda-national-nutrient-database-for-standard-reference/

Run this to make sure your system can read the required nutrient database tables.

```{r}
 weight <- read.table("WEIGHT.txt", header = F, sep = "^", quote = "~")
  names(weight) <- c("NDB_No", "Seq", "Amount", "Msre_Desc", "Gm_Wgt", "Num_Data_Pts", "Std_Dev")
  head(weight)
summary(weight)
```


```{r}
 food <- read.table("FOOD_DES.txt", header = F, sep = "^", quote = "~")
  names(food) <- c("NDB_No", "FdGrp_Cd", "Long_Desc", "Shrt_Desc", "ComName", "ManuFacName", "Survey", 
                     "Ref_desc", "Refuse",
                     "SciName", "N_Factor", "Pro_Factor", "Fat_Factor", "CHO_Factor")
  
 head(food)
summary(food)
```

# Part I

 ***Assigned Recipes: Tapioca Custard, Molded Custard**

# Create data tables for the required recipes for the respective years in the form Amount, Unit and Ingredient. Let Measure be Unit.

```{r setup, include=FALSE}
# creating data tables for the required recipes 

molded_custard_36 <- data.frame(Amount = c(1.00, 1.75, 2.00, 0.33, 1.00), 
                                Unit = c("Tbs", "cup", "lrg", "cup", "tsp"), 
                                Ingredient = c('gelatin, dry, unsweetened', 
                                                'milk, whole, 3.25%', 
                                                'egg, raw',
                                                'sugar, white, granulated, granulated',
                                                'extract, vanilla'))
molded_custard_36

molded_custard_06 <- data.frame(Amount = c(3.00, 2.25, 0.05, 3.00, 1.00), 
                                Unit = c("cup", "tsp", "cup", "lrg", "tsp"),
                                Ingredient = c('milk, whole, 3.25%', 
                                                'gelatin, dry, unsweetened', 
                                                'sugar, white, granulated, granulated', 
                                                'egg, raw',
                                                'extract, vanilla'))
molded_custard_06

tapioca_custad_36 <- data.frame(Amount = c(0.33, 0.50, 2.00, 4.00, 1.00),
                                Unit = c("cup", "cup", "lrg", "cup", "tsp"),
                                Ingredient = c('tapioca, quick cooking, dry',
                                                'sugar, white, granulated, granulated',
                                                'egg, raw',
                                                'milk, whole, 3.25%',
                                                'extract, vanilla'))
tapioca_custad_36

tapioca_custad_06 <- data.frame(Amount = c(2.50, 0.33, 3.00, 1.00),
                                Unit = c("cup", "cup", "Tbs", "tsp"),
                                Ingredient = c('milk, whole, 3.25%',
                                                'sugar, white, granulated, granulated',
                                                'tapioca, quick cooking, dry',
                                                'extract, vanilla'))
tapioca_custad_06

```

# Part II

# Next, write each of your assigned recipes to a tab delimited file. Make sure that you have your created delimited files name as a combination for the recipes name and year
 


```{r}
# writing a tab delimiter
write.table(molded_custard_36, "molded_custard_36.tab", sep="\t")
write.table(molded_custard_06, "molded_custard_06.tab", sep="\t")
write.table(tapioca_custad_36, "tapioca_custad_36.tab", sep="\t")
write.table(tapioca_custad_06, "tapioca_custad_06.tab", sep="\t")
```


# Part III

# Now, you need to read the tab delimited files and  merge the tables with the example Beef Stroganoff tables

```{r}
# reading data into R
BeefStroganoff.1997 <- read.table("BeefStroganoff.1997.tab", sep="", header = TRUE)
BeefStroganoff.1997
BeefStroganoff.2006 <- read.table("BeefStroganoff.2006.tab", sep = "", header = TRUE)
BeefStroganoff.2006
moldedCustard_36 <- read.table("molded_custard_36.tab", sep="", header = TRUE)
moldedCustard_36
moldedCustard_06 <- read.table("molded_custard_06.tab", sep="", header = TRUE)
moldedCustard_06
tapiocaCustad_36 <- read.table("tapioca_custad_36.tab", sep="", header = TRUE)
tapiocaCustad_36
tapiocaCustad_06 <- read.table("tapioca_custad_06.tab", sep="", header = TRUE)
tapiocaCustad_06

# Merging all files
merg <-Reduce(function(...) merge(..., all=TRUE, by="Ingredient"), 
           list(BeefStroganoff.1997, BeefStroganoff.2006,  moldedCustard_36, 
                moldedCustard_06, tapiocaCustad_36, tapiocaCustad_06))
merg
```

# Part IV

Show that the ingredients and units can be cross referenced into the WEIGHT and FOOD_DES table

```{r}
table(merg$Unit.x,merg$Unit.y)
plot(merg$Amount.x, merg$Amount.y)
```
